### 2025-01-11 presntations/talksforaha.Rmd
Update everything

---
title: "data"
output: html_document
date: "2024-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
# Load required libraries
library(ggplot2)
library(reshape2)
library(dplyr)
library(viridis)

Y_sum_df=readRDS("~/Dropbox (Personal)/icd10phe_lab.rds")

Y_sum_df=merge(Y_sum_df,ATM::disease_info_phecode_icd10[,c("phecode","phenotype","exclude_name")],by.x="names",by.y="phecode")
Y_sum_df$names=rownames(Y_sum)
Y_sum_df=data.frame(Y_sum)

# Function to plot disease events for a given set of individuals
plot_disease_events <- function(Y, individuals = 1:10, diseases = NULL, 
                              time_range = NULL, title = "Disease Events") {
  # Get dimensions
  N <- dim(Y)[1]
  D <- dim(Y)[2]
  T <- dim(Y)[3]
  # Default ranges if not specified
  if (is.null(diseases)) diseases <- 1:D
  if (is.null(time_range)) time_range <- 1:T

  # Convert tensor slice to long format
  data_long <- expand.grid(
    Individual = individuals,
    Disease = paste("Disease", diseases),
    Time = time_range
  ) %>%
    mutate(Event = as.vector(Y[individuals, diseases, time_range]))

  # Create the plot with improved visibility
  p <- ggplot(data_long, aes(x = Time, y = interaction(Disease, Individual))) +
    # Increase tile size and make events more prominent
    geom_tile(aes(fill = factor(Event)), width = 0.8, height = 0.8) +
    # Use a more contrasting color scheme
    scale_fill_manual(values = c("0" = "white", "1" = "darkred"),
                     name = "Event",
                     labels = c("No event", "Disease onset")) +
    theme_classic() +
    labs(title = title,
         x = "Time",
         y = "") +
    # Improve text and grid visibility
    theme(
      axis.text.y = element_text(size = 7),  # Smaller text for y-axis
      axis.text.x = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 14),
      panel.grid = element_blank(),  # Remove grid lines
      panel.background = element_rect(fill = "white"),  # Light grey background
      legend.position = "right",
      plot.margin = margin(10, 10, 10, 10)
    )

plot_summed_events <- function(Y, title = "Total Disease Events Over Time") {
  # Calculate sum across individuals
  Y_sum <- apply(Y, c(2,3), sum)

  # Convert to long format for ggplot
  data_long <- reshape2::melt(Y_sum) %>%
    dplyr::rename(
      Disease = Var1,
      Time = Var2,
      Count = value
    ) %>%
    dplyr::mutate(Disease = factor(Disease, levels = nrow(Y_sum):1))  # Reverse disease order for plotting

  # Create heatmap
  p <- ggplot(data_long, aes(x = Time, y = Disease)) +
    geom_tile(aes(fill = Count)) +
    scale_fill_viridis_c(option = "magma", direction = -1) +  # Use magma color palette
    theme_minimal() +
    labs(
      title = title,
      x = "Time",
      y = "Disease",
      fill = "Number of\nevents"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14),
      axis.text = element_text(size = 10),
      panel.grid = element_blank(),
      legend.position = "right",
      plot.margin = margin(10, 10, 10, 10)
    )

  return(p)
}

#Example usage:
Y_sum <- apply(Y2, c(2,3), sum)
p <- plot_summed_events(Y)
print(p)
ggsave("disease_events_sum.pdf", p, width = 12, height = 6, dpi = 300)

  return(p)
}

# Function to plot disease patterns over time
plot_disease_patterns <- function(Y, type = "onset") {
  # Get dimensions
  N <- dim(Y)[1]
  D <- dim(Y)[2]
  T <- dim(Y)[3]

  if (type == "onset") {
    # Calculate proportion of onsets at each time for each disease
    onset_props <- sapply(1:D, function(d) {
      colMeans(Y[,d,])
    })

    data_long <- melt(onset_props) %>%
      rename(Time = Var1, 
             Disease = Var2, 
             Proportion = value) %>%
      mutate(Disease = paste("Disease", Disease))

    p <- ggplot(data_long, aes(x = Time, y = Disease, fill = Proportion)) +
      geom_tile() +
      scale_fill_viridis(name = "Proportion\nof onsets") +
      theme_minimal() +
      labs(title = "Disease Onset Patterns",
           x = "Time",
           y = "Disease") +
      theme(panel.grid = element_blank())

  } else if (type == "cumulative") {
    # Calculate cumulative incidence
    cum_props <- sapply(1:D, function(d) {
      colMeans(t(apply(Y[,d,], 1, cumsum)))
    })

    data_long <- melt(cum_props) %>%
      rename(Time = Var1, 
             Disease = Var2, 
             Proportion = value) %>%
      mutate(Disease = paste("Disease", Disease))

    p <- ggplot(data_long, aes(x = Time, y = Disease, fill = Proportion)) +
      geom_tile() +
      scale_fill_viridis(name = "Cumulative\nincidence") +
      theme_minimal() +
      labs(title = "Cumulative Disease Patterns",
           x = "Time",
           y = "Disease") +
      theme(panel.grid = element_blank())
  }

  return(p)
}

# Function to plot disease co-occurrence within a time window
plot_disease_cooccurrence <- function(Y, window = 5) {
  N <- dim(Y)[1]
  D <- dim(Y)[2]

  # Calculate time to first event for each individual-disease pair
  first_events <- apply(Y, c(1,2), function(x) {
    event_times <- which(x == 1)
    if(length(event_times) > 0) min(event_times) else NA
  })

  # Calculate co-occurrence matrix
  cooc_matrix <- matrix(0, D, D)
  for(i in 1:N) {
    times <- first_events[i,]
    for(d1 in 1:D) {
      for(d2 in 1:D) {
        if(!is.na(times[d1]) && !is.na(times[d2])) {
          cooc_matrix[d1,d2] <- cooc_matrix[d1,d2] + 
            (abs(times[d1] - times[d2]) <= window)
        }
      }
    }
  }
  cooc_matrix <- cooc_matrix/N

  # Convert to long format for plotting
  data_long <- melt(cooc_matrix) %>%
    mutate(Var1 = paste("Disease", Var1),
           Var2 = paste("Disease", Var2))

  # Plot
  p <- ggplot(data_long, aes(x = Var1, y = Var2, fill = value)) +
    geom_tile() +
    scale_fill_viridis(name = paste0("Co-occurrence\nfrequency\n(within ", 
                                    window, " time units)")) +
    theme_minimal() +
    labs(title = "Disease Co-occurrence Patterns",
         x = "Disease 1",
         y = "Disease 2") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid = element_blank())

  return(p)
}

# Create a combined visualization
plot_disease_analysis <- function(Y, num_individuals = 10) {
  # Create all plots
  p1 <- plot_disease_events(Y, individuals = 1:num_individuals) +
    ggtitle("A) Individual Disease Events")

  p2 <- plot_disease_patterns(Y, type = "onset") +
    ggtitle("B) Disease Onset Patterns")

  p3 <- plot_disease_patterns(Y, type = "cumulative") +
    ggtitle("C) Cumulative Disease Patterns")

  p4 <- plot_disease_cooccurrence(Y) +
    ggtitle("D) Disease Co-occurrence")

  # Combine plots using cowplot
  library(cowplot)
  combined <- plot_grid(p1, p2, p3, p4, ncol = 2)

  return(combined)
}

# Example usage:
# Assuming Y is your N x D x T tensor of binary events

# For individual plots:
p1 <- plot_disease_events(Y)
p2 <- plot_disease_patterns(Y, type = "onset")
p3 <- plot_disease_patterns(Y, type = "cumulative")
p4 <- plot_disease_cooccurrence(Y)

# For combined visualization:
# p_all <- plot_disease_analysis(Y)

# To save:
# ggsave("disease_analysis.pdf", p_all, width = 15, height = 12)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)

# Function to find and plot most prevalent diseases
plot_top_diseases <- function(Y_sum_df, n_diseases = 5) {
  # Calculate total cases for each disease
  disease_totals <- Y_sum_df %>%
    select(-Group, -phenotype, -diag_icd10) %>%  # Remove non-numeric columns
    mutate(total = rowSums(.)) %>%
    bind_cols(Y_sum_df %>% select(phenotype, Group, diag_icd10)) %>%
    arrange(desc(total))

  # Get top n diseases
  top_diseases <- disease_totals %>%
    head(n_diseases)

  top_diseases=disease_totals[disease_totals$phenotype %in% c("Coronary athrosclerosis","Myocardial infarction","Heart failure NOS"),]

  # Reshape data for plotting
  plot_data <- top_diseases %>%
    select(-total) %>%
    pivot_longer(
      cols = matches("^\\d+$"),  # Columns that are just numbers
      names_to = "Age",
      values_to = "Count"
    ) %>%
    mutate(Age = as.numeric(Age))

  # Create plot
  p <- ggplot(plot_data, aes(x = Age, y = Count, fill = Pheno)) +
    geom_area(alpha = 0.6, position = "identity") +
    scale_fill_viridis(discrete = TRUE) +
    theme_minimal() +
    labs(
      title = paste("Age Distribution of Top", n_diseases, "Most Prevalent Diseases"),
      x = "Age of Onset",
      y = "Number of Cases",
      fill = "Disease"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14),
      legend.position = "bottom",
      legend.box = "horizontal",
      panel.grid.minor = element_blank(),
      axis.text = element_text(size = 10),
      legend.text = element_text(size = 8)
    )

  return(p)
}

# Function to plot diseases by group
plot_disease_groups <- function(Y_sum_df, selected_groups = NULL, n_per_group = 3) {
  # If no groups specified, use all groups
  if(is.null(selected_groups)) {
    selected_groups <- unique(Y_sum_df$Group)
  }

  # Calculate totals and get top diseases per group
  top_by_group <- Y_sum_df %>%
    filter(Group %in% selected_groups) %>%
    group_by(Group) %>%
    mutate(total = rowSums(select(., matches("^\\d+$")))) %>%
    arrange(desc(total)) %>%
    slice_head(n = n_per_group) %>%
    ungroup()

  # Reshape data for plotting
  plot_data <- top_by_group %>%
    select(-total) %>%
    pivot_longer(
      cols = matches("^\\d+$"),
      names_to = "Age",
      values_to = "Count"
    ) %>%
    mutate(
      Age = as.numeric(Age),
      Group = factor(Group),
      Pheno = factor(Pheno)
    )

  # Create faceted plot
  p <- ggplot(plot_data, aes(x = Age, y = Count, fill = Pheno)) +
    geom_area(alpha = 0.6) +
    facet_wrap(~Group, scales = "free_y") +
    scale_fill_viridis(discrete = TRUE) +
    theme_minimal() +
    labs(
      title = "Disease Onset Patterns by Group",
      x = "Age of Onset",
      y = "Number of Cases",
      fill = "Disease"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14),
      legend.position = "bottom",
      strip.text = element_text(size = 10),
      panel.grid.minor = element_blank(),
      axis.text = element_text(size = 8),
      legend.text = element_text(size = 7)
    )

  return(p)
}

# Example usage:
# 1. Plot top 5 most prevalent diseases overall
p1 <- plot_top_diseases(Y_sum_df, n_diseases = 5)

# 2. Plot top diseases by selected groups
selected_groups <- c("neoplasms", "infectious diseases", "cardiovascular")
p2 <- plot_disease_groups(Y_sum_df, 
                         selected_groups = selected_groups,
                         n_per_group = 3)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
# Function to plot diseases by group with proper data handling

plot_disease_groups <- function(Y_sum_df, selected_groups = NULL, n_per_group = 3) {
  # Ensure numeric columns are properly converted
  numeric_cols <- grep("^\\d+$", names(Y_sum_df), value = TRUE)

  # Calculate totals and get top diseases per group
  top_by_group <- Y_sum_df %>%
    # Ensure numeric columns are numeric
    mutate(across(all_of(numeric_cols), as.numeric)) %>%
    # Filter for selected groups
    filter(Group %in% selected_groups) %>%
    # Calculate totals
    group_by(Group) %>%
    mutate(total = rowSums(pick(all_of(numeric_cols)))) %>%
    arrange(desc(total)) %>%
    slice_head(n = n_per_group) %>%
    ungroup()

  # Reshape data for plotting
  plot_data <- top_by_group %>%
    select(-total) %>%
    pivot_longer(
      cols = all_of(numeric_cols),
      names_to = "Age",
      values_to = "Count"
    ) %>%
    mutate(
      Age = as.numeric(Age),
      Group = factor(Group),
      Pheno = factor(Pheno)
    )

  # Create faceted plot
  p <- ggplot(plot_data, aes(x = Age, y = Count, color = Pheno)) +
    geom_line(linewidth = 1.2) +
    facet_wrap(~Group, scales = "free_y", ncol = 1) +
    scale_color_viridis_d() +
    theme_minimal() +
    labs(
      title = "Disease Onset Patterns by Group",
      x = "Age of Onset",
      y = "Number of Cases",
      color = "Disease"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14),
      legend.position = "right",
      strip.text = element_text(size = 12, face = "bold"),
      panel.grid.minor = element_blank(),
      axis.text = element_text(size = 10),
      legend.text = element_text(size = 8)
    )

  return(p)
}

# Example usage:
selected_groups <- c("neoplasms", "digestive", "circulatory system")
p2 <- plot_disease_groups(Y_sum_df, 
                         selected_groups = selected_groups,
                         n_per_group = 3)

# To save the plot:
# ggsave("disease_groups.pdf", p2, width = 12, height = 10)
```

```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(viridis)
library(patchwork)

# Function to create density plot for top diseases within a group
create_group_density_plot <- function(data, group_name, n_top = 4) {
  # Get top n diseases in this group
  top_diseases <- data %>%
    filter(exclude_name == group_name) %>%
    group_by(phenotype) %>%
    summarise(
      n = n(),
      mean_age = mean(age_diag, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    arrange(desc(n)) %>%
    slice_head(n = n_top)

  # Filter data for just these diseases
  plot_data <- data %>%
    filter(exclude_name == group_name,
           phenotype %in% top_diseases$phenotype)

  # Create plot
  p <- ggplot(plot_data, aes(x = age_diag, fill = phenotype)) +
    geom_density(alpha = 0.5) +
    theme_minimal() +
    theme(
      strip.text = element_text(size = 10, face = "bold"),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 8),
      legend.position = "right",
      legend.text = element_text(size = 8),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)
    ) +
    scale_x_continuous(
      name = "Age of onset",
      limits = c(40, 80),
      breaks = seq(40, 80, by = 10)
    ) +
    scale_fill_viridis_d(
      option = "D",
      name = "Disease",
      labels = function(x) {
        counts <- top_diseases %>%
          filter(phenotype %in% x) %>%
          arrange(match(phenotype, x))
        paste0(x, "\n(n=", counts$n, ")")
      }
    ) +
    labs(
      title = paste("Top", n_top, "Diseases in", group_name),
      y = "Density"
    )

  # Add text with disease counts
  counts_text <- top_diseases %>%
    mutate(
      desc = paste0(phenotype, ": n=", n, 
                   " (mean age=", round(mean_age, 1), ")")
    ) %>%
    pull(desc) %>%
    paste(collapse = "\n")

  return(p)
}

# Get unique disease groups
groups <- unique(Y_sum_df$exclude_name)
groups=c("circulatory system","respiratory","digestive","neoplasms","neurological","congenital anomalies")

# Create plots for each group
plots <- list()
summaries <- list()

for(group in groups) {
  # Get summary for this group
  group_summary <- Y_sum_df %>%
    filter(exclude_name == group) %>%
    group_by(phenotype) %>%
    summarise(
      n = n(),
      mean_age = mean(age_diag, na.rm = TRUE),
      median_age = median(age_diag, na.rm = TRUE),
      sd_age = sd(age_diag, na.rm = TRUE)
    ) %>%
    arrange(desc(n)) %>%
    slice_head(n = 4)

  summaries[[group]] <- group_summary
  plots[[group]] <- create_group_density_plot(Y_sum_df, group)
}

# Combine all summaries into one dataframe
all_summaries <- bind_rows(summaries, .id = "disease_group")

# Save plots
for(i in seq_along(groups)) {
  ggsave(
    paste0("top_diseases_", groups[i], ".pdf"),
    plots[[i]],
    width = 6,
    height = 4,
    dpi = 300
  )
}
# Combined plot using patchwork
n_groups <- length(groups)
n_cols <- min(2, n_groups)
n_rows <- ceiling(n_groups / n_cols)

combined_plot <- wrap_plots(plots, ncol = n_cols)

ggsave(
  "all_groups_top_diseases.pdf",
  combined_plot,
  width = 10 * n_cols,
  height = 6 * n_rows,
  dpi = 300
)

# Save summary statistics
#write.csv(all_summaries, "top_diseases_summary.csv", row.names = FALSE)

# Print text summary
cat("\nTop 4 diseases by group:\n\n")
for(group in groups) {
  cat("\n", group, ":\n", sep="")
  group_data <- summaries[[group]]
  for(i in 1:nrow(group_data)) {
    cat(sprintf("%s: n=%d (mean age=%.1f)\n",
                group_data$phenotype[i],
                group_data$n[i],
                group_data$mean_age[i]))
  }
}
```

```{r}

# Define mapping between PRS columns and disease names
prs_mapping <- list(

    "AMD" = "Macular degeneration (senile) of retina NOS",
    "AD" = "Alzheimer's disease",
    "AST" = "Asthma",
    "AF" = "Atrial fibrillation",
    "BD" = "Bipolar",
    "BMI" = "Overweight, obesity and other hyperalimentation",
    "CRC" = "Colorectal cancer",
    "BC" = "Malignant neoplasm of female breast",

    "CED" = "Celiac disease",  
    "CAD" = "Coronary atherosclerosis",

    "EOC" = "Malignant neoplasm of ovary",
    "T1D" = "Type 1 diabetes",
    "T2D" = "Type 2 diabetes",
    "UC" = "Ulcerative colitis",
    "VTE" = "Other venous embolism and thrombosis",
    "RA" = "Rheumatoid arthritis",
    "SLE" = "Systemic lupus erythematosus",
    "SCZ" = "Schizophrenia",
    "MS" =  "Multiple sclerosis",
    "PD" = "Parkinson's disease"
)

# Function to standardize disease names
standardize_name <- function(x) {
    x <- tolower(x)
    x <- gsub("[[:punct:]]", "", x)
    x <- gsub("\\s+", "", x)
    return(x)
}

std_mapping <- lapply(prs_mapping, standardize_name)
Y_sum_df$std_phenotype <- standardize_name(Y_sum_df$phenotype)

# Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(pheatmap)
library(viridis)
library(broom)

# Read the data
Y_sum_df <- read.csv("Y_sum_df.csv")
prs_df <- read.table("prs_df", header=TRUE)

# Define mapping between PRS columns and disease names
prs_mapping <- list(
    "AAM" = "age at menopause",
    "AMD" = "age-related macular degeneration",
    "AD" = "alzheimer's disease",
    "AST" = "asthma",
    "AF" = "atrial fibrillation",
    "BD" = "bipolar disorder",
    "BMI" = "body mass index",
    "CRC" = "bowel cancer",
    "BC" = "breast cancer",
    "CVD" = "cardiovascular disease",
    "CED" = "coeliac disease",
    "CAD" = "coronary artery disease",
    "CD" = "crohn's disease",
    "EOC" = "epithelial ovarian cancer",
    "T1D" = "type 1 diabetes",
    "T2D" = "type 2 diabetes",
    "UC" = "ulcerative colitis",
    "VTE" = "venous thromboembolic disease",
    "RA" = "rheumatoid arthritis",
    "SLE" = "systemic lupus erythematosus",
    "SCZ" = "schizophrenia",
    "MS" = "multiple sclerosis",
    "PD" = "parkinson's disease"
)

# Function to standardize disease names
standardize_name <- function(x) {
    x <- tolower(x)
    x <- gsub("[[:punct:]]", "", x)
    x <- gsub("\\s+", "", x)
    return(x)
}

# Create standardized versions of both mapping and phenotype names
std_mapping <- lapply(prs_mapping, standardize_name)
Y_sum_df$std_phenotype <- standardize_name(Y_sum_df$phenotype)

# Function to analyze one matched pair
analyze_matched_pair <- function(pheno, prs_col, Y_data, prs_data) {
    # Get phenotype data
    pheno_data <- Y_data %>%
        filter(std_phenotype == standardize_name(pheno))

    if(nrow(pheno_data) < 20) return(NULL)

    # Merge with PRS data
    merged_data <- pheno_data %>%
        inner_join(prs_data %>%
                    select(!!sym(prs_col), `Identifier`) %>%
                    rename(eid = `Identifier`),
                  by = "eid")

    # Run regression
    model <- lm(age_diag ~ get(prs_col), data = merged_data)
    stats <- tidy(model) %>%
        filter(term != "(Intercept)")

    result <- stats %>%
        mutate(
            n = nrow(merged_data),
            phenotype = unique(pheno_data$phenotype)[1],
            prs = prs_col,
            mean_age = mean(merged_data$age_diag),
            sd_age = sd(merged_data$age_diag)
        )

    # Create violin plot split by PRS quintiles
    merged_data$prs_quintile <- cut(merged_data[[prs_col]], 
                                   breaks = quantile(merged_data[[prs_col]], probs = seq(0, 1, 0.2)),
                                   labels = c("Q1", "Q2", "Q3", "Q4", "Q5"),
                                   include.lowest = TRUE)

    p1 <- ggplot(merged_data, aes(x = get(prs_col), y = age_diag)) +
        geom_point(alpha = 0.3) +
        geom_smooth(method = "lm", color = "red") +
        theme_minimal() +
        labs(
            title = paste("Age of Onset vs PRS for", unique(pheno_data$phenotype)),
            subtitle = sprintf("n = %d\nEffect = %.2f years per SD (p = %.3g)",
                             result$n, result$estimate, result$p.value),
            x = "PRS Score (standardized)",
            y = "Age of Onset (years)"
        )

    p2 <- ggplot(merged_data, aes(x = prs_quintile, y = age_diag)) +
        geom_violin(fill = "lightblue", alpha = 0.5) +
        geom_boxplot(width = 0.2, fill = "white", alpha = 0.7) +
        theme_minimal() +
        labs(
            title = "Age of Onset by PRS Quintile",
            x = "PRS Quintile",
            y = "Age of Onset (years)"
        )

    return(list(stats = result, scatter = p1, violin = p2, data = merged_data))
}

# Analyze all matched pairs
results_list <- list()
plot_list <- list()

for(prs_col in names(prs_mapping)) {
    result <- analyze_matched_pair(
        prs_mapping[[prs_col]],
        prs_col,
        Y_sum_df,
        prs_df
    )

    if(!is.null(result)) {
        results_list[[prs_col]] <- result$stats
        plot_list[[prs_col]] <- list(scatter = result$scatter, 
                                    violin = result$violin)
    }
}

# Combine results
results_df <- bind_rows(results_list) %>%
    arrange(p.value)

# Save results
write.csv(results_df, "matched_prs_onset_results.csv", row.names = FALSE)

# Save plots for significant associations
for(prs in names(plot_list)[results_df$p.value < 0.05]) {
    # Save scatter plot
    ggsave(
        paste0("prs_scatter_", tolower(prs), ".pdf"),
        plot_list[[prs]]$scatter,
        width = 8,
        height = 6
    )

    # Save violin plot
    ggsave(
        paste0("prs_violin_", tolower(prs), ".pdf"),
        plot_list[[prs]]$violin,
        width = 8,
        height = 6
    )
}

# Create summary plot of effect sizes
p_effects <- ggplot(results_df, 
                   aes(x = reorder(phenotype, estimate), 
                       y = estimate)) +
    geom_point() +
    geom_errorbar(aes(ymin = estimate - 1.96*std.error,
                      ymax = estimate + 1.96*std.error)) +
    coord_flip() +
    theme_minimal() +
    labs(
        title = "PRS Effect Sizes on Age of Onset",
        subtitle = "Error bars show 95% confidence intervals",
        x = "Disease",
        y = "Effect Size (years per SD of PRS)"
    )

ggsave("prs_effects_summary.pdf", p_effects, width = 10, height = 8)

# Create heatmap for visualization
results_mat <- matrix(NA, 
                     nrow = 1, 
                     ncol = length(results_df$phenotype),
                     dimnames = list("Effect", results_df$phenotype))
results_mat[1,] <- results_df$estimate

# Add stars for significance
stars_mat <- matrix("", 
                   nrow = 1, 
                   ncol = length(results_df$phenotype),
                   dimnames = list("", results_df$phenotype))
stars_mat[1, results_df$p.value < 0.05] <- "*"

pdf("prs_effects_heatmap.pdf", width = 12, height = 4)
pheatmap(results_mat,
         display_numbers = stars_mat,
         cluster_cols = TRUE,
         cluster_rows = FALSE,
         color = colorRampPalette(c("blue", "white", "red"))(100),
         main = "PRS Effects on Age of Onset (years per SD)")
dev.off()

# Print summary
cat("\nResults summary (sorted by significance):\n")
print(results_df %>%
    select(phenotype, estimate, p.value, n) %>%
    mutate(
        estimate = round(estimate, 2),
        p.value = format.pval(p.value, digits = 2)
    ))

###
```

```{r}
library(dplyr)
library(ggplot2)
library(quantreg)
library(viridis)
library(patchwork)

# Function to analyze early vs late onset effects
analyze_onset_patterns <- function(pheno, prs_col, Y_data, prs_data) {
    # Standardize names and merge data
    std_pheno <- tolower(gsub("[[:punct:]]|\\s+", "", pheno))
    pheno_data <- Y_data %>%
        mutate(std_phenotype = tolower(gsub("[[:punct:]]|\\s+", "", phenotype))) %>%
        filter(std_phenotype == std_pheno)

    if(nrow(pheno_data) < 100) return(NULL)  # Need enough cases for quantile analysis

    # Merge with PRS data
    merged_data <- pheno_data %>%
        inner_join(prs_data %>%
                    select(!!sym(prs_col), `Identifier`) %>%
                    rename(eid = `Identifier`),
                  by = "eid")

    # Define early onset threshold (25th percentile of age)
    early_threshold <- quantile(merged_data$age_diag, 0.25)
    late_threshold <- quantile(merged_data$age_diag, 0.75)

    # Add onset category
    merged_data <- merged_data %>%
        mutate(
            onset_cat = case_when(
                age_diag <= early_threshold ~ "Early",
                age_diag >= late_threshold ~ "Late",
                TRUE ~ "Middle"
            )
        )

    # Fit quantile regression for different percentiles
    taus <- c(0.1, 0.25, 0.5, 0.75, 0.9)
    qr_fits <- lapply(taus, function(tau) {
        fit <- rq(age_diag ~ get(prs_col), tau = tau, data = merged_data)
        coef(summary(fit))[2,]
    })
    qr_results <- do.call(rbind, qr_fits)
    rownames(qr_results) <- paste0("Q", taus*100)

    # Regular regression for comparison
    lm_fit <- lm(age_diag ~ get(prs_col), data = merged_data)

    # Create visualizations

    # 1. Scatter plot with separate regression lines for early/late
    p1 <- ggplot(merged_data, aes(x = get(prs_col), y = age_diag, color = onset_cat)) +
        geom_point(alpha = 0.3) +
        geom_smooth(method = "lm", se = FALSE) +
        scale_color_viridis_d() +
        theme_minimal() +
        labs(
            title = paste("Age of Onset vs", prs_col, "for", pheno),
            subtitle = paste("n =", nrow(merged_data)),
            x = "PRS Score (standardized)",
            y = "Age of Onset",
            color = "Onset Category"
        )

    # 2. Quantile regression effects plot
    qr_df <- data.frame(
        tau = taus,
        estimate = qr_results[,1],
        lower = qr_results[,1] - 1.96*qr_results[,2],
        upper = qr_results[,1] + 1.96*qr_results[,2]
    )

    p2 <- ggplot(qr_df, aes(x = tau, y = estimate)) +
        geom_line() +
        geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
        geom_hline(yintercept = coef(lm_fit)[2], linetype = "dashed", color = "red") +
        theme_minimal() +
        labs(
            title = "PRS Effect Size by Age Quantile",
            x = "Age Quantile",
            y = "Effect Size (years per SD of PRS)"
        )

    # Calculate effect sizes for early vs late onset
    early_effect <- coef(lm(age_diag ~ get(prs_col), 
                           data = filter(merged_data, onset_cat == "Early")))[2]
    late_effect <- coef(lm(age_diag ~ get(prs_col), 
                          data = filter(merged_data, onset_cat == "Late")))[2]

    return(list(
        plots = list(scatter = p1, quantile = p2),
        effects = data.frame(
            phenotype = pheno,
            prs = prs_col,
            early_effect = early_effect,
            late_effect = late_effect,
            ratio = early_effect/late_effect,
            qr_effects = list(qr_results),
            n = nrow(merged_data)
        )
    ))
}

# Analyze all matched pairs
results_list <- list()
plot_list <- list()

for(prs_col in names(prs_mapping)) {
    result <- analyze_onset_patterns(
        prs_mapping[[prs_col]],
        prs_col,
        Y_sum_df,
        prs_df
    )

    if(!is.null(result)) {
        results_list[[prs_col]] <- result$effects
        plot_list[[prs_col]] <- result$plots
    }
}

# Combine results
results_df <- bind_rows(results_list)

# Create summary visualization
p_summary <- ggplot(results_df, 
                   aes(x = reorder(phenotype, early_effect), 
                       y = early_effect)) +
    geom_point(aes(color = "Early Onset"), size = 3) +
    geom_point(aes(y = late_effect, color = "Late Onset"), size = 3) +
    coord_flip() +
    theme_minimal() +
    labs(
        title = "PRS Effects on Early vs Late Disease Onset",
        x = "Disease",
        y = "Effect Size (years per SD of PRS)",
        color = "Onset Type"
    )

# Save outputs
ggsave("prs_early_vs_late_effects.pdf", p_summary, width = 12, height = 8)

# Save individual plots for each disease
for(prs in names(plot_list)) {
    # Combined plot with patchwork
    combined_plot <- plot_list[[prs]]$scatter + plot_list[[prs]]$quantile

    ggsave(
        paste0("prs_onset_patterns_", tolower(prs), ".pdf"),
        combined_plot,
        width = 15,
        height = 6
    )
}
```

```{r}
# Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(igraph)
library(viridis)

# Function to extract disease sequences for each individual
get_disease_sequences <- function(data) {
    # Order by individual and age of diagnosis
    sequences <- data %>%
        arrange(eid, age_diag) %>%
        group_by(eid) %>%
        summarise(
            sequence = list(exclude_name),
            ages = list(age_diag),
            n_diseases = n()
        ) %>%
        filter(n_diseases >= 3)  # Only look at individuals with multiple diseases

    return(sequences)
}

# Function to analyze transitions
analyze_transitions <- function(sequences) {
    # Create transition matrix
    transitions <- data.frame()

    for(i in 1:nrow(sequences)) {
        seq <- unlist(sequences$sequence[i])
        if(length(seq) >= 3) {
            for(j in 1:(length(seq)-1)) {
                transitions <- bind_rows(transitions,
                    data.frame(
                        from = seq[j],
                        to = seq[j+1],
                        eid = sequences$eid[i]
                    )
                )
            }
        }
    }

    # Count transitions
    transition_counts <- transitions %>%
        group_by(from, to) %>%
        summarise(
            weight = n(),
            .groups = 'drop'
        ) %>%
        filter(weight >= 5)  # Only keep transitions that occur at least 5 times

    return(list(
        all_transitions = transitions,
        counts = transition_counts
    ))
}

# Function to analyze genetic effects on transitions
analyze_genetic_effects <- function(transitions, prs_data) {
    # Merge with PRS data
    transitions_prs <- transitions %>%
        inner_join(prs_data, by = c("eid" = "Identifier"))

    # Analyze PRS effects on specific transitions
    results <- list()

    for(prs_col in names(prs_data)[1:(ncol(prs_data)-1)]) {
        # Look at how PRS affects each transition type
        transition_effects <- transitions_prs %>%
            group_by(from, to) %>%
            summarise(
                n = n(),
                prs_effect = cor(get(prs_col), 1:n(), method="spearman"),
                .groups = 'drop'
            ) %>%
            filter(n >= 20)  # Only analyze transitions with enough data

        results[[prs_col]] <- transition_effects
    }

    return(results)
}

# Get sequences
sequences <- get_disease_sequences(Y_sum_df)

# Analyze transitions
transition_results <- analyze_transitions(sequences[1:1000,])  # Limit to first 1000 for speed

# Create network visualization
g <- graph_from_data_frame(
    transition_results$counts,
    directed = TRUE,
    vertices = unique(c(transition_results$counts$from, 
                       transition_results$counts$to))
)

# Calculate node sizes based on frequency
node_sizes <- degree(g, mode="in") + degree(g, mode="out")

# Create network plot
pdf("disease_transitions_network.pdf", width = 12, height = 12)
plot(g,
     edge.width = log1p(E(g)$weight),
     vertex.size = 20 * node_sizes/max(node_sizes),
     vertex.color = viridis(vcount(g)),
     vertex.label.cex = 0.6,
     layout = layout_with_fr,
     edge.arrow.size = 0.5,
     main = "Disease Transition Network")
dev.off()

# Analyze genetic effects
genetic_effects <- analyze_genetic_effects(
    transition_results$all_transitions,
    prs_df
)

# Create visualization of genetic effects
pdf("genetic_effects_on_transitions.pdf", width = 15, height = 10)
par(mfrow=c(2,2))
for(prs in names(genetic_effects)[1:4]) {  # Show top 4 PRS effects
    effects_df <- genetic_effects[[prs]]

    if(nrow(effects_df) > 0) {
        # Create heatmap of effects
        effect_matrix <- reshape2::acast(effects_df, from ~ to, value.var = "prs_effect")

        heatmap(effect_matrix,
                col = colorRampPalette(c("blue", "white", "red"))(100),
                main = paste("PRS Effects on Transitions -", prs),
                cexRow = 0.7,
                cexCol = 0.7)
    }
}
dev.off()

# First, get all possible disease categories
all_categories <- unique(c(
    unique(transition_results$all_transitions$from),
    unique(transition_results$all_transitions$to)
))
all_categories <- all_categories[all_categories != ""]  # Remove empty category if present

# Modified plotting function
plot_prs_heatmaps <- function(genetic_effects, top_n = 35 ) {
    par(mfrow=c(2,2))

    for(prs in names(genetic_effects)[1:top_n]) {
        effects_df <- genetic_effects[[prs]]

        if(nrow(effects_df) > 0) {
            # Create complete matrix with all possible transitions
            full_matrix <- matrix(0,  # Fill with zeros (white in heatmap)
                                nrow = length(all_categories),
                                ncol = length(all_categories),
                                dimnames = list(all_categories, all_categories))

            # Fill in observed transitions
            for(i in 1:nrow(effects_df)) {
                if(effects_df$from[i] != "" && effects_df$to[i] != "") {
                    full_matrix[effects_df$from[i], effects_df$to[i]] <- effects_df$prs_effect[i]
                }
            }

            # Create heatmap
            heatmap(full_matrix,
                   col = colorRampPalette(c("blue", "white", "red"))(100),
                   main = paste("PRS Effects on Transitions -", prs),
                   cexRow = 0.7,
                   cexCol = 0.7,
                   scale = "none")  # Don't scale the matrix
        }
    }
}

# Run the visualization
pdf("prs_transition_heatmaps.pdf", width = 12, height = 12)
plot_prs_heatmaps(genetic_effects)
dev.off()

# Create summary of most common transition patterns
common_patterns <- sequences %>%
    filter(n_diseases >= 3) %>%  # Look at sequences of 3 or more diseases
    mutate(
        pattern = sapply(sequence, paste, collapse = "->")
    ) %>%
    count(pattern, sort = TRUE) %>%
    head(20)  # Top 20 patterns

# Save results
write.csv(common_patterns, "common_disease_patterns.csv", row.names = FALSE)

# Print summary
cat("\nMost common disease transition patterns:\n")
print(common_patterns)

# Analysis of time between transitions
time_between <- sequences %>%
    rowwise() %>%
    mutate(
        time_gaps = list(diff(unlist(ages)))
    ) %>%
    unnest(time_gaps)

# Create visualization of time between transitions
p_time <- ggplot(time_between, aes(x = time_gaps)) +
    geom_histogram(bins = 30, fill = "lightblue", color = "black") +
    theme_minimal() +
    labs(
        title = "Time Between Disease Transitions",
        x = "Years between diagnoses",
        y = "Count"
    )

ggsave("transition_times.pdf", p_time, width = 8, height = 6)

```

```{r}

analyze_transitions <- function(sequences) {
    transitions <- data.frame()

    for(i in 1:nrow(sequences)) {
        seq <- unlist(sequences$sequence[i])
        if(length(seq) >= 3) {
            for(j in 1:(length(seq)-1)) {
                transitions <- bind_rows(transitions,
                    data.frame(
                        from = seq[j],
                        to = seq[j+1],
                        eid = sequences$eid[i],
                        position_in_sequence = j  # Add this to track actual position
                    )
                )
            }
        }
    }

    # Count transitions
    transition_counts <- transitions %>%
        group_by(from, to) %>%
        summarise(
            weight = n(),
            .groups = 'drop'
        ) %>%
        filter(weight >= 5)

    return(list(
        all_transitions = transitions,
        counts = transition_counts
    ))
}

# Then modify analyze_genetic_effects
analyze_genetic_effects <- function(transitions, prs_data) {
    transitions_prs <- transitions %>%
        inner_join(prs_data, by = c("eid" = "Identifier"))

    results <- list()

    for(prs_col in names(prs_data)[1:(ncol(prs_data)-1)]) {
        transition_effects <- transitions_prs %>%
            group_by(from, to) %>%
            summarise(
                n = n(),
                # Now using actual position in sequence
                prs_effect = cor(get(prs_col), position_in_sequence, method="spearman"),
                .groups = 'drop'
            ) %>%
            filter(n >= 20)

        results[[prs_col]] <- transition_effects
    }

    return(results)
}

```

```{r}

sequences <- get_disease_sequences(Y_sum_df)

# Analyze transitions
transition_results <- analyze_transitions(sequences[1:20000,])
# Limit to first 1000 for speed

genetic_effects <- analyze_genetic_effects(
    transition_results$all_transitions,
    prs_df
)

# First, get all possible disease categories
all_categories <- unique(c(
    unique(transition_results$all_transitions$from),
    unique(transition_results$all_transitions$to)
))
all_categories <- all_categories[all_categories != ""]  # Remove empty category if present

plot_prs_heatmaps <- function(genetic_effects, top_n = 36) {
    # Set up plotting parameters for a 2x2 grid
    par(mfrow=c(2,2))

    for(prs in names(genetic_effects)[1:top_n]) {
        effects_df <- genetic_effects[[prs]]

        if(nrow(effects_df) > 0) {
            # Create complete matrix with all possible transitions
            full_matrix <- matrix(0,
                                nrow = length(all_categories),
                                ncol = length(all_categories),
                                dimnames = list(all_categories, all_categories))

            # Fill in observed transitions
            for(i in 1:nrow(effects_df)) {
                if(effects_df$from[i] != "" && effects_df$to[i] != "") {
                    full_matrix[effects_df$from[i], effects_df$to[i]] <- effects_df$prs_effect[i]
                }
            }

            # Get the range of correlation values for the legend
            val_range <- range(full_matrix, na.rm = TRUE)
            breaks <- seq(min(val_range), max(val_range), length.out = 101)

            # Create color palette
            col_palette <- colorRampPalette(c("blue", "white", "red"))(100)

            # Set up the layout to include space for the legend
            layout(matrix(c(1,2), ncol=2), widths=c(4,1))

            # Create heatmap
            heatmap(full_matrix,
                   col = col_palette,
                   main = paste("PRS Effects on Transitions -", prs),
                   cexRow = 0.7,
                   cexCol = 0.7,
                   scale = "none")

            # Add color key
            par(mar=c(5,1,4,2))
            plot(c(0,2), c(min(val_range), max(val_range)), 
                 type = 'n', axes = FALSE, xlab = "", ylab = "")
            title("Correlation\nStrength", line = 1)
            for(i in 1:100) {
                y1 <- breaks[i]
                y2 <- breaks[i+1]
                col <- col_palette[i]
                rect(0, y1, 1, y2, col = col, border = NA)
            }
            axis(4, at = seq(min(val_range), max(val_range), length.out = 5),
                 labels = round(seq(min(val_range), max(val_range), length.out = 5), 2))
        }
    }
}

# Run the visualization
#pdf("prs_transition_heatmaps.pdf", width = 12, height = 12)
plot_prs_heatmaps(genetic_effects)
#dev.off()
```

```{r}

# Improved function for analyzing genetic effects with better error handling
analyze_genetic_effects <- function(transitions, prs_data) {
    # Merge with PRS data
    transitions_prs <- transitions %>%
        inner_join(prs_data, by = c("eid" = "Identifier"))

    # List to store results for each PRS
    results <- list()

    for(prs_col in names(prs_data)[1:(ncol(prs_data)-1)]) {
        # Analyze each transition type with more statistics and error handling
        transition_effects <- transitions_prs %>%
            group_by(from, to) %>%
            filter(n() >= 20) %>%  # Filter before calculating correlations
            summarise(
                n = n(),
                # Safely calculate correlation and p-value
                prs_effect = tryCatch({
                    cor(get(prs_col), 1:n(), method="spearman")
                }, error = function(e) NA),
                p_value = tryCatch({
                    cor.test(get(prs_col), 1:n(), method="spearman")$p.value
                }, error = function(e) NA),
                .groups = 'drop'
            ) %>%
            filter(!is.na(prs_effect) & !is.na(p_value)) %>%  # Remove NA results
            mutate(
                sig = p_value < 0.05,
                adj_p = p.adjust(p_value, method="BH")
            )

        results[[prs_col]] <- transition_effects
    }

    return(results)
}

```

# Function to plot the genetic effects
plot_genetic_effects <- function(genetic_effects, prs_name) {
    # Get data for specific PRS
    effects_df <- genetic_effects[[prs_name]]

    if(nrow(effects_df) > 0) {
        # Filter for significant transitions and create matrix
        sig_effects <- effects_df %>%
            filter(adj_p < 0.05) %>%  # Use adjusted p-values
            filter(from != "" & to != "") %>%  # Remove empty categories
            mutate(
                from = factor(from),
                to = factor(to)
            )

        if(nrow(sig_effects) > 0) {  # Check if we have any significant effects
            # Create main plot
            p <- ggplot(sig_effects, aes(x = from, y = to)) +
                geom_tile(aes(fill = prs_effect)) +
                geom_text(aes(label = sprintf("%.2f\n(n=%d)", prs_effect, n)), 
                         size = 2.5) +
                scale_fill_gradient2(
                    low = "blue", 
                    high = "red", 
                    mid = "white",
                    midpoint = 0,
                    name = "Correlation"
                ) +
                theme_minimal() +
                theme(
                    axis.text.x = element_text(angle = 45, hjust = 1),
                    plot.title = element_text(hjust = 0.5),
                    panel.grid.major = element_blank(),
                    panel.grid.minor = element_blank()
                ) +
                labs(
                    title = paste("PRS Effects on Disease Transitions -", prs_name),
                    x = "From Disease",
                    y = "To Disease"
                )

            return(p)
        }
    }
    return(NULL)
}

# Let's print some diagnostic information first
print_diagnostic_info <- function(transitions, prs_data) {
    cat("Transition counts by category:\n")
    print(table(transitions$from))
    print(table(transitions$to))

    cat("\nNumber of unique individuals:", length(unique(transitions$eid)))
    cat("\nNumber of total transitions:", nrow(transitions))

    # Check transition pairs with at least 20 observations
    transition_counts <- transitions %>%
        group_by(from, to) %>%
        summarise(
            n = n(),
            .groups = 'drop'
        ) %>%
        filter(n >= 20)

    cat("\n\nTransition pairs with ≥20 observations:", nrow(transition_counts))
    print(head(arrange(transition_counts, desc(n)), 10))
}

# Run the analysis with diagnostic information
print_diagnostic_info(transition_results$all_transitions, prs_df)

# Then run the genetic effects analysis
genetic_effects <- analyze_genetic_effects(
    transition_results$all_transitions,
    prs_df
)

# Create visualizations
pdf("genetic_effects_plots.pdf", width = 12, height = 8)
for(prs in names(genetic_effects)) {
    p <- plot_genetic_effects(genetic_effects, prs)
    if(!is.null(p)) {
        print(p)

        # Print summary statistics for this PRS
        effects_df <- genetic_effects[[prs]]
        cat("\nSummary for", prs, ":\n")
        cat("Total transitions analyzed:", nrow(effects_df), "\n")
        cat("Significant transitions (p < 0.05):", sum(effects_df$p_value < 0.05), "\n")
        cat("Significant transitions (adj_p < 0.05):", sum(effects_df$adj_p < 0.05), "\n")
        cat("Range of effects:", round(range(effects_df$prs_effect, na.rm=TRUE), 3), "\n\n")
    }
}
dev.off()
```

This improved version:

1. Added error handling for correlation calculations
2. Filters for minimum sample size before attempting correlations
3. Removes empty categories and NA results
4. Adds diagnostic information to help understand the data structure
5. Includes more informative console output

The main changes are:
- Pre-filtering for sufficient sample size
- Using tryCatch for correlation calculations
- Better handling of empty categories
- Added diagnostic information printing
- More robust plotting function

Try running this and let me know if you get any errors. If you do, we can look at the diagnostic information to understand what's happening in your data and adjust accordingly.
```

```{r}

library(ggplot2)
library(dplyr)
library(tidyr)

# Create toy data for two contrasting trajectories
trajectory1 <- data.frame(
  ID = "Patient 1",
  Disease = c("CVD", "Neoplasm", "Inflammatory"),
  Age = c(45, 52, 58),
  Path = "CVD → Neo → Inflammatory"
)

trajectory2 <- data.frame(
  ID = "Patient 2",
  Disease = c("Metabolic", "CVD", "Neoplasm"),
  Age = c(48, 55, 63),
  Path = "Metabolic → CVD → Neo"
)

# Combine trajectories
toy_data <- rbind(trajectory1, trajectory2)

# Set disease order
toy_data$Disease <- factor(toy_data$Disease, 
                          levels = c("Metabolic", "CVD", "Neoplasm", "Inflammatory"))

# Create the visualization
p <- ggplot(toy_data, aes(x = Age, y = Disease, color = ID)) +
  geom_point(size = 4) +
  geom_path(size = 1.2, arrow = arrow(length = unit(0.3, "cm"), type = "closed")) +
  facet_wrap(~ID, ncol = 1) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14, face = "bold"),
    legend.position = "none"
  ) +
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  labs(
    x = "Age at Diagnosis",
    y = "Disease Category",
    title = "Example Disease Trajectories",
    subtitle = "Two common patterns of disease progression"
  ) +
  scale_x_continuous(breaks = seq(40, 70, by = 5))

# Save the plot
ggsave("toy_trajectories.pdf", p, width = 10, height = 8)

# Create alternate version with multiple patients per pattern
multi_patients <- data.frame(
  ID = rep(paste("Patient", 1:6), each = 3),
  Pattern = rep(c("Pattern 1: CVD → Neo → Inflammatory",
                 "Pattern 2: Metabolic → CVD → Neo"), each = 9),
  Disease = c(rep(c("CVD", "Neoplasm", "Inflammatory"), 3),
              rep(c("Metabolic", "CVD", "Neoplasm"), 3)),
  Age = c(
    # Pattern 1 patients
    45, 52, 58,  # Patient 1
    42, 50, 55,  # Patient 2
    48, 54, 61,  # Patient 3
    # Pattern 2 patients
    48, 55, 63,  # Patient 4
    45, 53, 60,  # Patient 5
    50, 58, 65   # Patient 6
  )
)

# Set disease order
multi_patients$Disease <- factor(multi_patients$Disease, 
                               levels = c("Metabolic", "CVD", "Neoplasm", "Inflammatory"))

# Create plot with multiple patients per pattern
p2 <- ggplot(multi_patients, aes(x = Age, y = Disease, group = ID, color = ID)) +
  geom_point(size = 3) +
  geom_path(size = 1, arrow = arrow(length = unit(0.2, "cm"), type = "closed")) +
  facet_wrap(~Pattern, ncol = 1) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "right"
  ) +
  scale_color_viridis_d() +
  labs(
    x = "Age at Diagnosis",
    y = "Disease Category",
    title = "Common Disease Trajectories",
    subtitle = "Multiple examples of each progression pattern"
  ) +
  scale_x_continuous(breaks = seq(40, 70, by = 5))

# Save the second plot
ggsave("toy_trajectories_multiple.pdf", p2, width = 12, height = 10)

# Create a version showing timing distributions
set.seed(42)
n_per_pattern <- 50

# Function to generate ages with some random variation
generate_ages <- function(base_ages, n) {
  t(sapply(1:n, function(i) {
    base_ages + rnorm(length(base_ages), mean = 0, sd = 2)
  }))
}

# Generate data for many patients
pattern1_ages <- generate_ages(c(45, 52, 58), n_per_pattern)
pattern2_ages <- generate_ages(c(48, 55, 63), n_per_pattern)

distribution_data <- data.frame(
  ID = rep(paste("Patient", 1:(2*n_per_pattern)), each = 3),
  Pattern = rep(c("Pattern 1: CVD → Neo → Inflammatory",
                 "Pattern 2: Metabolic → CVD → Neo"), 
               each = n_per_pattern * 3),
  Disease = c(rep(c("CVD", "Neoplasm", "Inflammatory"), n_per_pattern),
              rep(c("Metabolic", "CVD", "Neoplasm"), n_per_pattern)),
  Age = c(pattern1_ages, pattern2_ages)
)

# Set disease order
distribution_data$Disease <- factor(distribution_data$Disease, 
                                  levels = c("Metabolic", "CVD", "Neoplasm", "Inflammatory"))

# Create violin plot showing distributions
p3 <- ggplot(distribution_data, aes(x = Disease, y = Age, fill = Disease)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  geom_boxplot(width = 0.2, alpha = 0.5) +
  facet_wrap(~Pattern, ncol = 1) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "none"
  ) +
  scale_fill_viridis_d() +
  labs(
    x = "Disease Category",
    y = "Age at Diagnosis",
    title = "Age Distribution of Disease Onset",
    subtitle = "Showing typical timing patterns for each trajectory"
  )

# Save the third plot
ggsave("toy_trajectories_distributions.pdf", p3, width = 12, height = 10)

```

```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)

# Create example patient trajectories
# Generate toy data for 3 different patient patterns
set.seed(42)

# Function to create smooth transitions between timepoints
interpolate_loadings <- function(start_loads, end_loads, n_points = 10) {
  t <- seq(0, 1, length.out = n_points)
  sapply(1:length(start_loads), function(i) {
    start_loads[i] + (end_loads[i] - start_loads[i]) * t
  })
}

# Define main disease categories
disease_cats <- c("circulatory system", "neoplasms", "digestive", 
                 "respiratory", "mental disorders", "neurological",
                 "endocrine/metabolic", "infectious diseases")

# Create three different patient trajectories
create_patient_trajectory <- function(pattern_type, n_timepoints = 50) {
  age_seq <- seq(40, 70, length.out = n_timepoints)

  if(pattern_type == 1) {
    # Pattern 1: CVD → Neoplasm progression
    start_loads <- c(0.3, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1)
    mid_loads <- c(0.4, 0.2, 0.1, 0.1, 0.05, 0.05, 0.05, 0.05)
    end_loads <- c(0.3, 0.4, 0.1, 0.05, 0.05, 0.05, 0.025, 0.025)
  } else if(pattern_type == 2) {
    # Pattern 2: Metabolic → CVD progression
    start_loads <- c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.3, 0.1)
    mid_loads <- c(0.2, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1)
    end_loads <- c(0.4, 0.1, 0.1, 0.1, 0.05, 0.05, 0.15, 0.05)
  } else {
    # Pattern 3: Inflammatory/Infectious → Multiple systems
    start_loads <- c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.3)
    mid_loads <- c(0.15, 0.15, 0.2, 0.15, 0.1, 0.1, 0.1, 0.15)
    end_loads <- c(0.2, 0.2, 0.2, 0.2, 0.05, 0.05, 0.05, 0.05)
  }

  # Create smooth transitions
  first_half <- t(interpolate_loadings(start_loads, mid_loads, n_timepoints/2))
  second_half <- t(interpolate_loadings(mid_loads, end_loads, n_timepoints/2))
  loadings <- rbind(first_half, second_half[-1,])

  # Create dataframe
  data.frame(
    Age = age_seq,
    Pattern = paste("Patient Type", pattern_type),
    t(loadings)
  ) %>%
    setNames(c("Age", "Pattern", disease_cats))
}

# Generate data for each pattern
pattern1 <- create_patient_trajectory(1)
pattern2 <- create_patient_trajectory(2)
pattern3 <- create_patient_trajectory(3)

# Combine all patterns
all_patterns <- bind_rows(pattern1, pattern2, pattern3)

# Convert to long format for plotting
plot_data <- all_patterns %>%
  pivot_longer(
    cols = -c(Age, Pattern),
    names_to = "Disease",
    values_to = "Loading"
  )

# Create stacked area plot
p1 <- ggplot(plot_data, aes(x = Age, y = Loading, fill = Disease)) +
  geom_area(position = "stack") +
  facet_wrap(~Pattern, ncol = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) +
  labs(
    title = "Disease Loading Trajectories",
    subtitle = "Showing proportional disease burden over time",
    x = "Age",
    y = "Disease Loading",
    fill = "Disease Category"
  )

# Save the plot
ggsave("disease_loading_trajectories.pdf", p1, width = 12, height = 12)

# Create alternative visualization with heatmap style
p2 <- ggplot(plot_data, aes(x = Age, y = Disease, fill = Loading)) +
  geom_tile() +
  facet_wrap(~Pattern, ncol = 1) +
  scale_fill_viridis_c() +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) +
  labs(
    title = "Disease Loading Patterns",
    subtitle = "Heatmap showing disease burden intensity over time",
    x = "Age",
    y = "Disease Category",
    fill = "Loading"
  )

# Save the heatmap version
ggsave("disease_loading_heatmap.pdf", p2, width = 12, height = 12)

# Create small multiples showing individual disease trajectories
p3 <- ggplot(plot_data, aes(x = Age, y = Loading, color = Pattern)) +
  geom_line(size = 1) +
  facet_wrap(~Disease, scales = "free_y") +
  scale_color_viridis_d() +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "Individual Disease Category Trajectories",
    x = "Age",
    y = "Loading",
    color = "Patient Type"
  )

# Save the small multiples version
ggsave("disease_loading_individual.pdf", p3, width = 15, height = 10)

````

```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)

# Define disease categories
disease_cats <- c(
  "infectious.diseases",
  "neoplasms",
  "endocrine.metabolic",
  "circulatory.system",
  "respiratory",
  "digestive",
  "mental.disorders",
  "neurological"
)

# Function to ensure loadings sum to 1
normalize_loadings <- function(x) {
  x / sum(x)
}

# Function to create smooth transitions between timepoints
create_trajectory <- function(age_range = seq(40, 70, length.out = 50), 
                            keypoints) {
  # Interpolate between keypoints
  result <- matrix(nrow = length(age_range), ncol = length(disease_cats))

  for(i in 1:(nrow(keypoints)-1)) {
    start_idx <- (i-1) * length(age_range)/(nrow(keypoints)-1) + 1
    end_idx <- i * length(age_range)/(nrow(keypoints)-1)
    indices <- round(start_idx:end_idx)

    # Create smooth transition
    for(j in 1:length(disease_cats)) {
      result[indices, j] <- approx(
        c(1, length(indices)),
        c(keypoints[i,j], keypoints[i+1,j]),
        1:length(indices)
      )$y
    }
  }

  # Normalize each row to sum to 1
  result <- t(apply(result, 1, normalize_loadings))
  colnames(result) <- disease_cats

  return(data.frame(
    Age = age_range,
    as.data.frame(result)
  ))
}

# Create three distinct patient trajectories
# Patient 1: Circulatory -> Neoplasm progression
p1_keypoints <- matrix(c(
  0.1, 0.1, 0.1, 0.3, 0.1, 0.1, 0.1, 0.1,  # Start: high circulatory
  0.1, 0.2, 0.1, 0.3, 0.1, 0.1, 0.05, 0.05,  # Mid: increasing neoplasm
  0.05, 0.4, 0.05, 0.2, 0.1, 0.1, 0.05, 0.05  # End: high neoplasm
), nrow = 3, byrow = TRUE)

# Patient 2: Metabolic -> Circulatory progression
p2_keypoints <- matrix(c(
  0.1, 0.1, 0.3, 0.1, 0.1, 0.1, 0.1, 0.1,  # Start: high metabolic
  0.1, 0.1, 0.2, 0.3, 0.1, 0.1, 0.05, 0.05,  # Mid: increasing circulatory
  0.05, 0.1, 0.15, 0.4, 0.1, 0.1, 0.05, 0.05  # End: high circulatory
), nrow = 3, byrow = TRUE)

# Patient 3: Infectious -> Multiple systems
p3_keypoints <- matrix(c(
  0.3, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1,  # Start: high infectious
  0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.05, 0.05,  # Mid: spreading
  0.1, 0.2, 0.15, 0.15, 0.15, 0.15, 0.05, 0.05  # End: multiple systems
), nrow = 3, byrow = TRUE)

# Generate trajectories
traj1 <- create_trajectory(keypoints = p1_keypoints) %>% mutate(Pattern = "Patient 1: CVD → Neoplasm")
traj2 <- create_trajectory(keypoints = p2_keypoints) %>% mutate(Pattern = "Patient 2: Metabolic → CVD")
traj3 <- create_trajectory(keypoints = p3_keypoints) %>% mutate(Pattern = "Patient 3: Infectious → Multiple")

# Combine trajectories
all_trajectories <- bind_rows(traj1, traj2, traj3)

# Convert to long format
plot_data <- all_trajectories %>%
  pivot_longer(
    cols = all_of(disease_cats),
    names_to = "Disease",
    values_to = "Loading"
  )

# Create streamgraph
p1 <- ggplot(plot_data, aes(x = Age, y = Loading, fill = Disease)) +
  geom_area(position = "fill") +
  facet_wrap(~Pattern, ncol = 1) +
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) +
  labs(
    title = "Disease Loading Trajectories",
    subtitle = "Showing relative disease burden over time",
    x = "Age",
    y = "Proportion of Disease Burden",
    fill = "Disease Category"
  ) +
  scale_y_continuous(labels = scales::percent)

# Save the main plot
ggsave("disease_proportions_stream.pdf", p1, width = 12, height = 12)

# Create heatmap version
p2 <- ggplot(plot_data, aes(x = Age, y = Disease, fill = Loading)) +
  geom_tile() +
  facet_wrap(~Pattern, ncol = 1) +
  scale_fill_viridis_c(limits = c(0, 0.4)) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12)
  ) +
  labs(
    title = "Disease Loading Patterns",
    subtitle = "Heatmap showing relative disease burden over time",
    x = "Age",
    y = "Disease Category",
    fill = "Proportion"
  )

# Save the heatmap
ggsave("disease_proportions_heatmap.pdf", p2, width = 12, height = 12)

# Create individual trajectories plot
p3 <- ggplot(plot_data, aes(x = Age, y = Loading, color = Pattern)) +
  geom_line(size = 1) +
  facet_wrap(~Disease) +
  scale_color_viridis_d() +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 8)
  ) +
  labs(
    title = "Disease Category Trajectories",
    x = "Age",
    y = "Proportion of Disease Burden",
    color = "Trajectory Pattern"
  ) +
  scale_y_continuous(limits = c(0, 0.4), labels = scales::percent)

# Save the individual trajectories
ggsave("disease_proportions_individual.pdf", p3, width = 15, height = 10)
```

```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(RColorBrewer)

# Use same data generation code as before, but change visualization
# [Previous data generation code remains the same until plotting]

# Create heatmap with brighter colors
p2 <- ggplot(plot_data, aes(x = Age, y = Disease, fill = Loading)) +
  geom_tile() +
  facet_wrap(~Pattern, ncol = 1) +
  # Use a brighter color palette
  scale_fill_gradientn(
    colors = c(
      "#FFFFFF",  # white for zero
      "#FFF7BC",  # very light yellow
      "#FFA07A",  # light salmon
      "#FF4500",  # orange red
      "#FF0000"   # bright red
    ),
    limits = c(0, 0.4),
    breaks = seq(0, 0.4, by = 0.1),
    labels = scales::percent
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    panel.background = element_rect(fill = "white", color = NA),
    strip.background = element_rect(fill = "gray95", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  ) +
  labs(
    title = "Disease Loading Patterns",
    subtitle = "Heatmap showing relative disease burden over time",
    x = "Age",
    y = "Disease Category",
    fill = "Disease\nBurden"
  )

# Try alternative color scheme
p3 <- ggplot(plot_data, aes(x = Age, y = Disease, fill = Loading)) +
  geom_tile() +
  facet_wrap(~Pattern, ncol = 1) +
  # Use a more vibrant diverging palette
  scale_fill_gradientn(
    colors = c(
      "#FFFFFF",      # white
      "#FFE29D",      # light yellow
      "#FFA07A",      # light salmon
      "#FF4E50",      # coral red
      "#E31A1C"       # bright red
    ),
    limits = c(0, 0.4),
    breaks = seq(0, 0.4, by = 0.1),
    labels = scales::percent
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    panel.background = element_rect(fill = "white", color = NA),
    strip.background = element_rect(fill = "gray95", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  ) +
  labs(
    title = "Disease Loading Patterns",
    subtitle = "Heatmap showing relative disease burden over time",
    x = "Age",
    y = "Disease Category",
    fill = "Disease\nBurden"
  )

# Save both versions
ggsave("disease_heatmap_bright1.pdf", p2, width = 12, height = 12)
ggsave("disease_heatmap_bright2.pdf", p3, width = 12, height = 12)

# Create a third version with an even more vibrant scheme
p4 <- ggplot(plot_data, aes(x = Age, y = Disease, fill = Loading)) +
  geom_tile() +
  facet_wrap(~Pattern, ncol = 1) +
  # Use a highly saturated diverging palette
  scale_fill_gradientn(
    colors = c(
      "#FFFFFF",      # white
      "#FFF200",      # bright yellow
      "#FF9933",      # bright orange
      "#FF3333",      # bright red
      "#990000"       # dark red
    ),
    limits = c(0, 0.4),
    breaks = seq(0, 0.4, by = 0.1),
    labels = scales::percent
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid = element_blank(),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    panel.background = element_rect(fill = "white", color = NA),
    strip.background = element_rect(fill = "gray95", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  ) +
  labs(
    title = "Disease Loading Patterns",
    subtitle = "Heatmap showing relative disease burden over time",
    x = "Age",
    y = "Disease Category",
    fill = "Disease\nBurden"
  )

ggsave("disease_heatmap_bright3.pdf", p4, width = 12, height = 12)
```

```{r}
# Get the column names from d
d_cols <- colnames(d)

# Get the mapping from ATM package
phecode_map <- ATM::disease_info_phecode_icd10[, c("phecode", "phenotype")]

# Convert phecodes to match format in d_cols
phecode_map$phecode <- as.character(phecode_map$phecode)

# Create a mapping function to handle different formats
format_phecode <- function(x) {
    # Convert decimal points to match format
    if(grepl("\\.", x)) {
        # If already has decimal, just convert to character
        return(as.character(x))
    } else {
        # If no decimal, add .0
        return(paste0(x, ".0"))
    }
}

# Apply formatting to both sets
d_cols_formatted <- sapply(d_cols, format_phecode)
phecode_map$phecode_formatted <- sapply(phecode_map$phecode, format_phecode)

# Match and create named vector
disease_names <- phecode_map$phenotype[match(d_cols_formatted, phecode_map$phecode_formatted)]

# Check first few matches
head(data.frame(
    code = d_cols,
    phenotype = disease_names
))

# Save for later use
saveRDS(disease_names, "~/Dropbox (Personal)/disease_names.rds")

```

```{r}
library(igraph)
library(ggplot2)
library(dplyr)

# Function to create transition network and PRS association plot
analyze_disease_patterns <- function(sequences, prs_data, min_transitions = 20) {
    # Create transition network
    transitions <- data.frame()
    for(i in 1:nrow(sequences)) {
        seq <- unlist(sequences$sequence[i])
        if(length(seq) >= 2) {  # Changed to 2 to see more transitions
            for(j in 1:(length(seq)-1)) {
                transitions <- bind_rows(transitions,
                    data.frame(
                        from = seq[j],
                        to = seq[j+1],
                        eid = sequences$eid[i]
                    )
                )
            }
        }
    }

    # Count transitions and filter for common ones
    transition_counts <- transitions %>%
        group_by(from, to) %>%
        summarise(
            weight = n(),
            .groups = 'drop'
        ) %>%
        filter(weight >= min_transitions)

    # Create network
    g <- graph_from_data_frame(
        transition_counts,
        directed = TRUE
    )

    # Calculate node statistics
    V(g)$size <- degree(g, mode="total") * 2
    V(g)$color <- ifelse(V(g)$name %in% c("circulatory system", "cardiovascular", "heart disease"), 
                        "red", "lightblue")
    E(g)$width <- log(E(g)$weight)

    # Plot network
    pdf("disease_transitions.pdf", width = 12, height = 8)
    plot(g,
         layout = layout_with_fr,
         vertex.label.cex = 0.8,
         vertex.label.color = "black",
         edge.arrow.size = 0.5,
         main = "Disease Transition Network")
    dev.off()

    # Analyze PRS associations with disease categories
    # Merge sequence data with PRS
    disease_prs <- data.frame()
    for(i in 1:nrow(sequences)) {
        seq <- unlist(sequences$sequence[i])
        disease_prs <- bind_rows(disease_prs,
            data.frame(
                eid = sequences$eid[i],
                disease = seq
            )
        )
    }

    disease_prs <- disease_prs %>%
        inner_join(prs_data, by = c("eid" = "Identifier"))

    # Calculate average PRS scores for each disease category
    prs_summary <- disease_prs %>%
        group_by(disease) %>%
        summarise(
            n = n(),
            mean_cad_prs = mean(CAD),
            sd_cad_prs = sd(CAD),
            .groups = 'drop'
        ) %>%
        filter(n >= min_transitions) %>%
        arrange(desc(mean_cad_prs))

    # Create PRS association plot
    p <- ggplot(prs_summary, aes(x = reorder(disease, mean_cad_prs), y = mean_cad_prs)) +
        geom_bar(stat = "identity", fill = "lightblue") +
        geom_errorbar(aes(ymin = mean_cad_prs - sd_cad_prs, 
                         ymax = mean_cad_prs + sd_cad_prs), 
                     width = 0.2) +
        coord_flip() +
        theme_minimal() +
        labs(
            title = "Average CAD PRS by Disease Category",
            x = "Disease Category",
            y = "Mean CAD PRS Score"
        ) +
        theme(
            axis.text.y = element_text(size = 8)
        )

    ggsave("prs_by_disease.pdf", p, width = 10, height = 8)

    return(list(
        network = g,
        prs_summary = prs_summary
    ))
}

# Run the analysis
results <- analyze_disease_patterns(sequences, prs_df)

```

```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Function to plot age of onset distributions with overlapping densities
plot_age_distributions <- function(data) {
  cv_conditions <- c(
    "Coronary atherosclerosis",
    "Myocardial infarction",
    "Angina pectoris",
    "Essential hypertension",
    "Other chronic ischemic heart disease,unspecified"
  )

  plot_data <- data %>%
    filter(phenotype %in% cv_conditions)

  # Get counts for legend
  counts <- plot_data %>%
    group_by(phenotype) %>%
    summarise(n = n()) %>%
    mutate(legend_label = paste0(phenotype, " (n=", format(n, big.mark=","), ")"))

  # Create mapping of old to new labels
  label_map <- setNames(counts$legend_label, counts$phenotype)

  # Create the plot
  ggplot(plot_data, aes(x = age_diag, y = phenotype, fill = phenotype)) +
    geom_density_ridges(alpha = 0.2) +
    scale_x_continuous(limits = c(40, 80), breaks = seq(40, 80, by = 10)) +
    scale_color_brewer(palette = "Set2", labels = label_map) +
    scale_fill_brewer(palette = "Set2", labels = label_map) +
    labs(
      title = "Age of Onset Distribution for Cardiovascular Conditions",
      x = "Age of Onset",
      y = "Density",
      color = "Condition",
      fill = "Condition"
    ) +
    theme_minimal() +
    theme(
      legend.position = "right",
      plot.title = element_text(hjust = 0.5, size = 14),
      axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      legend.text = element_text(size = 10)
    )
}

# Function to create summary statistics table
summarize_ages <- function(data) {
  data %>%
    group_by(phenotype) %>%
    summarise(
      n = n(),
      mean_age = mean(age_diag, na.rm = TRUE),
      median_age = median(age_diag, na.rm = TRUE),
      q25 = quantile(age_diag, 0.25, na.rm = TRUE),
      q75 = quantile(age_diag, 0.75, na.rm = TRUE)
    ) %>%
    mutate(
      IQR = paste(round(q25, 1), "-", round(q75, 1)),
      mean_age = round(mean_age, 1),
      median_age = round(median_age, 1)
    ) %>%
    select(phenotype, n, mean_age, median_age, IQR) %>%
    arrange(mean_age)  # Sort by mean age
}

# Example usage:
# data <- read.csv("your_data.csv")
# 
# # Create and save the plot
 p <- plot_age_distributions(data)
# ggsave("age_distributions.pdf", p, width = 10, height = 6)
# 
# # Get summary statistics
# summary_stats <- summarize_ages(data)
# print(summary_stats)

# Alternative version with individual facets
plot_age_distributions_faceted <- function(data) {
  cv_conditions <- c(
    "Coronary atherosclerosis",
    "Myocardial infarction",
    "Angina pectoris",
    "Essential hypertension",
    "Other chronic ischemic heart disease,unspecified"
  )

  plot_data <- data %>%
    filter(phenotype %in% cv_conditions)

  # Get counts for annotations
  counts <- plot_data %>%
    group_by(phenotype) %>%
    summarise(
      n = n(),
      median = median(age_diag),
      mean = mean(age_diag)
    ) %>%
    mutate(
      count_label = paste0("n=", format(n, big.mark=",")),
      stat_label = paste0("Mean: ", round(mean,1), "\nMedian: ", round(median,1))
    )

  ggplot(plot_data, aes(x = age_diag)) +
    geom_density_ridges(fill = "blue", alpha = 0.5) +
    facet_wrap(~phenotype, ncol = 1, scales = "free_y") +
    geom_text(
      data = counts,
      aes(x = 75, y = Inf, label = count_label),
      hjust = 1, vjust = 2
    ) +
    geom_text(
      data = counts,
      aes(x = 75, y = Inf, label = stat_label),
      hjust = 1, vjust = 4
    ) +
    scale_x_continuous(limits = c(40, 80), breaks = seq(40, 80, by = 10)) +
    labs(
      title = "Age of Onset Distribution for Cardiovascular Conditions",
      x = "Age of Onset",
      y = "Density"
    ) +
    theme_minimal() +
    theme(
      strip.text = element_text(size = 12, face = "bold"),
      plot.title = element_text(hjust = 0.5, size = 14),
      axis.text = element_text(size = 12),
      axis.title = element_text(size = 12)
    )
}
```

```{r}

library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)

create_signature_plots <- function() {
  # Define cardiovascular diseases with exact colors
  diseases_cv <- tibble(
    Disease = c(
      "Myocardial infarction",
      "Essential hypertension",
      "Congestive heart failure",
      "Atrial fibrillation",
      "Angina pectoris"
    ),
    Peak_Age = c(55, 45, 60, 65, 50),
    Color = c(
      "#2196F3",  # Blue for MI
      "#4CAF50",  # Green for HTN
      "#F44336",  # Red for CHF
      "#9C27B0",  # Purple for AF
      "#FF9800"   # Orange for Angina
    )
  )

  # Define neoplastic diseases with exact colors
  diseases_neo <- tibble(
    Disease = c(
      "Secondary malignancy",
      "Neutropenia",
      "Malignant neoplasm of breast",
      "Cancer of bronchus/lung"
    ),
    Peak_Age = c(60, 55, 58, 65),
    Color = c(
      "#E91E63",  # Pink for Secondary
      "#795548",  # Brown for Neutropenia
      "#F44336",  # Red for Breast
      "#9E9E9E"   # Grey for Lung
    )
  )
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)

create_signature_plots <- function() {
  # Define cardiovascular diseases with exact colors
  diseases_cv <- tibble(
    Disease = c(
      "Myocardial infarction",
      "Essential hypertension",
      "Congestive heart failure",
      "Atrial fibrillation",
      "Angina pectoris"
    ),
    Peak_Age = c(55, 45, 60, 65, 50),
    Color = c(
      "#2196F3",  # Blue for MI
      "#4CAF50",  # Green for HTN
      "#F44336",  # Red for CHF
      "#9C27B0",  # Purple for AF
      "#FF9800"   # Orange for Angina
    )
  )

  # Define neoplastic diseases with exact colors
  diseases_neo <- tibble(
    Disease = c(
      "Secondary malignancy",
      "Neutropenia",
      "Malignant neoplasm of breast",
      "Cancer of bronchus/lung"
    ),
    Peak_Age = c(60, 55, 58, 65),
    Color = c(
      "#E91E63",  # Pink for Secondary
      "#795548",  # Brown for Neutropenia
      "#F44336",  # Red for Breast
      "#9E9E9E"   # Grey for Lung
    )
  )

  # Function to generate probability progression
  generate_probs <- function(peak_age, max_prob = 0.95, steepness = 0.2) {
    age_points <- seq(35, 75, by = 10)  # Adjust for 5 time points
    age_points %>%
      map_dbl(~{
        base_prob <- max_prob * exp(-steepness * (. - peak_age)^2 / 100)
        if (. >= peak_age - 10) max(0, base_prob) else max(0, base_prob)
      })
  }

  # Function to create plot data
  create_plot_data <- function(diseases) {
    age_points <- seq(35, 75, by = 10)  # 5 time points

    diseases %>%
      group_by(Disease, Color, Peak_Age) %>%
      summarise(
        TimePoint = list(age_points),
        Probabilities = list(generate_probs(Peak_Age))
      ) %>%
      unnest(cols = c(TimePoint, Probabilities)) %>%
      group_by(Disease) %>%
      mutate(TimeIndex = row_number())
  }

  # Generate plot data
  cv_data <- create_plot_data(diseases_cv)
  neo_data <- create_plot_data(diseases_neo)

  # Function to create signature plot
  create_signature_plot <- function(data, title) {
    ggplot(data, aes(y = reorder(Disease, desc(Peak_Age)))) +
      # Add horizontal lines
      geom_segment(aes(x = min(TimePoint), xend = max(TimePoint), 
                      y = Disease, yend = Disease),
                  color = "gray80", size = 0.3) +
      # Add filled circles with outline
      geom_point(aes(x = TimePoint, fill = Disease,
                     size = Probabilities, alpha = Probabilities),
                 shape = 21, stroke = 0.5, color = Color) +  # white outline
      geom_point(aes(x = TimePoint, color = Disease,
                     size = Probabilities),
                 alpha = 1) +  # filled circles
      # Add probability labels
      geom_text(aes(x = TimePoint, label = sprintf("φ%d%d", TimeIndex, 
                    as.numeric(factor(Disease)))),
                size = 2.5, vjust = 2.2, color = "black") +
      scale_size_continuous(range = c(8, 15),  # Larger circles
                           breaks = c(0.00, 0.25, 0.50, 0.75, 0.95)) +
      scale_alpha_continuous(range = c(0.3, 1),
                           breaks = c(0.00, 0.25, 0.50, 0.75, 0.95)) +
      scale_color_manual(values = data$Color) +
      scale_fill_manual(values = data$Color) +
      scale_x_continuous(breaks = seq(35, 75, by = 10),
                        limits = c(35, 75)) +
      labs(title = title,
           x = "Age",
           y = "",
           size = "Probability",
           alpha = "Probability",
           color = "Disease",
           fill = "Disease") +
      theme_minimal() +
      theme(
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "bottom",
        legend.box = "vertical",
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.margin = margin(t = 5, r = 10, b = 5, l = 5),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8)
      )
  }

  # Create probability scale legend
  prob_scale <- tibble(
    label = c("0.00", "0.25", "0.50", "0.75", "0.95"),
    size = c(0.00, 0.25, 0.50, 0.75, 0.95)
  )

  prob_legend <- ggplot(prob_scale, aes(x = 1:5, y = 1)) +
    geom_point(aes(size = size),
               color = "#666666",
               fill = "#666666",
               alpha = seq(0.3, 1, length.out = 5)) +
    scale_size_continuous(range = c(8, 15)) +
    scale_x_continuous(breaks = 1:5, labels = prob_scale$label) +
    labs(x = "Probability Scale") +
    theme_void() +
    theme(
      axis.text.x = element_text(),
      legend.position = "none"
    )

  # Create both plots
  cv_plot <- create_signature_plot(cv_data, "Cardiovascular Disease Progression")
  neo_plot <- create_signature_plot(neo_data, "Neoplastic Disease Progression")

  return(list(
    cv_plot = cv_plot, 
    neo_plot = neo_plot, 
    prob_legend = prob_legend
  ))
}

# Example usage:
# plots <- create_signature_plots()
# ggsave("cardiovascular_progression.pdf", plots$cv_plot, width = 10, height = 5)
# ggsave("neoplastic_progression.pdf", plots$neo_plot, width = 10, height = 4)
# ggsave("probability_legend.pdf", plots$prob_legend, width = 6, height = 1)
plots <- create_signature_plots()
```

```{r}

  # Define cardiovascular diseases with exact colors
  diseases_cv <- tibble(
    Disease = c(
      "Myocardial infarction",
      "Essential hypertension",
      "Atrial fibrillation",
      "Angina pectoris"
    ),
    Peak_Age = c(55, 45, 60, 65, 50),
    Color = c(
      "#2196F3",  # Blue for MI
      "#4CAF50",  # Green for HTN
      "#F44336",  # Red for CHF
      "#9C27B0",  # Purple for AF
      "#FF9800"   # Orange for Angina
    )
  )

  # Define neoplastic diseases with exact colors
  diseases_neo <- tibble(
    Disease = c(
      "Secondary malignancy",
      "Neutropenia",
      "Malignant neoplasm of breast",
      "Cancer of bronchus/lung"
    ),
    Peak_Age = c(60, 55, 58, 65),
    Color = c(
      "#E91E63",  # Pink for Secondary
      "#795548",  # Brown for Neutropenia
      "#F44336",  # Red for Breast
      "#9E9E9E"   # Grey for Lung
    )
  )

  # Function to generate probability progression
  generate_probs <- function(peak_age, max_prob = 0.95, steepness = 0.2) {
    age_points <- seq(35, 75, by = 10)  # Adjust for 5 time points
    age_points %>%
      map_dbl(~{
        base_prob <- max_prob * exp(-steepness * (. - peak_age)^2 / 100)
        if (. >= peak_age - 10) max(0, base_prob) else max(0, base_prob)
      })
  }

  # Function to create plot data
  create_plot_data <- function(diseases) {
    age_points <- seq(35, 75, by = 10)  # 5 time points

    diseases %>%
      group_by(Disease, Color, Peak_Age) %>%
      summarise(
        TimePoint = list(age_points),
        Probabilities = list(generate_probs(Peak_Age))
      ) %>%
      unnest(cols = c(TimePoint, Probabilities)) %>%
      group_by(Disease) %>%
      mutate(TimeIndex = row_number())
  }

  # Generate plot data
  cv_data <- create_plot_data(diseases_cv)
  neo_data <- create_plot_data(diseases_neoplastic)

                               create_signature_plot <- function(data, title) {
                                 ggplot(data, aes(y = reorder(Disease, desc(Peak_Age)))) +
                                   # Add horizontal lines
                                   geom_segment(
                                     aes(
                                       x = min(TimePoint),
                                       xend = max(TimePoint),
                                       y = Disease,
                                       yend = Disease
                                     ),
                                     color = "gray80",
                                     size = 0.3
                                   ) +
                                   # Single geom_point layer with both fill and color
                                   geom_point(
                                     aes(x = TimePoint, size = Probabilities),
                                     fill = data$Color,
                                     # Direct color assignment
                                     color = "white",
                                     # White outline
                                     alpha = data$Probabilities,
                                     # Consistent alpha
                                     shape = 21,
                                     # Filled circle with outline
                                     stroke = 0.5
                                   ) +       # Outline thickness
                                   # Add probability labels
                                   geom_text(
                                     aes(
                                       x = TimePoint,
                                       label = sprintf("φ%d%d", TimeIndex, as.numeric(factor(Disease)))
                                     ),
                                     size = 2.5,
                                     vjust = 2.2,
                                     color = "black"
                                   ) +
                                   scale_size_continuous(range = c(8, 15),
                                                         breaks = c(0.00, 0.25, 0.50, 0.75, 0.95)) +
                                   scale_x_continuous(breaks = seq(35, 75, by = 10), limits = c(35, 75)) +
                                   labs(
                                     title = title,
                                     x = "Age",
                                     y = "",
                                     size = "Probability"
                                   ) +
                                   theme_classic(base_size = 20) +
                                   theme(
                                     panel.grid.minor = element_blank(),
                                     panel.grid.major.y = element_blank(),
                                     legend.position = "bottom",
                                     legend.box = "vertical",
                                     plot.title = element_text(
                                       hjust = 0.5,
                                       size = 14,
                                       face = "bold"
                                     ),
                                     plot.margin = margin(
                                       t = 5,
                                       r = 10,
                                       b = 5,
                                       l = 5
                                     ),
                                     legend.title = element_text(size = 10),
                                     legend.text = element_text(size = 8)
                                   )
                               }

```

```{r}

  # Function to find matches for one disease
find_cv_matches <- function(theta, phi, Y, disease_names, 
                           diseases = c("Coronary atherosclerosis", "Myocardial infarction"),
                           n_examples = 5) {

  # Convert phi to probabilities
  phi_prob <- plogis(phi)

  # Get dimensions
  N <- dim(theta)[1]
  K <- dim(theta)[2]
  T <- dim(theta)[3]

  # Find disease indices
  disease_idx <- match(diseases, disease_names)

  # Function to find matches for one disease
  find_disease_matches <- function(d_idx) {
    disease_name <- disease_names[d_idx]

    # Find individuals who get this disease
    disease_cases <- which(rowSums(Y[,d_idx,]) > 0)

    if(length(disease_cases) == 0) return(NULL)

    # For each case, compute correlation
    matches <- lapply(disease_cases, function(i) {
      # Get time of disease onset - fixed this line
      onset_time <- which(Y[i,d_idx,] > 0)[1]  # Added > 0 comparison

      # Compute correlations for each signature
      sig_corrs <- sapply(1:K, function(k) {
        # Get data up to onset
        indiv_theta <- theta[i,k,1:onset_time]
        disease_phi <- phi_prob[k,d_idx,1:onset_time]
        cor(indiv_theta, disease_phi)
      })

      list(
        individual = i,
        disease = d_idx,
        disease_name = disease_name,
        onset_time = onset_time,
        correlations = sig_corrs,
        max_corr = max(sig_corrs),
        best_sig = which.max(sig_corrs)
      )
    })

    # Sort by correlation and take top examples
    matches <- matches[order(sapply(matches, function(x) x$max_corr), decreasing=TRUE)]
    matches[1:min(n_examples, length(matches))]
  }

  # Find matches for specified diseases
  all_matches <- do.call(c, lapply(disease_idx, find_disease_matches))

  # Plot function for one match
  plot_match <- function(match) {
    i <- match$individual
    d <- match$disease
    t <- match$onset_time
    k <- match$best_sig

       # Create data frames for plotting individual signature loadings
    theta_df <- melt(theta[i,,], varnames=c("signature", "time")) 
    theta_df$t=t

                   individual_prob = matrix(0, nrow = K, ncol = T)
                   for (k in 1:K) {
                     for (t in 1:T) {
                       individual_prob[k, t] <- theta[i, k, t] * phi_prob[k, d, t]
                     }
                   }

    # Best signature comparison
    best_df <- data.frame(
      time = 1:T,
      Individual_Loading = individual_prob[k,],
      Disease_Probability = phi_prob[k,d,]  # Changed to actual individual probability
    ) %>%
      pivot_longer(-time, names_to="type", values_to="value")    
    # Color palette
      # Create an extended color palette for K > 8
  create_palette <- function(n) {
    if(n <= 8) {
      return(RColorBrewer::brewer.pal(n, "Set2"))
    } else {
      # Use a combination of Set2 and additional colors
      base_colors <- RColorBrewer::brewer.pal(8, "Set2")
      # Add more distinct colors if needed
      extra_colors <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
                       "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf")
      return(c(base_colors, extra_colors)[1:n])
    }
  }

  # Use the new palette
  sig_colors <- create_palette(K)
  phi_df$t=t
  best_df$t=t

    # Create plots
    p1 <- ggplot(theta_df, aes(x=time, y=value, color=as.factor(signature))) +
      geom_line(aes(alpha=ifelse(signature==paste("Signature", k), 1, 0.3)), size=1) +
      geom_vline(xintercept=t, linetype="dashed", color="red") +
      scale_color_manual(values=sig_colors) +
      scale_alpha_identity() +
      labs(title=paste("Individual", i, "Signature Loadings"),
           subtitle=paste("Disease:", match$disease_name),
           y="Loading") +
      theme_minimal() +
      theme(legend.position="bottom")

    p2 <- ggplot(phi_df, aes(x=time, y=value, color=as.factor(signature))) +
      geom_line(aes(alpha=ifelse(signature==paste("Signature", k), 1, 0.3)), size=1) +
      geom_vline(xintercept=t, linetype="dashed", color="red") +
      scale_color_manual(values=sig_colors) +
      scale_alpha_identity() +
      labs(title="Disease Probabilities in Signatures",
           subtitle=paste("Correlation:", round(match$max_corr, 3)),
           y="Probability") +
      theme_minimal() +
      theme(legend.position="bottom")

    p3 <- ggplot(best_df, aes(x=time, y=value, color=type)) +
      geom_line(size=1) +
      geom_vline(xintercept=t, linetype="dashed", color="red") +
      scale_color_manual(values=c("Individual_Loading"=sig_colors[k], 
                                 "Disease_Probability"="darkred")) +
      labs(title=paste("Best Match: Signature", k),
           y="Value") +
      theme_minimal() +
      theme(legend.position="bottom")

    # Combine plots
    gridExtra::grid.arrange(p1, p2, p3, ncol=3,
                           top=grid::textGrob(paste("Disease Trajectory Example:", match$disease_name),
                                            gp=grid::gpar(fontsize=14, font=2)))

    # Print match details
    cat("\nMatch Details:\n")
    cat("Individual:", i, "\n")
    cat("Disease:", match$disease_name, "\n")
    cat("Onset Time:", t, "\n")
    cat("Best Signature:", k, "\n")
    cat("Correlation:", round(match$max_corr, 3), "\n")
    cat(paste(rep("=", 50), collapse=""), "\n")
  }

  # Plot all matches
  invisible(lapply(all_matches, plot_match))

  return(all_matches)
}

# Usage:
cv_matches <- find_cv_matches(theta, phi, Y2, disease_names)
```

```{r}
find_mi_matches <- function(theta, phi, Y, disease_names, n_examples = 5) {

  # Convert phi to probabilities
  phi_prob <- plogis(phi)

  # Get dimensions
  N <- dim(theta)[1]
  K <- dim(theta)[2]
  T <- dim(theta)[3]
# Find MI index
d_idx <- match("Myocardial infarction", disease_names)

# Find individuals who get MI before age 55 (index 25)
early_mi_cases <- which(apply(Y[,d_idx,1:25], 1, function(x) any(x > 0)))

if(length(early_mi_cases) == 0) {
  cat("No early MI cases found\n")
  return(NULL)
}

# For each case, check if max phi signature matches max theta signature
matches <- sapply(early_mi_cases, function(i) {
  # Get time of disease onset (only looking up to age 55)
  onset_time <- which(Y[i,d_idx,1:25] > 0)[1]

  # Find signature with max loading for this individual at onset
  theta_values <- theta[i,,onset_time]
  max_theta_sig <- which.max(theta_values)

  # Find signature with max phi for MI at onset
  phi_values <- phi_prob[,d_idx,onset_time]
  max_phi_sig <- which.max(phi_values)

  # Check if they match
  max_theta_sig == max_phi_sig
})

# Get the matching cases
matching_cases <- early_mi_cases[matches]

# Print summary
cat("Found", length(matching_cases), "early MI cases with matching signatures\n")

  # Plot function for one match
  plot_match <- function(match) {
    i <- match
    d <- disease_idx
    onset_t_name <- as.numeric(as.character(names(which(Y[match,d,]==1))))
    onset_t=which(Y[match,d,]==1)
    k <- which.max(theta[i,,onset_t])

    # Create data frames for plotting individual signature loadings
    theta_df <- melt(theta[i,,], varnames=c("signature", "time"))

    # Create phi_df
    phi_df <- melt(phi_prob[,d,], varnames=c("signature", "time"))

    # Calculate individual probabilities
    individual_prob = matrix(0, nrow = K, ncol = T)
    for (sig in 1:K) {
      for (t in 1:T) {
        individual_prob[sig, t] <- theta[i, sig, t] * phi_prob[sig, d, t]
      }
    }

    # Create color palette
    create_palette <- function(n) {
      if(n <= 8) {
        return(RColorBrewer::brewer.pal(n, "Set2"))
      } else {
        base_colors <- RColorBrewer::brewer.pal(8, "Set2")
        extra_colors <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
                         "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf")
        return(c(base_colors, extra_colors)[1:n])
      }
    }
    sig_colors <- create_palette(K)

    # Create plots

    theta_df$time=theta_df$time+30
    phi_df$time=phi_df$time+30

    p1 <- ggplot(theta_df, aes(x=time, y=value, color=as.factor(signature))) +
      geom_line() +
      geom_vline(xintercept=onset_t_name, linetype="dashed", color="red") +
      scale_color_manual(values=sig_colors, name="Signature") +
      #scale_alpha_identity() +
      labs(title=paste("Individual", i, "Signature Weights"),
           y="Weight",
           x="Age") +
      theme_minimal() +
      theme(legend.position="bottom")

    p2 <- ggplot(phi_df, aes(x=time, y=value, color=as.factor(signature))) +
      geom_line() +
      geom_vline(xintercept=onset_t_name, linetype="dashed", color="red") +
      scale_color_manual(values=sig_colors, name="Signature") +
      #scale_alpha_identity() +
      labs(title="MI Probabilities by Signature",
           y="Probability",
           x="Age") +
      theme_minimal() +
      theme(legend.position="bottom")

       # Combine plots
    gridExtra::grid.arrange(p1, p2,  ncol=2)

      }

  # Plot all matches

  invisible(lapply(matches, plot_match))

  return(matches)
}

# Usage:
mi_matches <- find_mi_matches(theta, phi, Y, disease_names)

```

```{r}
at=function(individual_id, Y, G, theta, disease_names_list, prs_names_list, age_range=30:80) {
  # Get first occurrence of each disease for ordering
  disease_timeline <- data.frame(
    disease = disease_names_list,
    first_occurrence = apply(Y[individual_id,,], 1, function(x) min(which(x == 1)))
  ) %>%
    mutate(
      first_occurrence = ifelse(is.infinite(first_occurrence), Inf, first_occurrence + 30),
      ever_occurs = first_occurrence < Inf
    )

  # Get diseases that occur for this person
  person_disease <- disease_timeline %>%
    filter(ever_occurs) %>%
    arrange(first_occurrence) %>%
    pull(disease)

  disease_indices <- match(person_disease, disease_names_list)

  # 1. Topic trajectories
  topics_df <- melt(theta_pred[individual_id,,], varnames=c("Topic", "Age")) 
  topics_df$Age =topics_df$Age+30

  p1 <- ggplot(topics_df, aes(x = Age, y = value, color = as.factor(Topic))) +
    geom_line(size = 1) +
    labs(title = "Signature Trajectories",
         y = "Topic Weight") +
    theme_minimal() +
    theme(legend.position = "right") +
    scale_color_d3()

  # 2. Disease events - now ordered by first occurrence
  diseases_df <- as.data.frame(t(Y[individual_id, disease_indices, ])) %>%
    setNames(person_disease) %>%
    mutate(Age = age_range) %>%
    pivot_longer(-Age, 
                names_to = "Disease", 
                values_to = "Status") %>%
    mutate(Disease = factor(Disease, levels = rev(person_disease)))  # Reverse order for bottom-up timeline

  p2 <- ggplot(diseases_df, aes(x = Age, y = Disease, fill = as.factor(Status))) +
    geom_tile() +
    labs(title = "Disease Events",
         y = "Disease Status",fill="Status") +
    theme_minimal() +scale_fill_manual(values = c("0" = "white", "1" = "red"))+
    theme(legend.position = "right")

  # 3. Improved genetic risk scores plot
  genetic_df <- data.frame(t(
    G[individual_id,]
  ));
  genetic_df$PRS=rownames(genetic_df)
  colnames(genetic_df)=c("Value","PRS")
  genetic_df=genetic_df%>%mutate(
      Category = case_when(
        abs(Value) >= 2 ~ "Very High",
        abs(Value) >= 1 ~ "High",
        abs(Value) >= 0.5 ~ "Moderate",
        TRUE ~ "Low"
      ),
      Direction = ifelse(Value >= 0, "Positive", "Negative"))

  p3 <- ggplot(genetic_df, aes(x = reorder(PRS, abs(Value)), y = Value)) +
    geom_col(aes(fill = Direction)) +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    geom_hline(yintercept = c(-2, -1, 1, 2), linetype = "dotted", alpha = 0.3) +
    scale_fill_manual(values = c("Negative" = "darkred", "Positive" = "steelblue")) +
    labs(title = "Genetic Risk Scores",
         subtitle = "Dotted lines at ±1 and ±2 SD",
         y = "PRS Value (SD units)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "right") +
    coord_flip()

grid.arrange(p1, p2, p3, ncol = 3)
  # Combine plots with adjusted heights
    #combined_plot <- p1 / p2 / p3 +
    #plot_layout(heights = c(2, 2, 2))

}
```

```{r}
plot_match=
function(match) {
    i <- match
    d <- disease_idx
    onset_t_name <- as.numeric(as.character(names(which(Y[match,d,]==1))))
    onset_t=which(Y[match,d,]==1)
    k <- which.max(theta[i,,onset_t])

    # Create data frames for plotting individual signature loadings
    theta_df <- melt(theta[i,,], varnames=c("signature", "time"))

    # Create phi_df
    phi_df <- melt(phi_prob[,d,], varnames=c("signature", "time"))

    # Calculate individual probabilities
    individual_prob = matrix(0, nrow = K, ncol = T)
    for (sig in 1:K) {
      for (t in 1:T) {
        individual_prob[sig, t] <- theta[i, sig, t] * phi_prob[sig, d, t]
      }
    }

    # Create color palette
    create_palette <- function(n) {
      if(n <= 8) {
        return(RColorBrewer::brewer.pal(n, "Set2"))
      } else {
        base_colors <- RColorBrewer::brewer.pal(8, "Set2")
        extra_colors <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
                         "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf")
        return(c(base_colors, extra_colors)[1:n])
      }
    }
    sig_colors <- create_palette(K)

    # Create plots

    theta_df$time=theta_df$time+30
    phi_df$time=phi_df$time+30

    p1 <- ggplot(theta_df, aes(x=time, y=value, color=as.factor(signature))) +
      geom_line() +
      geom_vline(xintercept=onset_t_name, linetype="dashed", color="red") +
      scale_color_manual(values=sig_colors, name="Signature") +
      #scale_alpha_identity() +
      labs(title=paste("Individual", i, "Signature Weights"),
           y="Weight",
           x="Age") +
      theme_minimal() +
      theme(legend.position="bottom")

    p2 <- ggplot(phi_df, aes(x=time, y=value, color=as.factor(signature))) +
      geom_line() +
      geom_vline(xintercept=onset_t_name, linetype="dashed", color="red") +
      scale_color_manual(values=sig_colors, name="Signature") +
      #scale_alpha_identity() +
      labs(title="MI Probabilities by Signature",
           y="Probability",
           x="Age") +
      theme_minimal() +
      theme(legend.position="bottom")

    individual_id=i
     # Get first occurrence of each disease for ordering
  disease_timeline <- data.frame(
    disease = disease_names_list,
    first_occurrence = apply(Y[individual_id,,], 1, function(x) min(which(x == 1)))
  ) %>%
    mutate(
      first_occurrence = ifelse(is.infinite(first_occurrence), Inf, first_occurrence + 30),
      ever_occurs = first_occurrence < Inf
    )

  # Get diseases that occur for this person
  person_disease <- disease_timeline %>%
    filter(ever_occurs) %>%
    arrange(first_occurrence) %>%
    pull(disease)

  disease_indices <- match(person_disease, disease_names_list)

  # 2. Disease events - now ordered by first occurrence
  diseases_df <- as.data.frame(t(Y[individual_id, disease_indices, ])) %>%
    setNames(person_disease) %>%
    mutate(Age = age_range) %>%
    pivot_longer(-Age, 
                names_to = "Disease", 
                values_to = "Status") %>%
    mutate(Disease = factor(Disease, levels = rev(person_disease)))  # Reverse order for bottom-up timeline

  p3 <- ggplot(diseases_df, aes(x = Age, y = Disease, fill = as.factor(Status))) +
    geom_tile() +
    labs(title = "Disease Events",
         y = "Disease Status",fill="Status") +
    theme_minimal() +scale_fill_manual(values = c("0" = "white", "1" = "red"))+
    theme(legend.position = "right")

  # 3. Improved genetic risk scores plot
  genetic_df <- data.frame(t(
    G[individual_id,]
  ));
  genetic_df$PRS=rownames(genetic_df)
  colnames(genetic_df)=c("Value","PRS")
  genetic_df=genetic_df%>%mutate(
      Category = case_when(
        abs(Value) >= 2 ~ "Very High",
        abs(Value) >= 1 ~ "High",
        abs(Value) >= 0.5 ~ "Moderate",
        TRUE ~ "Low"
      ),
      Direction = ifelse(Value >= 0, "Positive", "Negative"))

  p4 <- ggplot(genetic_df, aes(x = reorder(PRS, abs(Value)), y = Value)) +
    geom_col(aes(fill = Direction)) +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    geom_hline(yintercept = c(-2, -1, 1, 2), linetype = "dotted", alpha = 0.3) +
    scale_fill_manual(values = c("Negative" = "darkred", "Positive" = "steelblue")) +
    labs(title = "Genetic Risk Scores",
         subtitle = "Dotted lines at ±1 and ±2 SD",
         y = "PRS Value (SD units)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "right") +
    coord_flip()

grid.arrange(p1,p2,p3 ,p4, nrow = 2)
  # Combine plots with adjusted heights
    #combined_plot <- p1 / p2 / p3 +
    #plot_layout(heights = c(2, 2, 2)) 
}

    ```

      }