### 2025-01-11 presntations/aladynoulliplots.qmd
Update everything

---
title: "plottingAladynoulli"
format: html
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
```{r}

library(tidyverse)
library(reticulate)

# 2. Activate the environment
use_condaenv("r-tensornoulli")

# 3. Install required packages
#conda_install("r-tensornoulli", packages = c("numpy", "pandas", "scipy"))

# 4. Verify installation
py_config()  # Should show your new environment

conda_install("r-tensornoulli", packages = c("torch", "torchvision"))
# Use your conda environment
use_condaenv("r-tensornoulli")
library(reticulate)

# First, make sure we're using the right environment
use_condaenv("r-tensornoulli")

# Install PyTorch using pip
py_install("torch", pip = TRUE)

# Load the PyTorch model
model_path <- "~/Dropbox (Personal)/mgbbtopic/model.pt"
model_data <- torch$load(model_path)

# Load the saved model data
model_data <- py_load_object("~/Dropbox (Personal)//aladyn_model_full.pkl")
lambda=model_data$lambda
model_data$metadata$smoke=ifelse(model_data$metadata$smoke==1,1,0)
phi=model_data$phi
Y=model_data$Y
event_times=model_data$event_times
disease_names=model_data$disease_names
groups=model_data$groups
```

```{r}
library(reshape2)
theta=aperm(apply(lambda,c(1,3),function(x){exp(x)/sum(exp(x))}),c(2,1,3))
G=readRDS("~/tensornoulli_ehr_new/data/prs.rds")
l=melt(theta)
colnames(l)=c("Individual","Signature","Time","value")
p=ggplot(l[l$Individual%in%sample(1:9661,5),],aes(Time,value,fill=as.factor(Signature)))+geom_area()+facet_wrap(~Individual,scales = "free")+scale_fill_futurama()+labs(x="Time (years from 30)",y="Probability",title="Signature probabilities over time",fill="Signature")

ggsave(p, filename = "signature_probabilities_indvidiaul.pdf", width = 6, height = 4)

```

You can add options to executable code like this

```{r}

disease_names=readRDS("~/Dropbox (Personal)/disease_names.rds")
phi_logit=aperm(plogis(phi),c(2,1,3))
phi_logit=aperm(phi,c(2,1,3))
rownames(phi_logit)=disease_names
l=melt(phi_logit)
colnames(l)=c("Disease","Signature","Time","value")
p2=ggplot(l[l$Disease%in%c("Coronary atherosclerosis","Myocardial infarction","Angina pectors","Atrial fibrillation and flutter","Congestive heart failure (CHF) NOS"),],aes(Time,value,fill=as.factor(Signature)))+
  geom_area()+facet_wrap(~Disease)+scale_fill_futurama()+labs(x="Time (years from 30)",y="Probability",title="Signature probabilities over time",fill="Signature")
ggsave(p2, filename = "signature_probabilities_disease.pdf", width = 6, height = 4)

```

```{r}
library(ggplot2)
library(viridis)  # for better color gradients

# First get your data as before
phi = model_data$phi
phi_logit = aperm(phi,c(2,1,3))
rownames(phi_logit) = disease_names
l = melt(plogis(phi_logit))
colnames(l) = c("Disease","Signature","Time","value")

# Calculate mean values for each disease within each signature
disease_means <- l %>%
  group_by(Signature, Disease) %>%
  summarize(mean_value = mean(value), .groups = "drop") %>%
  group_by(Signature) %>%
  arrange(Signature, desc(mean_value)) %>%
  slice_head(n = 20)

# Create the filtered dataset
l_top <- l %>%
  inner_join(disease_means, by = c("Signature", "Disease")) %>%
  merge(groups, by.x = "Disease", by.y = "phenotype", all.x = TRUE)

# Create factor levels separately for each signature
l_top <- l_top %>%
  group_split(Signature) %>%
  map_dfr(~{
    sig_means <- disease_means %>% 
      filter(Signature == first(.x$Signature)) %>%
      arrange(desc(mean_value))

    .x %>% 
      mutate(Disease = factor(Disease, levels = sig_means$Disease))
  })

# Plot
p2 <- ggplot(l_top, aes(Time, Disease, fill = value)) +
  geom_tile() +
  facet_wrap(~ Signature, scales = "free_y", drop = TRUE) +
  labs(x = "Time (years from 30)", y = NULL,
       title = "Top diseases per signature over time", 
       fill = "Weight") +
  scale_fill_viridis_c(option = "B", direction = -1) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.text = element_text(size = 14),
    axis.title.x = element_text(size = 14),
    axis.ticks.y = element_blank(),
    panel.spacing = unit(1, "lines")
  )

print(p2)

ggsave(p2, filename = "signature_probabilities_disease_to2p20.pdf", width = 20, height = 6)
```

```{r}

# Function to calculate remaining lifetime risk from annual transitions
calc_remaining_risk <- function(pi_pred, from_age) {
  # Convert from 3D array to more manageable form
  N <- dim(pi_pred)[1]
  D <- dim(pi_pred)[2]
  T <- dim(pi_pred)[3]

  # For each person and disease, calculate probability of developing disease 
  # between from_age and end of follow-up
  remaining_risk <- array(0, dim=c(N, D))

  for(n in 1:N) {
    for(d in 1:D) {
      # Get annual transition probabilities from from_age onwards
      yearly_pi <- pi_pred[n, d, from_age:T]
      # Probability of developing = 1 - prob of never developing
      remaining_risk[n,d] <- 1 - prod(1 - yearly_pi)
    }
  }

  return(remaining_risk)
}

# Calculate remaining risk at different ages
ages <- seq(30, 75, by=5)
aladyn_risks <- list()

# Calculate remaining risk at different ages

for(age in ages) {
  time_idx <- age - 29  # Convert age to time index
  aladyn_risks[[as.character(age)]] <- calc_remaining_risk(pi_pred, time_idx)
}

fr=readRDS("~/Dropbox (Personal)/fr30.rds")
fr=fr[fr$eid%in%rownames(Y2),]

```

```{r}
# Function to fit Cox model and get predicted risks
# Function to fit Cox model and get predicted risks
fit_cox_risk <- function(event_times, event, sex, smoke, from_age) {
  N <- length(event_times)
  all_risks <- rep(NA, N)  # Initialize risks for everyone

  # Create survival data
  df <- data.frame(
    id = 1:N,
    time = as.numeric(event_times),
    event = event,
    sex = sex,
    smoke = smoke
  )

  # Get indices of people still at risk
  at_risk_idx <- which(df$time > from_age)

  if(length(at_risk_idx) > 0) {
    # Create filtered dataset
    df_risk <- df[at_risk_idx,]
    df_risk$time <- df_risk$time - from_age

    # Fit Cox and get risks for at-risk people at a given time (i.e., time varying covariate predictions conditional on risk free age)
    cox_fit <- coxph(Surv(time, event) ~ sex + smoke, data=df_risk)
    at_risk_risks <- 1 - exp(-predict(cox_fit, type="expected"))

    # Assign risks back to full population
    all_risks[at_risk_idx] <- at_risk_risks
  }

  return(all_risks)
}

# Calculate risks at each age
cox_risks <- list()
for(age in ages) {
  time_idx <- age - 29
  cox_risks[[as.character(age)]] <- matrix(NA, nrow=N, ncol=D)

  for(d in 1:D) {
    cox_risks[[as.character(age)]][,d] <- fit_cox_risk(
      event_times = model_data$event_times[,d]+1,
      event = rowSums(Y[,d,time_idx:T]),  # Sum future events
      sex = model_data$metadata$sex,
      smoke = model_data$metadata$smoke,
      from_age = time_idx
    )
  }
}
```

```{r}
library(pROC)

library(pROC)
# Print summary
disease_name_list <- disease_names
# Compare for each age and disease
results <- data.frame()
for(age in ages) {
  for(d in 1:D) {
    # Get predictions
    aladyn_pred <- aladyn_risks[[as.character(age)]][,d]
    cox_pred <- cox_risks[[as.character(age)]][,d]

    # Get actual outcomes (future events)
    time_idx <- age - 29
    true_outcome <- rowSums(Y[,d,time_idx:T]) > 0

    # Calculate AUCs for non-NA predictions and where we have both outcomes
    valid_idx <- !is.na(cox_pred)
    if(sum(valid_idx) > 0) {
      # Check if we have both positive and negative cases
      if(sum(true_outcome[valid_idx]) > 0 && 
         sum(!true_outcome[valid_idx]) > 0) {

        auc_aladyn <- auc(true_outcome[valid_idx], aladyn_pred[valid_idx])
        auc_cox <- auc(true_outcome[valid_idx], cox_pred[valid_idx])

        results <- rbind(results, data.frame(
          age = age,
          disease = d,
          disease_name = disease_name_list[d],
          auc_aladyn = auc_aladyn,
          auc_cox = auc_cox,
          n_at_risk = sum(valid_idx),
          n_events = sum(true_outcome[valid_idx])
        ))
      }
    }
  }
}

r=melt(results[,c("age","disease_name","auc_aladyn","auc_cox")],id=c("age","disease_name"))
groups=readRDS("~/Dropbox (Personal)/disease_groups.rds")
r=merge(r,groups,by.x="disease_name",by.y="phenotype",all.x=TRUE)
ggplot(r,aes(x=age,y=value,fill=variable))+geom_boxplot()+
  scale_fill_futurama()+labs(y="AUC",x="Age",fill="Model")+theme_classic()+theme(axis.text.x = element_text(angle = 90))

print(head(results[order(results$auc_aladyn,decreasing = T),]))

mean(results$auc_cox>results$auc_aladyn)

#saveRDS(results,"~/Dropbox (Personal)/results_without_cox_correction.rds")

```

### Now do it where we use the aladyn risks computed above as time varying predictors

```{r}
library(survival)

e2=event_times[rownames(event_times)%in%fr$eid,]
yf=Y2[rownames(Y2)%in%fr$eid,,]
all.equal(as.character(fr$eid),as.character(rownames(e2)))

disease_name_list=model_data$disease_names
# Function to fit Cox models and get predicted risks
fit_cox_models <- function(event_times,
                          event,
                          sex,
                          smoke,
                          aladyn_pred,
                          from_age) {
  N <- length(event_times)
  all_risks <- data.frame(
    cox_basic = rep(NA, N),
    cox_aladyn = rep(NA, N),
    cox_combined = rep(NA, N)
  )

  all_cs <- data.frame(
    cox_basic = NA,    # Changed: Single value per model
    cox_aladyn = NA,
    cox_combined = NA
  )

  # Create survival data
  df <- data.frame(
    id = 1:N,
    time = as.numeric(event_times),
    event = event,
    sex = sex,
    smoke = smoke,
    aladyn_pred = aladyn_pred
  )

  # Get indices of people still at risk
  at_risk_idx <- which(df$time > from_age)

  if (length(at_risk_idx) > 0) {
    df_risk <- df[at_risk_idx, ]
    df_risk$time <- df_risk$time - from_age

    # Fit models and get predictions only if we have enough data
    if(sum(df_risk$event) > 0) {  # Added check for events
      # Fit three models
      cox_basic <- coxph(Surv(time, event) ~ sex + smoke, data = df_risk)
      cox_aladyn <- coxph(Surv(time, event) ~ aladyn_pred, data = df_risk)
      cox_combined <- coxph(Surv(time, event) ~ sex + smoke + aladyn_pred, data = df_risk)

      # Get risks
      all_risks$cox_basic[at_risk_idx] <- 1 - exp(-predict(cox_basic, type = "expected"))
      all_risks$cox_aladyn[at_risk_idx] <- 1 - exp(-predict(cox_aladyn, type = "expected"))
      all_risks$cox_combined[at_risk_idx] <- 1 - exp(-predict(cox_combined, type = "expected"))

      # Store concordance
      all_cs$cox_basic <- cox_basic$concordance["concordance"]
      all_cs$cox_aladyn <- cox_aladyn$concordance["concordance"]
      all_cs$cox_combined <- cox_combined$concordance["concordance"]
    }
  }

  return(list("all_risks" = all_risks, "all_cs" = all_cs))
}

T=max(event_times) 
D=348
N=9661
# Calculate risks for all models at each age
all_risks <- list()
all_risk_cs <- list()
for(age in ages) {
  time_idx <- age - 29
  N=dim(Y)[1]
  D=dim(Y)[2]
  T=dim(Y)[3]
  all_risks[[as.character(age)]] <- array(NA, dim=c(N, D, 3))  # 3 models
  all_risk_cs[[as.character(age)]] <- array(NA, dim=c(D, 3))   # Changed: One value per disease/model

  for(d in 1:D) {
    risks <- fit_cox_models(
      event_times = event_times[,d]+1,
      event = rowSums(Y[,d,time_idx:T]),
      sex = metadata$sex,
      smoke = metadata$smoke,
      aladyn_pred = aladyn_risks[[as.character(age)]][,d],
      from_age = time_idx
    )

    all_risks[[as.character(age)]][,d,] <- as.matrix(risks$all_risks)
    all_risk_cs[[as.character(age)]][d,] <- as.numeric(risks$all_cs)  # Changed: Store concordance correctly
  }
}

# Compare performance
results <- data.frame()
for(age in ages) {
  for(d in 1:D) {
    # Get predictions from all models
    basic_pred <- all_risks[[as.character(age)]][,d,1]
    aladyn_pred <- all_risks[[as.character(age)]][,d,2]
    combined_pred <- all_risks[[as.character(age)]][,d,3]

    # Get actual outcomes
    time_idx <- age - 29
    true_outcome <- rowSums(Y[,d,time_idx:T]) > 0

    # Calculate AUCs
    valid_idx <- !is.na(basic_pred)
    if(sum(valid_idx) > 0 && 
       sum(true_outcome[valid_idx]) > 0 && 
       sum(!true_outcome[valid_idx]) > 0) {

      auc_basic <- auc(true_outcome[valid_idx], basic_pred[valid_idx])
      auc_aladyn <- auc(true_outcome[valid_idx], aladyn_pred[valid_idx])
      auc_combined <- auc(true_outcome[valid_idx], combined_pred[valid_idx])

      results <- rbind(results, data.frame(
        age = age,
        disease = d,
        disease_name = disease_name_list[d],
        auc_basic = auc_basic,
        auc_aladyn = auc_aladyn,
        auc_combined = auc_combined,
        n_at_risk = sum(valid_idx),
        n_events = sum(true_outcome[valid_idx])
      ))
    }
  }
}

# Calculate means and SDs
library(reshape2)

r=melt(results[,c("age","disease_name","auc_basic","auc_aladyn","auc_combined")],id=c("age","disease_name"))

ggplot(r,aes(x=variable,y=value,fill=variable))+geom_boxplot()+
  scale_fill_futurama()+labs(y="AUC Across all Diseases of remaining risk",x="Model",fill="Model")+theme_classic()+theme(axis.text.x = element_text(angle = 90))

# Create nicer plot
barplot(means, 
        ylim=c(0,1.1),
        col=c("lightgray","skyblue","lightgreen"),
        names.arg=c("Basic\n(sex+smoke)", "Aladynoulli", "Combined"),

        ylab="AUC Across all Diseases of remaining risk")

# Add error bars
arrows(seq(0.7,2.7,1), means-sds, seq(0.7,2.7,1), means+sds, 
       code=3, angle=90, length=0.05)

# Add numeric values
text(seq(0.7,2.7,1), means+0.05, 
     sprintf("%.3f", means))

results2 <- data.frame()
for(age in ages) {
    # Check if we have data for this age
    if(!is.null(all_risk_cs[[as.character(age)]]) && 
       !all(is.na(all_risk_cs[[as.character(age)]]))) {

        # Create temporary data frame for this age
        temp_df <- data.frame(
            age = rep(age, D),
            disease = 1:D,
            disease_name = disease_name_list
        )

        # Add concordance scores
        temp_df$c_basic <- all_risk_cs[[as.character(age)]][,1]
        temp_df$c_aladyn <- all_risk_cs[[as.character(age)]][,2]
        temp_df$c_combined <- all_risk_cs[[as.character(age)]][,3]

        # Append to results
        results2 <- rbind(results2, temp_df)
    }
}

# Calculate means and SDs
sum <- results[,c("auc_basic","auc_aladyn","auc_combined")]
means <- apply(sum,2,function(x){mean(na.omit(x))})
sds <- apply(sum,2,function(x){sd(na.omit(x))})

# Create nicer plot
barplot(means, 
        ylim=c(0,1.1),
        col=c("lightgray","skyblue","lightgreen"),
        names.arg=c("Basic\n(sex+smoke)", "Aladynoulli", "Combined"),

        ylab="Concordance Across all Diseases of remaining risk")

# Add error bars
arrows(seq(0.7,2.7,1), means-sds, seq(0.7,2.7,1), means+sds, 
       code=3, angle=90, length=0.05)

# Add numeric values
text(seq(0.7,2.7,1), means+0.05, 
     sprintf("%.3f", means))

```

```{r}
groups=readRDS("~/Dropbox (Personal)/disease_groups.rds")
results$auc_improvement <- results$auc_aladyn - results$auc_basic
results$rel_improvement <- (results$auc_aladyn - results$auc_basic)/results$auc_basic

# Look at where Aladyn helps most
top_improvements <- results[order(-results$auc_improvement),]
head(top_improvements[,c("disease_name", "age", "auc_basic", "auc_combined", 
                        "auc_improvement", "n_events")])

# 2. Disease-group specific analysis
results_by_group <- merge(results, groups[,c("phenotype","exclude_name")], 
                         by.x="disease_name", by.y="phenotype")

ggsave(ggplot(results_by_group,aes(x=exclude_name,y=auc_improvement,fill=exclude_name))+geom_boxplot()+theme_classic()+labs(y="AUC Improvement",x="Group",fill="Group")+theme(axis.text.x = element_text(angle = 90)),filename = "auc_improvement_by_group.pdf",width = 6, height = 4)

ggsave(
  ggplot(results_by_group,aes(x=age,y=auc_improvement,fill=as.factor(age),group=as.factor(age)))+geom_boxplot()+theme_classic()+labs(y="AUC Improvement",x="Age",fill="Age")+theme(axis.text.x = element_text(angle = 90))
                                                                                                                                       ,filename="auc_improvement_by_age.pdf",width = 6, height = 4)
```

```{r}
pi_pred=model_data$pi_pred
analyze_conditionals <- function(d1, d2, age) {
  time_idx <- age - 29

  # Get people at risk for d2
  at_risk <- event_times[,d2] > time_idx

  # Calculate probabilities
  marginal <- mean(pi_pred[at_risk, d2, time_idx])

  # Conditional on having d1
  has_d1 <- rowSums(Y[, d1, 1:time_idx]) > 0
  cond_yes <- mean(pi_pred[at_risk & has_d1, d2, time_idx])
  cond_no <- mean(pi_pred[at_risk & !has_d1, d2, time_idx])

  # Risk ratios
  rr <- cond_yes/cond_no

  return(data.frame(
    disease1 = disease_name_list[d1],
    disease2 = disease_name_list[d2],
    age = age,
    marginal = marginal,
    cond_yes = cond_yes,
    cond_no = cond_no,
    risk_ratio = rr
  ))
}

ages=seq(31,80,by=5)
disease_name_list <- disease_names
D=length(disease_name_list)
# Calculate for all pairs of diseases
# Look at all disease pairs at different ages
pairs_analysis <- data.frame()
for(age in ages) {
  for(d1 in 1:D) {
    for(d2 in 1:D) {
      if(d1 != d2) {
        res <- analyze_conditionals(d1, d2, age)
        pairs_analysis <- rbind(pairs_analysis, res)
      }
    }
  }
}

# Find strongest associations
top_pairs <- pairs_analysis[order(-pairs_analysis$risk_ratio),]

```

```{r}
library(tidyverse)
library(viridis)

# Function 1: Topic-specific weight importance over other groups
plot_group_deviations_within_topic <- function(l, groups, target_groups = c("circulatory system", "neoplasms", "neurological")) {

  # Calculate mean and std for each signature and time
  topic_stats <- l %>%
    group_by(Signature, Time) %>%
    summarize(mean_weight = mean(value),
             std_weight = sd(value),
             .groups = 'drop')

  # Calculate deviations for each group
  group_deviations <- map_df(target_groups, function(group) {
    # Get diseases in this group
    group_diseases <- groups %>%
      filter(exclude_name == group) %>%
      pull(phenotype)

    # Calculate group deviation
    l %>%
      filter(Disease %in% group_diseases) %>%
      left_join(topic_stats, by = c("Signature", "Time")) %>%
      group_by(Signature, Time) %>%
      summarize(
        deviation = mean((value - mean_weight) / std_weight),
        group = group,
        .groups = 'drop'
      )
  })

  # Plot
  ggplot(group_deviations, aes(x = Time, y = deviation, color = factor(Signature))) +
    geom_line(alpha = 0.7) +
    facet_wrap(~group, ncol = 1, scales = "free_y") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.3) +
     scale_color_npg() +
    labs(x = "Time", y = "Deviation (σ)", color = "Signature",
         title = "Group Deviation Within Topics") +
    theme_minimal() +
    theme(legend.position = "right")
}

# Function 2: Topic-specific group importance vs other topics
plot_group_deviations_across_topics <- function(l, groups, target_groups = c("circulatory system", "neoplasms", "neurological")) {

  # Calculate deviations for each group
  group_deviations <- map_df(target_groups, function(group) {
    # Get diseases in this group
    group_diseases <- groups %>%
      filter(exclude_name == group) %>%
      pull(phenotype)

    # Calculate mean weight across signatures for these diseases
    signature_means <- l %>%
      filter(Disease %in% group_diseases) %>%
      group_by(Time, Disease) %>%
      summarize(
        mean_across_sigs = mean(value),
        std_across_sigs = sd(value),
        .groups = 'drop'
      )

    # Calculate deviation from other signatures
    l %>%
      filter(Disease %in% group_diseases) %>%
      left_join(signature_means, by = c("Time", "Disease")) %>%
      group_by(Signature, Time) %>%
      summarize(
        deviation = mean((value - mean_across_sigs) / std_across_sigs),
        group = group,
        .groups = 'drop'
      )
  })

  # Plot
  ggplot(group_deviations, aes(x = Time, y = deviation, color = factor(Signature))) +
    geom_line(alpha = 0.7) +
    facet_wrap(~group, ncol = 1, scales = "free_y") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.3) +
    scale_color_npg() +
    labs(x = "Time", y = "Deviation from mean across signatures (σ)", 
         color = "Signature",
         title = "Group Deviation Across Topics") +
    theme_minimal() +
    theme(legend.position = "right")
}

# Usage:
p1 <- plot_group_deviations_within_topic(l, groups)
p2 <- plot_group_deviations_across_topics(l, groups)

# Display plots
print(p1)
print(p2)

# Optional: Combine plots using patchwork
library(patchwork)
combined_plot <- p1 / p2
print(combined_plot)

ggsave(p1, filename = "group_deviationsz_within.pdf", width = 8, height = 6,dpi=300)
ggsave(p2, filename = "group_deviationsz_between.pdf", width = 8, height = 6,dpi=300)
```

```{r}
library(tidyverse)
library(patchwork)
theta_pred=theta
prs=readRDS("~/tensornoulli_ehr_new/data/prs.rds")
prs_names=names(prs)

library(tidyverse)
library(patchwork)

plot_individual_analysis <- function(theta_pred, Y, G, disease_names, prs_names, 
                                   individual_id=4426, start_age=30,B=model_data$gamma) {

    # Debug prints
  cat("Dimensions and types:\n")
  cat("G dimensions:", dim(G), "\n")
  cat("B dimensions:", dim(B), "\n")
  cat("G class:", class(G), "\n")
  cat("B class:", class(B), "\n")

  # Ensure numeric matrices
  G <- as.matrix(G)
  B <- as.matrix(B)

  # Calculate number of time points and topics
  n_times <- dim(theta_pred)[3]
  n_topics <- dim(theta_pred)[2]
  age_range <- seq(start_age, start_age + n_times - 1)

  # Get individual's genetics and ensure numeric
  g_vec <- as.numeric(G[individual_id,])

  # Calculate number of time points
  n_times <- dim(theta_pred)[3]  # assuming theta_pred is [individuals, topics, time]
  age_range <- seq(start_age, start_age + n_times - 1)

  # 1. Topic trajectories
  topics_df <- as.data.frame(t(theta_pred[individual_id,,])) %>%
    setNames(paste0("Topic", 1:ncol(.))) %>%
    mutate(Age = age_range) %>%
    pivot_longer(-Age, 
                names_to = "Topic", 
                values_to = "Weight")

  p1 <- ggplot(topics_df, aes(x = Age, y = Weight, color = Topic)) +
    geom_line() +
    labs(title = paste("Topic Trajectories for Individual", individual_id),
         y = "Topic Weight") +
    theme_minimal() +
    theme(legend.position = "right")

  # 2. Disease events
  cv_diseases <- c('Coronary atherosclerosis', 'Myocardial infarction', 
                  'Angina pectoris', 'Cerebrovascular disease', 
                  'Cerebral ischemia', 'Occlusion of cerebral arteries')

  disease_indices <- match(cv_diseases, disease_names)

  diseases_df <- as.data.frame(t(Y[individual_id, disease_indices, ])) %>%
    setNames(cv_diseases) %>%
    mutate(Age = age_range) %>%
    pivot_longer(-Age, 
                names_to = "Disease", 
                values_to = "Status")

  disease_indices <- which(rowSums(Y[individual_id,,])>0)

  person_disease=disease_names[disease_indices]

  diseases_df <- as.data.frame(t(Y[individual_id, disease_indices, ])) %>%
    setNames(person_disease) %>%
    mutate(Age = age_range) %>%
    pivot_longer(-Age, 
                names_to = "Disease", 
                values_to = "Status")

  p2 <- ggplot(diseases_df, aes(x = Age, y = Disease, fill = as.factor(Status)))+
    geom_tile() +
    labs(title = "Disease Events",
         y = "Disease Status") +
    theme_minimal() +
    theme(legend.position = "right")+scale_fill_d3()

  # 3. Genetic risk scores
genetic_df <- G[individual_id,] %>%
    as.data.frame() %>%
    gather(key = "PRS", value = "Value") %>%
    mutate(PRS = str_remove(PRS, "Value."))  # Remove "Value." prefix

  p3 <- ggplot(genetic_df, aes(x = reorder(PRS, Value), y = Value)) +  # reorder by value
    geom_col(aes(fill = Value > 0)) +  # Color by positive/negative
    scale_fill_manual(values = c("darkred", "steelblue")) +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    labs(title = "Genetic Risk Scores",
         y = "PRS Value (SD units)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    coord_flip()  # Flip coordinates for better label readability

  # Combine plots
  combined_plot <- p1 / p2 / p3 +
    plot_layout(heights = c(2, 1, 1.5))  # Adjusted height for genetic plot

  # Print top genetic values
  cat("\nTop 5 highest absolute PRS values for Individual", individual_id, ":\n")
  genetic_df %>%
    arrange(desc(abs(Value))) %>%
    head(5) %>%
    with(cat(paste(PRS, ": ", sprintf("%.3f", Value), "\n", sep="")))
  return(combined_plot)

}

# Usage:
p=plot_individual_analysis(theta_pred = theta_pred, Y = Y, G = prs, disease_names, prs_names =prs_names )
ggsave(p, filename = "individual_analysis.pdf", width = 20, height = 12,dpi=300)
```

```{r}

analyze_individual_trajectory <- function(individual_id, Y, G, theta, disease_names_list, prs_names_list, age_range=30:80) {
  # Get first occurrence of each disease for ordering
  disease_timeline <- data.frame(
    disease = disease_names_list,
    first_occurrence = apply(Y[individual_id,,], 1, function(x) min(which(x == 1)))
  ) %>%
    mutate(
      first_occurrence = ifelse(is.infinite(first_occurrence), Inf, first_occurrence + 30),
      ever_occurs = first_occurrence < Inf
    )

  # Get diseases that occur for this person
  person_disease <- disease_timeline %>%
    filter(ever_occurs) %>%
    arrange(first_occurrence) %>%
    pull(disease)

  disease_indices <- match(person_disease, disease_names_list)

  # 1. Topic trajectories
  topics_df <- melt(theta_pred[individual_id,,], varnames = c("Signature", "Age"), value.name = "Weight")

  p1 <- ggplot(topics_df, aes(x = Age, y = Weight, color = as.factor(Signature)))+
    geom_line(size = 1) +
    labs(title = "Signature Trajectories",
         y = "Topic Weight") +
    theme_minimal() +
    theme(legend.position = "right") +
    scale_color_d3()

  # 2. Disease events - now ordered by first occurrence
  diseases_df <- as.data.frame(t(Y[individual_id, disease_indices, ])) %>%
    setNames(person_disease) %>%
    mutate(Age = age_range) %>%
    pivot_longer(-Age, 
                names_to = "Disease", 
                values_to = "Status") %>%
    mutate(Disease = factor(Disease, levels = rev(person_disease)))  # Reverse order for bottom-up timeline

  p2 <- ggplot(diseases_df, aes(x = Age, y = Disease, fill = as.factor(Status))) +
    geom_tile() +
    labs(title = "Disease Events",
         y = "Disease Status",fill="Status") +
    theme_minimal() +scale_fill_manual(values = c("0" = "white", "1" = "red"))+
    theme(legend.position = "right")

  # 3. Improved genetic risk scores plot
  genetic_df <- data.frame(t(
    G[individual_id,]
  ));
  genetic_df$PRS=rownames(genetic_df)
  colnames(genetic_df)=c("Value","PRS")
  genetic_df=genetic_df%>%mutate(
      Category = case_when(
        abs(Value) >= 2 ~ "Very High",
        abs(Value) >= 1 ~ "High",
        abs(Value) >= 0.5 ~ "Moderate",
        TRUE ~ "Low"
      ),
      Direction = ifelse(Value >= 0, "Positive", "Negative"))

  p3 <- ggplot(genetic_df, aes(x = reorder(PRS, abs(Value)), y = Value)) +
    geom_col(aes(fill = Direction)) +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    geom_hline(yintercept = c(-2, -1, 1, 2), linetype = "dotted", alpha = 0.3) +
    scale_fill_manual(values = c("Negative" = "darkred", "Positive" = "steelblue")) +
    labs(title = "Genetic Risk Scores",
         subtitle = "Dotted lines at ±1 and ±2 SD",
         y = "PRS Value (SD units)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "right") +
    coord_flip()

grid.arrange(p1, p2, p3, ncol = 3,widths=c(2,4,2))
  # Combine plots with adjusted heights
    #combined_plot <- p1 / p2 / p3 +
    #plot_layout(heights = c(2, 2, 2))

  return(cp)
}

prs_names=names(G)
individual_id=846
a=analyze_individual_trajectory(individual_id = individual_id,Y,G,theta,disease_names,prs_names,age_range = 30:80)
ggsave(a, filename = "individual_trajectory_earlyCAD.pdf", width = 8, height = 20,dpi=600)
```

```{r}
library(tidyverse)
library(patchwork)

plot_individual_analysis <- function(theta_pred, Y, G, B, disease_names, prs_names, 
                                   individual_id=501, start_age=30) {

  # Ensure numeric matrices and proper dimensions
  G <- as.matrix(G)
  B <- as.matrix(B)
  g_vec <- as.numeric(G[individual_id,])

  # Calculate number of time points and topics
  n_times <- dim(theta_pred)[3]
  n_topics <- dim(theta_pred)[2]
  age_range <- seq(start_age, start_age + n_times - 1)

  # 1. Topic trajectories
  topics_df <- as.data.frame(t(theta_pred[individual_id,,])) %>%
    setNames(paste0("Topic", 1:ncol(.))) %>%
    mutate(Age = age_range) %>%
    pivot_longer(-Age, 
                names_to = "Topic", 
                values_to = "Weight")

  p1 <- ggplot(topics_df, aes(x = Age, y = Weight, color = Topic)) +
    geom_line(linewidth = 1) +
    labs(title = paste("Topic Trajectories for Individual", individual_id),
         y = "Topic Weight") +
    theme_minimal() +
    theme(legend.position = "right")

  # 2. Disease events
  cv_diseases <- c('Coronary atherosclerosis', 'Myocardial infarction', 
                  'Angina pectoris', 'Cerebrovascular disease', 
                  'Cerebral ischemia', 'Occlusion of cerebral arteries')

  disease_indices <- match(cv_diseases, disease_names)

  diseases_df <- as.data.frame(t(Y[individual_id, disease_indices, ])) %>%
    setNames(cv_diseases) %>%
    mutate(Age = age_range) %>%
    pivot_longer(-Age, 
                names_to = "Disease", 
                values_to = "Status")

  # Find first occurrence of each disease
  disease_onsets <- diseases_df %>%
    group_by(Disease) %>%
    filter(Status == 1) %>%
    summarize(Onset = first(Age)) %>%
    filter(!is.na(Onset))

  p2 <- ggplot(diseases_df, aes(x = Age, y = Status, color = Disease)) +
    geom_line(linewidth = 1) +
    geom_vline(data = disease_onsets, 
               aes(xintercept = Onset, color = Disease),
               linetype = "dashed", alpha = 0.5) +
    labs(title = "Disease Events",
         y = "Disease Status") +
    theme_minimal() +
    theme(legend.position = "right")

  # 3. Genetic risk scores
  genetic_df <- data.frame(
    PRS = prs_names,
    Value = g_vec
  )

  p3 <- ggplot(genetic_df, aes(x = reorder(PRS, Value), y = Value)) +
    geom_col(aes(fill = Value > 0)) +
    scale_fill_manual(values = c("darkred", "steelblue")) +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    labs(title = "Genetic Risk Scores",
         y = "PRS Value (SD units)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    coord_flip()

  # 4. Calculate genetic contributions to topics
  contrib_list <- list()
  for(k in 1:n_topics) {
    b_vec <- as.numeric(B[,k])
    contrib <- g_vec * b_vec
    contrib_list[[k]] <- data.frame(
      PRS = prs_names,
      Topic = paste0("Topic", k),
      Contribution = contrib
    )
  }

  contrib_df <- do.call(rbind, contrib_list)

  # Calculate total contribution for each topic
  topic_totals <- contrib_df %>%
    group_by(Topic) %>%
    summarize(Total = sum(Contribution))

  p4 <- ggplot(contrib_df, aes(x = reorder(Topic, -Contribution), y = Contribution, fill = as.factor(PRS))) +
    geom_col(position = "stack") +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    #geom_text(data = topic_totals, 
              #aes(x = Topic, y = Total, label = sprintf("%.2f", Total)),
              #position = position_dodge(width = 0.9),
              #vjust = -0.5) +
    labs(title = "Genetic Contributions to Topics",
         y = "Combined Genetic Effect") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  # Combine plots
  combined_plot <- (p1 / p2 / p3 / p4) +
    plot_layout(heights = c(2, 1, 1.5, 1.5))

  # Print summary statistics
  cat("\nDisease Onsets:\n")
  print(disease_onsets)

  cat("\nTop 5 highest absolute PRS values:\n")
  genetic_df %>%
    arrange(desc(abs(Value))) %>%
    head(5) %>%
    with(cat(paste(PRS, ": ", sprintf("%.3f", Value), "\n", sep="")))

  cat("\nStrongest genetic contributions by topic:\n")
  contrib_df %>%
    group_by(Topic) %>%
    arrange(desc(abs(Contribution))) %>%
    slice(1) %>%
    ungroup() %>%
    arrange(Topic) %>%
    with(cat(paste(Topic, ": ", PRS, " (", sprintf("%.3f", Contribution), ")\n", sep="")))

  return(combined_plot)
}

# Usage:
p2=plot_individual_analysis(theta_pred, Y, G=prs, B=model_data$gamma, 
                        disease_names, prs_names)

ggsave(p2, filename = "individual_analysis_2.pdf", width = 8, height = 12,dpi=300)
```

```{r}
library(tidyverse)
library(pheatmap)
library(viridis)
library(RColorBrewer)
library(grid)
library(gridExtra)

plot_topic_group_enrichment <- function(phi, disease_names, groups, time_points=50) {
  phi=plogis(phi)
  # Define groups to exclude
  exclude_groups <- c("infectious diseases", "injuries & poisonings", "", "symptoms")

  # Filter groups
  groups_filtered <- groups %>%
    filter(!exclude_name %in% exclude_groups)

  # Create different color palettes for each topic
  color_palettes <- list(
    viridis(100),              # viridis
    magma(100),               # magma
    plasma(100),              # plasma
    inferno(100),             # inferno
    colorRampPalette(brewer.pal(9, "YlOrRd"))(100),   # Yellow-Orange-Red
    colorRampPalette(brewer.pal(9, "YlGnBu"))(100),   # Yellow-Green-Blue
    colorRampPalette(brewer.pal(9, "RdPu"))(100),     # Red-Purple
    colorRampPalette(brewer.pal(9, "PuBuGn"))(100),   # Purple-Blue-Green
    colorRampPalette(brewer.pal(9, "PuRd"))(100),     # Purple-Red
    colorRampPalette(brewer.pal(9, "GnBu"))(100)      # Green-Blue
  )

  # Convert phi to array if it's a tensor
  phi_array <- as.array(phi)
  n_topics <- dim(phi_array)[1]

  # Calculate mean logOR for each topic, group, and time
  enrichment_by_topic <- map(1:n_topics, function(k) {
    # Get weights for this topic
    topic_weights <- phi_array[k,,]

    # Calculate group means at each time point
    remaining_groups <- unique(groups_filtered$exclude_name)
    group_means <- matrix(nrow=length(remaining_groups), 
                         ncol=time_points)
    rownames(group_means) <- remaining_groups
    colnames(group_means) <- paste0("T", 1:time_points)

    for(t in 1:time_points) {
      weights_t <- topic_weights[,t]
      # Calculate mean weight for each group
      group_means[,t] <- sapply(remaining_groups, function(g) {
        group_diseases <- which(disease_names %in% 
                              groups_filtered$phenotype[groups_filtered$exclude_name == g])
        if(length(group_diseases) > 0) {
          mean(weights_t[group_diseases])
        } else {
          NA
        }
      })
    }

    return(group_means)
  })

  # Plot heatmap for each topic and convert to grob
  plot_grobs <- map(1:n_topics, function(k) {
    data <- enrichment_by_topic[[k]]

    # Create nicer row names
    rownames(data) <- stringr::str_to_title(rownames(data))

    # Get color palette for this topic
    palette_idx <- ((k-1) %% length(color_palettes)) + 1

    # Create pheatmap and convert to grob
    ph <- pheatmap(data,
                  main = paste("Topic", k, "Mean Probability by Group"),
                  color = color_palettes[[palette_idx]],
                  cluster_rows = FALSE,
                  cluster_cols = FALSE,
                  show_colnames = FALSE,
                  fontsize = 8,
                  angle_col = 45,
                  silent = TRUE)  # Important: return the pheatmap object

    # Convert to grob
    grid.grabExpr(grid.draw(ph$gtable))
  })

  # Return both the plots and the data
  return(list(
    plots = plot_grobs,
    data = enrichment_by_topic
  ))
}

# Usage:
result <- plot_topic_group_enrichment(phi, disease_names, groups)

# Arrange plots in a grid
do.call(grid.arrange, c(result$plots, ncol=2))

# Optionally, save the plot
# ggsave("topic_enrichment.pdf", marrangeGrob(grobs=result$plots, ncol=2, nrow=5), width=15, height=30)
```

```{r}

library(tidyverse)
library(pheatmap)
library(RColorBrewer)

plot_genetic_effects <- function(gamma, prs_names, n_topics) {
  # Convert gamma to matrix if needed
  gamma_matrix <- t(as.matrix(gamma))  # transpose to match Python version
  rownames(gamma_matrix) <- paste0("Topic ", 0:(n_topics-1))
  colnames(gamma_matrix) <- prs_names

  # Create color palette centered at 0
  colors <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100)

  # Create breaks that are symmetric around 0
  max_abs <- max(abs(gamma_matrix))
  breaks <- seq(-max_abs, max_abs, length.out = 101)

  # Create the heatmap
  ph <- pheatmap(gamma_matrix,
                 color = colors,
                 breaks = breaks,
                 cluster_rows = FALSE,      # cluster topics
                 cluster_cols = TRUE,      # cluster genetic effects
                 show_rownames = TRUE,
                 show_colnames = TRUE,
                 main = "Genetic Effects (Gamma) Across Topics",
                 fontsize = 8,
                 angle_col = 45,
                 treeheight_row = 20,      # height of dendrogram
                 treeheight_col = 20)

   # Return both plots
  return(
    heatmap = ph,

  )
}

# Usage:
plots <- plot_genetic_effects(model_data$gamma, prs_names, n_topics=10)

# Display plots
# The heatmap
plots$heatmap

# The Topic 7 barplot
plots$topic7_plot

# Optionally combine plots
library(gridExtra)
grid.arrange(
  grid.grabExpr(grid.draw(plots$heatmap$gtable)),
  plots$topic7_plot,
  ncol = 1,
  heights = c(1.5, 1)
)

# Get clustering order
cluster_order <- plots$heatmap$tree_col$order
print("PRS clustering order:")
print(prs_names_list[cluster_order])

```

```{r}
library(tidyverse)
library(pheatmap)
library(viridis)
library(RColorBrewer)

plot_topic_group_enrichment <- function(phi, disease_names, groups, time_points=50) {

  # Define groups to exclude
  exclude_groups <- c("infectious diseases", "injuries & poisonings", "", "symptoms")

  # Filter groups
  groups_filtered <- groups %>%
    filter(!exclude_name %in% exclude_groups)

  # Convert phi to array if it's a tensor
  phi_array <- as.array(phi)
  n_topics <- dim(phi_array)[1]

  # Calculate probabilities for each topic, group, and time
  enrichment_by_topic <- map(1:n_topics, function(k) {
    # Get weights for this topic
    topic_weights <- phi_array[k,,]

    # Calculate group means at each time point
    remaining_groups <- unique(groups_filtered$exclude_name)
    group_means <- matrix(nrow=length(remaining_groups), 
                         ncol=time_points)
    rownames(group_means) <- remaining_groups
    colnames(group_means) <- paste0("T", 1:time_points)

    for(t in 1:time_points) {
      weights_t <- topic_weights[,t]
      # Calculate mean probability for each group
      group_means[,t] <- sapply(remaining_groups, function(g) {
        group_diseases <- which(disease_names %in% 
                              groups_filtered$phenotype[groups_filtered$exclude_name == g])
        if(length(group_diseases) > 0) {
          mean(plogis(weights_t[group_diseases]))  # Convert to probability
        } else {
          NA
        }
      })
    }

    return(group_means)
  })

  # Find global min and max for consistent scale
  all_values <- unlist(enrichment_by_topic)
  value_range <- range(all_values, na.rm = TRUE)

  # Create color palette (using magma for better intensity visualization)
  colors <- viridis(100, option = "plasma")

  # Plot heatmap for each topic
  plot_grobs <- map(1:n_topics, function(k) {
    data <- enrichment_by_topic[[k]]

    # Create nicer row names
    rownames(data) <- stringr::str_to_title(rownames(data))

    # Create pheatmap with consistent scale
    ph <- pheatmap(data,
                  main = paste("Topic", k, "Mean Probability by Group"),
                  color = colors,
                  cluster_rows = FALSE,
                  cluster_cols = FALSE,
                  show_colnames = FALSE,
                  breaks = seq(value_range[1], value_range[2], length.out = 101),
                  fontsize = 8,
                  angle_col = 45,
                  silent = TRUE)

    # Convert to grob
    grid.grabExpr(grid.draw(ph$gtable))
  })

  # Return both the plots and the data
  return(list(
    plots = plot_grobs,
    data = enrichment_by_topic,
    value_range = value_range
  ))
}

# Usage:
result <- plot_topic_group_enrichment(phi, disease_names, groups)

# Print value range
cat("Probability range:", sprintf("%.4f to %.4f\n", result$value_range[1], result$value_range[2]))

# Arrange plots in a grid
do.call(grid.arrange, c(result$plots, ncol=3))

# Optionally save the plot
ggsave("topic_probabilities.pdf", marrangeGrob(grobs=result$plots, ncol=3, nrow=3),
       width=40, height=30, dpi=600)
```

```{r}
library(tidyverse)
library(pheatmap)
library(viridis)

plot_group_transitions <- function(Y, pi_pred, disease_names, groups, age_bins = seq(30, 80, 10)) {

  # Filter out unwanted groups
  exclude_groups <- c("infectious diseases", "injuries & poisonings", "", "symptoms","congenital anomalies")
  key_groups <- unique(groups$exclude_name[!groups$exclude_name %in% exclude_groups])

  # Create group dictionary
  group_dict <- map(key_groups, function(group) {
    which(disease_names %in% groups$phenotype[groups$exclude_name == group])
  })
  names(group_dict) <- key_groups

  # For each age window
  for(i in 2:(length(age_bins)-1)) {
    age_start <- age_bins[i] - 30
    age_end <- age_bins[i+1] - 30

    # Initialize matrices
    marginal_probs <- numeric(length(key_groups))
    conditional_matrix <- matrix(0, nrow = length(key_groups), ncol = length(key_groups))
    ratio_matrix <- matrix(0, nrow = length(key_groups), ncol = length(key_groups))
    counts_matrix <- matrix(0, nrow = length(key_groups), ncol = length(key_groups))

    names(marginal_probs) <- key_groups
    rownames(conditional_matrix) <- colnames(conditional_matrix) <- key_groups
    rownames(ratio_matrix) <- colnames(ratio_matrix) <- key_groups

    # Calculate marginal probabilities
    for(idx1 in 1:length(key_groups)) {
      group1 <- key_groups[idx1]
      diseases1 <- group_dict[[group1]]

      at_risk <- which(rowSums(Y[, diseases1, 1:age_start]) == 0)

      if(length(at_risk) >= 100) {
        group_risk <- rep(0, length(at_risk))
        for(d in diseases1) {
          # Extract probabilities for this disease
          yearly_pi <- pi_pred[at_risk, d, (age_start+1):age_end]
          # Calculate cumulative risk
          if(length(dim(yearly_pi)) == 2) {
            disease_risk <- 1 - apply(1 - yearly_pi, 1, prod)
          } else {
            disease_risk <- 1 - prod(1 - yearly_pi)
          }
          group_risk <- pmax(group_risk, disease_risk)
        }
        marginal_probs[idx1] <- mean(group_risk)
      }
    }

    # Calculate conditional probabilities
    for(idx1 in 1:length(key_groups)) {
      for(idx2 in 1:length(key_groups)) {
        if(idx1 != idx2) {
          group1 <- key_groups[idx1]
          group2 <- key_groups[idx2]

          diseases1 <- group_dict[[group1]]
          diseases2 <- group_dict[[group2]]

          at_risk <- which(rowSums(Y[, diseases1, 1:age_start]) == 0)

          if(age_start > 0) {
            had_prev <- which(rowSums(Y[, diseases2, 1:age_start]) > 0)
            mask <- intersect(at_risk, had_prev)
            counts_matrix[idx1, idx2] <- length(mask)

            if(length(mask) >= 100) {
              group_risk_cond <- rep(0, length(mask))
              for(d in diseases1) {
                # Extract probabilities for this disease
                yearly_pi <- pi_pred[mask, d, (age_start+1):age_end]
                # Calculate cumulative risk
                if(length(dim(yearly_pi)) == 2) {
                  disease_risk <- 1 - apply(1 - yearly_pi, 1, prod)
                } else {
                  disease_risk <- 1 - prod(1 - yearly_pi)
                }
                group_risk_cond <- pmax(group_risk_cond, disease_risk)
              }
              conditional_matrix[idx1, idx2] <- mean(group_risk_cond)

              if(marginal_probs[idx1] > 0) {
                ratio_matrix[idx1, idx2] <- conditional_matrix[idx1, idx2] / marginal_probs[idx1]
              }
            }
          }
        }
      }
    }

    # Create plots
    # 1. Marginal probabilities bar plot
    p1 <- ggplot(data.frame(Group = names(marginal_probs), 
                           Probability = marginal_probs)) +
      geom_col(aes(x = reorder(Group, Probability), y = Probability,fill=Group)) +
      coord_flip() +
      labs(title = sprintf("Marginal Risk by Group\nAge %d-%d", age_bins[i], age_bins[i+1]),
           x = "Disease Group", y = "Risk") +
      theme_minimal()

    # 2. Conditional probabilities heatmap
  # Ensure unique breaks
  max_cond <- max(conditional_matrix, na.rm = TRUE)
  breaks_cond <- unique(seq(0, max_cond * 1.1, length.out = 100))
  if(length(breaks_cond) < 2) breaks_cond <- c(0, max_cond * 1.1)

  p2 <- pheatmap(conditional_matrix,
                 main = sprintf("Conditional Risks\nAge %d-%d", age_bins[i], age_bins[i+1]),
                 color = viridis(length(breaks_cond) - 1, option ="H" ),
                 breaks = breaks_cond,
                 display_numbers = TRUE,
                 number_format = "%.3f",
                 cluster_rows = FALSE,
                 cluster_cols = FALSE)

    return(list(
      marginal = marginal_probs,
      conditional = conditional_matrix,

      counts = counts_matrix,
      plots = list(marginal = p1, conditional = p2)
    ))
  }
}

# Usage:
results <- plot_group_transitions(Y, pi_pred, disease_names_list, groups)

# Display plots
grid.arrange(
  results$plots$marginal,
  grid.grabExpr(grid.draw(results$plots$conditional$gtable)),
  grid.grabExpr(grid.draw(results$plots$ratio$gtable)),
  ncol = 3
)

results <- plot_group_transitions(Y, pi_pred, disease_names_list, groups)

```

```{r}
analyze_signature_transitions <- function(theta, threshold = 0.2) {
  # theta dimensions: [n_people, n_signatures, n_times]
  n_people <- dim(theta)[1]
  n_signatures <- dim(theta)[2]
  n_times <- dim(theta)[3]

  # Create plot layout
  par(mfrow = c(2,2))

  # 1. Average signature strengths over time
  mean_theta <- apply(theta, c(2,3), mean)  # Average across people
  matplot(1:n_times, t(mean_theta), type = "l", 
          main = "Average Signature Strengths Over Time",
          xlab = "Time", ylab = "Average θ")
  legend("topright", legend = paste("Signature", 0:(n_signatures-1)), 
         col = 1:n_signatures, lty = 1)

  # 2. Calculate transition matrix
  transitions <- matrix(0, n_signatures, n_signatures)

  for(person in 1:n_people) {
    for(t in 1:(n_times-1)) {
      # Get dominant signatures (above threshold)
      current <- which(theta[person, , t] > threshold)
      next_t <- which(theta[person, , t+1] > threshold)

      # Record transitions
      if(length(current) > 0 && length(next_t) > 0) {
        for(i in current) {
          for(j in next_t) {
            transitions[j, i] <- transitions[j, i] + 1
          }
        }
      }
    }
  }

  # Normalize by row sums (probability of transitioning TO each signature)
  transitions_norm <- sweep(transitions, 2, colSums(transitions) + 1e-10, "/")

  # Plot transition heatmap
  library(pheatmap)
  pheatmap(transitions,
           main = "Signature Transition Probabilities",
           display_numbers = TRUE,
           number_format = "%.2f",
           cluster_rows = FALSE,
           cluster_cols = FALSE,
           labels_row = paste("Sig", 0:(n_signatures-1)),
           labels_col = paste("Sig", 0:(n_signatures-1)))

  # 3. Example individual trajectories
  # Find people with most signature switches
  n_switches <- numeric(n_people)
  for(person in 1:n_people) {
    dominant_sigs <- apply(theta[person, , ], 2, which.max)
    n_switches[person] <- sum(diff(dominant_sigs) != 0)
  }

  interesting_people <- order(n_switches, decreasing = TRUE)[1:5]
  m=melt(theta[interesting_people, , ], varnames = c("Person", "Signature", "Time"))
  ggplot(m,aes(x=Time,y=value,fill=as.factor(Signature)))+scale_fill_simpsons()+
    geom_area()+
    facet_wrap(~Person)+
    labs(title="Example Individual Signature Trajectories\n(Top signature switchers)",
         x="Time",y="θ",fill="Signature")
  # Plot trajectories for interesting people
  matplot(1:n_times, t(theta[interesting_people[1], , ]), type = "l",
          main = "Example Individual Signature Trajectories\n(Top signature switchers)",
          xlab = "Time", ylab = "θ")

  return(transitions_norm)
}

# Usage:
transitions <- analyze_signature_transitions(theta)
```

```{r}
find_and_plot_exemplar_cases <- function(model, Y, G, disease_names_list, prs_names_list, n_cases=20) {
  K <- 10  # Number of topics

  # PRS to phenotype mapping
  prs_to_phenotype <- list(
    "AMD" = "Macular degeneration (senile) of retina NOS",
    "AD" = "Alzheimer's disease",
    "AST" = "Asthma",
    "AF" = "Atrial fibrillation",
    "BD" = "Bipolar",
    "BMI" = "Overweight, obesity and other hyperalimentation",
    "CRC" = "Colorectal cancer",
    "BC" = "Malignant neoplasm of female breast",
    "CED" = "Celiac disease",
    "CAD" = "Coronary atherosclerosis",
    "EOC" = "Malignant neoplasm of ovary",
    "T1D" = "Type 1 diabetes",
    "T2D" = "Type 2 diabetes",
    "UC" = "Ulcerative colitis",
    "VTE" = "Other venous embolism and thrombosis",
    "RA" = "Rheumatoid arthritis",
    "SLE" = "Systemic lupus erythematosus",
    "SCZ" = "Schizophrenia",
    "MS" = "Multiple sclerosis",
    "PD" = "Parkinson's disease"
  )

  # Get model parameters
  theta_np <- theta_pred  # Assuming these are already in R format
  gamma_np <- model_data$gamma

  # Calculate genetic contributions for each person
  genetic_contributions <- G %*% gamma_np  # N x K

  # Calculate disease burden
  disease_burden <- apply(Y, 1, sum)

  # Find individuals with many diseases and strong genetic alignment
  scores <- list()
  for(i in 1:dim(Y)[1]) {
    # Get their dominant topic
    dominant_topic <- which.max(colMeans(theta_np[i,,]))
    if(dominant_topic > K) dominant_topic <- K  # Safety check

    # Score based on genetic prediction and disease burden
    genetic_score <- genetic_contributions[i, dominant_topic]
    score <- genetic_score * disease_burden[i]
    scores[[i]] <- c(i, score)
  }

  # Get top cases
  scores_mat <- do.call(rbind, scores)
  top_cases <- scores_mat[order(scores_mat[,2], decreasing=TRUE)[1:n_cases],]

  # Plot each case
  for(case in 1:nrow(top_cases)) {
    idx <- top_cases[case,1]
    score <- top_cases[case,2]

    cat(sprintf("\nAnalyzing Individual %d\n", idx))

    # Set up plotting area
    par(mfrow=c(3,1), mar=c(4,10,4,2))

    # 1. Disease timeline
    disease_events <- list()
    for(d_idx in 1:dim(Y)[2]) {
      events <- which(Y[idx,d_idx,] == 1)
      if(length(events) > 0) {
        disease_name <- disease_names_list[d_idx]
        is_prs_disease <- disease_name %in% unlist(prs_to_phenotype)
        disease_events[[length(disease_events) + 1]] <- list(
          disease = disease_name,
          events = events + 30,
          is_prs_disease = is_prs_disease
        )
      }
    }

    # Sort by first occurrence
    first_events <- sapply(disease_events, function(x) min(x$events))
    disease_events <- disease_events[order(first_events)]

    # Plot disease timeline
    plot(NULL, xlim=c(30, 30+dim(Y)[3]), ylim=c(0, length(disease_events)),
         ylab="", xlab="Age", main="Disease Timeline (* = Has PRS)")

    for(i in 1:length(disease_events)) {
      d <- disease_events[[i]]
      points(d$events, rep(i, length(d$events)), 
             col=ifelse(d$is_prs_disease, "red", "blue"),
             pch=19, cex=1)
    }

    # Add disease labels
    disease_labels <- sapply(disease_events, function(x) 
    paste0(substr(x$disease, 1, 50), ifelse(x$is_prs_disease, "*", "")))
    axis(2, at=1:length(disease_events), labels=disease_labels, las=2)
    grid(nx=NULL, ny=NA)

    # 2. Topic trajectories
    matplot(30:(30+dim(Y)[3]-1), t(theta_np[idx,,]), type="l", lty=1,
            main=paste0("Topic Trajectories for",idx), xlab="Age", ylab="Topic Weight",col=1:K)
    legend("topright", legend=paste("Topic", 1:K), col=1:K, lty=1)
    grid()

    # 3. Genetic contributions
    individual_genetics <- G[idx,]
    topic_contributions <- individual_genetics %*% gamma_np
    barplot(topic_contributions, main="Genetic Contributions to Topics",
            xlab="Topic", ylab="Genetic Contribution")
    grid()

    # Print matching PRS and phenotypes
    cat("\nMatching PRS and phenotypes:\n")
    for(prs in names(prs_to_phenotype)) {
      phenotype <- prs_to_phenotype[[prs]]
      if(any(sapply(disease_events, function(x) x$disease == phenotype))) {
        prs_value <- individual_genetics[which(prs_names_list == prs)]
        cat(sprintf("%s (value: %.3f) -> %s\n", 
                   prs, prs_value, phenotype))
      }
    }
  }
}

# Usage:
find_and_plot_exemplar_cases(model, Y, G, disease_names_list, prs_names)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(viridis)

analyze_patient <- function(idx, Y, G, theta_pred, gamma_np, disease_names_list, 
                          prs_names_list, prs_to_phenotype, K=10) {

  # Create consistent colors for topics
  topic_colors <- scales::hue_pal()(K)

  # Helper function to create timeline data
  create_timeline_data <- function(Y, idx, disease_names_list) {
    timeline_data <- data.frame()

    for(d_idx in 1:dim(Y)[2]) {
      events <- which(Y[idx, d_idx,] == 1)
      if(length(events) > 0) {
        disease_name <- disease_names_list[d_idx]
        is_prs_disease <- disease_name %in% unlist(prs_to_phenotype)

        timeline_data <- rbind(timeline_data, 
                             data.frame(
                               disease = disease_name,
                               age = events + 30,
                               is_prs = is_prs_disease
                             ))
      }
    }

    # Sort by first occurrence
    timeline_data <- timeline_data %>%
      group_by(disease) %>%
      arrange(age) %>%
      ungroup()

    return(timeline_data)
  }

  # Create all visualizations
  timeline_data <- create_timeline_data(Y, idx, disease_names_list)

  # 1. Disease Timeline Plot
  timeline_plot <- ggplot(timeline_data, aes(x = age, y = reorder(disease, age))) +
    geom_point(aes(color = is_prs), size = 3) +
    scale_color_manual(values = c("FALSE" = "#2C3E50", "TRUE" = "#E74C3C"),
                      labels = c("No PRS", "Has PRS")) +
    theme_minimal() +
    theme(
      axis.text.y = element_text(size = 8),
      panel.grid.major.y = element_line(color = "gray90"),
      legend.position = "top"
    ) +
    labs(
      title = "Disease Timeline",
      x = "Age",
      y = NULL,
      color = "PRS Status"
    )

  # 2. Topic Trajectories Plot
  trajectory_data <- as.data.frame(t(theta_pred[idx,,]))
  trajectory_data$age <- 30:(30+dim(Y)[3]-1)
  trajectory_data_long <- trajectory_data %>%
    pivot_longer(-age, names_to = "topic", values_to = "weight") %>%
    mutate(topic = factor(topic))

  trajectory_plot <- ggplot(trajectory_data_long, 
                          aes(x = age, y = weight, color = topic)) +
    geom_line(size = 1) +
    scale_color_manual(values = topic_colors) +
    theme_minimal() +
    theme(legend.position = "right") +
    labs(
      title = paste("Topic Trajectories for Individual", idx),
      x = "Age",
      y = "Topic Weight",
      color = "Topic"
    )

  # 3. Genetic Contributions Plot
  genetic_data <- data.frame(
    topic = factor(1:K),
    contribution = as.vector(G[idx,] %*% gamma_np)
  )

  genetic_plot <- ggplot(genetic_data, aes(x = topic, y = contribution)) +
    geom_col(aes(fill = topic)) +
    scale_fill_manual(values = topic_colors) +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(
      title = "Genetic Contributions to Topics",
      x = "Topic",
      y = "Contribution"
    )

  # Print matching PRS and phenotypes
  print_prs_matches <- function() {
    cat("\nPRS-Phenotype Matches for Individual", idx, "\n")
    cat(rep("-", 50), "\n")

    matches <- data.frame()
    for(prs in names(prs_to_phenotype)) {
      phenotype <- prs_to_phenotype[[prs]]
      if(phenotype %in% timeline_data$disease) {
        prs_value <- G[idx, which(prs_names_list == prs)]
        matches <- rbind(matches, 
                        data.frame(
                          PRS = prs,
                          Value = round(prs_value, 3),
                          Phenotype = phenotype
                        ))
      }
    }

    if(nrow(matches) > 0) {
      print(matches, row.names = FALSE)
    } else {
      cat("No matching PRS-phenotype pairs found.\n")
    }
  }

  # Calculate disease burden and genetic score
  disease_burden <- sum(Y[idx,,])
  genetic_score <- max(rowMeans(theta_pred[idx,,]))  # Fixed to rowMeans

  # Print summary statistics
  cat("\nSummary Statistics for Individual", idx, "\n")
  cat(rep("-", 50), "\n")
  cat(sprintf("Disease Burden: %d conditions\n", disease_burden))
  cat(sprintf("Maximum Topic Weight: %.3f\n", genetic_score))

  # Arrange and display plots
  grid.arrange(
    timeline_plot, 
    trajectory_plot,
    genetic_plot,
    ncol = 1,
    heights = c(1.5, 1, 1)
  )

  # Print PRS matches
  print_prs_matches()
}

# Function to find and analyze top cases
analyze_top_cases <- function(Y, G, theta_pred, gamma_np, disease_names_list, 
                            prs_names_list, prs_to_phenotype, n_cases = 5) {

  # Calculate scores
  disease_burden <- apply(Y, 1, sum)
  genetic_contributions <- G %*% gamma_np
  scores <- numeric(dim(Y)[1])

  for(i in 1:dim(Y)[1]) {
    dominant_topic <- which.max(rowMeans(theta_pred[i,,]))  # Fixed to rowMeans
    genetic_score <- genetic_contributions[i, dominant_topic]
    scores[i] <- genetic_score * disease_burden[i]
  }

  # Get top cases
  top_indices <- order(scores, decreasing = TRUE)[1:n_cases]

  # Analyze each top case
  for(idx in top_indices) {
    cat("\n\nAnalyzing Individual", idx, "(Score:", round(scores[idx], 2), ")\n")
    cat(rep("=", 70), "\n")

    analyze_patient(idx, Y, G, theta_pred, gamma_np, disease_names_list,
                   prs_names_list, prs_to_phenotype)

    cat("\nPress [Enter] to continue...\n")
    readline()
  }
}

# Analyze top 5 cases
analyze_top_cases(Y, G, theta_pred, gamma_np, disease_names_list, 
                 prs_names, prs_to_phenotype, n_cases = 5)

# Or analyze a specific patient
analyze_patient(1, Y, G, theta_pred, gamma_np, disease_names_list,
               prs_names, prs_to_phenotype)
```

```{r}
 analyze_prediction_timing <- function(Y, pi_pred, window_size=5) {
  N <- dim(Y)[1]
  D <- dim(Y)[2]
  T <- dim(Y)[3]

  # Store predictions before events
  pre_event_preds <- list()
  control_preds <- list()

  for(d in 1:D) {
    for(i in 1:N) {
      # Find first event for this person and disease
      event_time <- which(Y[i,d,] == 1)[1]

      if(!is.na(event_time) && event_time > window_size) {
        # Get predictions leading up to event
        pre_event_window <- (event_time-window_size):(event_time-1)
        pre_event_preds[[length(pre_event_preds) + 1]] <- 
          pi_pred[i,d,pre_event_window]

        # Get matched control period from someone without the event
        control_ids <- which(apply(Y[,d,], 1, sum) == 0)
        if(length(control_ids) > 0) {
          control_id <- sample(control_ids, 1)
          control_preds[[length(control_preds) + 1]] <- 
            pi_pred[control_id,d,pre_event_window]
        }
      }
    }
  }

  # Convert lists to vectors
  pre_event_preds <- unlist(pre_event_preds)
  control_preds <- unlist(control_preds)

  # Plot
  par(mfrow=c(2,1))

  # 1. Distribution comparison
  plot(density(pre_event_preds), 
       main="Predictions Before Events vs Controls",
       xlab="Predicted Probability",
       col="red")
  lines(density(control_preds), col="blue")
  legend("topright", 
         legend=c("Pre-Event", "Controls"),
         col=c("red", "blue"),
         lty=1)

  # 2. Time-to-event analysis
  boxplot(split(pre_event_preds, 
                rep(window_size:1, length(pre_event_preds)/window_size)),
          main="Predictions Leading up to Events",
          xlab="Time to Event",
          ylab="Predicted Probability")

  # Statistical tests
  # 1. Compare pre-event vs control predictions
  wilcox_test <- wilcox.test(pre_event_preds, control_preds)

  # 2. Test for trend in pre-event predictions
  time_to_event <- rep(window_size:1, length(pre_event_preds)/window_size)
  trend_test <- cor.test(time_to_event, pre_event_preds, method="spearman")

  # Print results
  cat("\nStatistical Tests:\n")
  cat("-----------------\n")
  cat("Pre-event vs Control Wilcoxon test:\n")
  cat(sprintf("  p-value: %.4e\n", wilcox_test$p.value))
  cat("\nTrend analysis (Spearman correlation):\n")
  cat(sprintf("  rho: %.4f\n", trend_test$estimate))
  cat(sprintf("  p-value: %.4e\n", trend_test$p.value))

  return(list(
    pre_event_preds = pre_event_preds,
    control_preds = control_preds,
    wilcox_test = wilcox_test,
    trend_test = trend_test
  ))
 }
```

```{r}
# First, get top diseases
get_top_diseases <- function(phi, n_top=10) {
  mean_loadings <- apply(phi, c(1,2), mean)
  top_diseases <- lapply(1:dim(phi)[1], function(k) {
    topic_loadings <- scale(mean_loadings[k,])
    order(topic_loadings, decreasing=TRUE)[1:n_top]
  })
  return(top_diseases)
}

# Create dataframe for plotting
library(tidyr)
library(dplyr)
library(ggplot2)

# Assuming phi is your K x D x T array
K <- dim(phi)[1]  # number of topics
D <- dim(phi)[2]  # number of diseases
T <- dim(phi)[3]  # number of timepoints

times <- 30:80  # assuming these are your age ranges

# Get top diseases
top_diseases <- get_top_diseases(phi, 10)

# Create plotting dataframe
phi_df <- do.call(rbind, lapply(1:K, function(k) {
  do.call(rbind, lapply(top_diseases[[k]], function(d) {
    data.frame(
      topic = factor(paste0("Topic ", k)),
      disease = factor(disease_names[d]),
      time = times,
      loading = phi[k,d,]
    )
  }))
}))

# Create faceted plot
p1 <- ggplot(phi_df, aes(x=time, y=loading, color=disease)) +
  geom_line(size=1) +
  facet_wrap(~topic, scales="free_y", ncol=2) +
  theme_minimal() +
  labs(title="Top Disease Trajectories by Topic",
       x="Age",
       y="Disease Loading") +
  theme(legend.position="right",
        plot.title = element_text(hjust=0.5, size=16),
        strip.text = element_text(size=12)) +
  scale_x_continuous(breaks=seq(30, 80, by=10))

print(p1)

Siimlarly, an individual’s relative enrichment in any one of these signatures at a given time

may change as influenced by genetics, environmental and previous diagnostic history. Characterizing a patient as loaded on ‘one signature’ belies the often complicated and beautiful trajectory of a. patient’s unique story.

which(rowSums(Y[,c(114,111,252),])==3&G$CAD>0&model_data$event_times[,111]<10)
# Create heatmap
mean_phi <- apply(phi, c(1,2), mean)
top_n_all <- unique(unlist(top_diseases))

heatmap_data <- reshape2::melt(mean_phi[,top_n_all])
names(heatmap_data) <- c("Topic", "Disease", "Loading")

p2 <- ggplot(heatmap_data, aes(x=Topic, y=Disease, fill=Loading)) +
  geom_tile() +
  scale_fill_gradient2(low="blue", mid="white", high="red", midpoint=0) +
  theme_minimal() +
  labs(title="Mean Disease Loadings by Topic",
       x="Topic",
       y="Disease") +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        plot.title = element_text(hjust=0.5, size=16))

print(p2)

```
```{r}
library(Matrix)
library(tidyverse)

# Store the actual IDs and ICD codes in order
person_ids <- sort(unique(icdlab$eid))
disease_codes <- sort(unique(icdlab$diag_icd10))

# Create mapping dictionaries with names preserved
id_map <- setNames(seq_len(length(person_ids)), person_ids)
disease_map <- setNames(seq_len(length(disease_codes)), disease_codes)

# Create sparse matrix with same logic as before
sparse_data <- icdlab %>%
  mutate(
    id_idx = id_map[as.character(eid)],
    disease_idx = disease_map[as.character(diag_icd10)],
    age_idx = age_diag - 29
  ) %>%
  filter(age_idx >= 1, age_idx <= 52)

# Create sparse array using sparseMatrix, now with dimnames
sparse_arrays <- vector("list", 52)
for(t in 1:52) {
  temp_data <- sparse_data %>% filter(age_idx == t)
  sparse_arrays[[t]] <- sparseMatrix(
    i = temp_data$id_idx,
    j = temp_data$disease_idx,
    x = 1,
    dims = c(length(id_map), length(disease_map)),
    dimnames = list(person_ids, disease_codes)  # Add names to dimensions
  )
}

# Save the dimension information
array_dims <- list(
  people = person_ids,
  diseases = disease_codes,
  ages = 30:81
)

# Now you can access by name or index:
# By index: sparse_arrays[[t]][i,j]
# By name: sparse_arrays[[t]]["1000015", "272"]
```

```{r}

phiatm=ATM::UKB_HES_10topics[1:81,c(1:348),]
colnames(phiatm)=c(disease_names_list)
rownames(phiatm)=1:81
l=melt(phiatm)
names(l)=c("Time","Disease","Signature","value")

# Step 1: Calculate the average value for ranking
top_diseases <- l %>%
group_by(Signature, Disease) %>%
summarize(avg_value = mean(value, na.rm = TRUE)) %>%
group_by(Signature) %>%
arrange(Signature, desc(avg_value)) %>%
slice_head(n = 20) %>%
ungroup()
# Step 2: Create a unique Disease-Signature combination for filtering
l_top <- l %>%
inner_join(top_diseases, by = c("Signature", "Disease")) %>%
merge(groups, by.x = "Disease", by.y = "phenotype", all.x = TRUE)
# Step 3: Create separate disease orders for each signature
l_top <- l_top %>%
group_by(Signature) %>%
mutate(Disease = factor(Disease,
levels = unique(Disease[order(avg_value,decreasing = F)]))) %>%
ungroup()
# Step 4: Plot with free_y scales to allow different diseases per facet
p2 <- ggplot(l_top, aes(Time, Disease, fill = value)) +
geom_tile() +
facet_wrap(~ Signature, scales = "free_y", drop = TRUE) +
labs(x = "Time (years from 30)", y = NULL,
title = "Top 20 diseases per signature over time",
fill = "Weight") +
scale_fill_viridis_c(option = "C", direction = +1) +
theme_minimal() +
theme(axis.text.y = element_text(size = 8),
strip.text = element_text(size = 10),
axis.ticks.y = element_blank(),
panel.spacing = unit(0.5, "lines"))

```