### 2025-09-24 pyScripts/embddingslides.qmd
update

---
title: "Embedding"
format: html
editor: visual
---

```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(viridis)
library(reshape2)
library(tibble)

set.seed(42)

# =============================================================================
# SLIDE 1: OVERVIEW
# =============================================================================
print_slide <- function(title, content) {
  cat("\n")
  cat(paste(rep("=", 80), collapse = ""), "\n")
  cat(paste0("SLIDE: ", title), "\n")
  cat(paste(rep("=", 80), collapse = ""), "\n")
  cat(content, "\n")
  cat(paste(rep("-", 80), collapse = ""), "\n")
}

print_slide("SIMPLIFIED HIERARCHICAL EMBEDDINGS EXPLAINED", 
"This simulation shows the SIMPLIFIED approach where:
1. Diseases and signatures get embeddings in shared space
2. Psi = raw attention scores (E_d^T * W * E_k)
3. NO contextualization, NO softmax, NO projection layers

Key insight: Psi directly measures embedding compatibility!")

# =============================================================================
# SLIDE 2: DATA GENERATION 
# =============================================================================
generate_simplified_embedding_data <- function(N = 1000, D = 50, K = 5, L = 10) {

  print_slide("STEP 1: CREATE TRUE EMBEDDINGS", 
  paste("Creating interpretable ground truth embeddings with realistic scale:
  - N =", N, "patients, D =", D, "diseases, K =", K, "signatures, L =", L, "dimensions
  - Each signature gets", floor(L/K), "dedicated dimensions
  - Block structure should be clearly visible"))

  # Step 1: Create interpretable signature embeddings
  # Step 1: Create interpretable signature embeddings with cycling assignment
E_k_true <- matrix(0, K, L)
for(k in 1:K) {
  # Primary dimension: cycle through available dimensions
  primary_dim <- (k-1) %% L + 1
  E_k_true[k, primary_dim] <- 2.0

  # Add 1-2 secondary dimensions for richness
  secondary_dims <- setdiff(1:L, primary_dim)
  if(length(secondary_dims) > 0) {
    n_secondary <- min(2, length(secondary_dims))
    selected_secondary <- sample(secondary_dims, n_secondary)
    E_k_true[k, selected_secondary] <- 0.3
  }

  cat(sprintf("Signature %d: primary dim %d, secondary dims %s\n", 
              k, primary_dim, paste(selected_secondary, collapse = ", ")))
}
  print("Signature embeddings structure:")
  print(round(E_k_true, 2))

  # Step 2: Create disease embeddings based on signature assignments
  E_d_true <- matrix(0, D, L)
  disease_assignments <- rep(1:K, length.out = D)

  cat(sprintf("Disease assignment pattern (first 20): %s\n", 
              paste(disease_assignments[1:20], collapse = ", ")))
  cat(sprintf("Diseases per signature: %d\n", D %/% K))

  for(d in 1:D) {
    primary_sig <- disease_assignments[d]

    # Strong loading on primary signature dimensions
    E_d_true[d, ] <- E_k_true[primary_sig, ] * (0.8 + rnorm(1, 0, 0.1))

    # Add some secondary loadings for realism
    secondary_sigs <- setdiff(1:K, primary_sig)
    if(length(secondary_sigs) > 0) {
      sec_sig <- sample(secondary_sigs, 1)
      E_d_true[d, ] <- E_d_true[d, ] + 0.2 * E_k_true[sec_sig, ]
    }  # FIXED: Added missing closing brace

    # Add noise
    E_d_true[d, ] <- E_d_true[d, ] + rnorm(L, 0, 0.05)
  }

  print_slide("STEP 2: CREATE ATTENTION MATRIX W", 
  "The attention matrix W determines how embedding dimensions interact:
  - Mostly diagonal (1.5) to preserve structure
  - Within-signature-block connections (0.3) for realism
  - This is the KEY learnable component that transforms embeddings")

  # Step 3: Create attention matrix W with better structure
  W_true <- matrix(0, L, L)
  diag(W_true) <- 1.5  # Strong diagonal to preserve structure

  # Add within-signature-block connections
  for(k in 1:K) {
    start_dim <- (k-1) * (L %/% K) + 1
    end_dim <- min(k * (L %/% K), L)

    if(end_dim > start_dim) {
      dims <- start_dim:end_dim
      for(i in dims) {
        for(j in dims) {
          if(i != j) {
            W_true[i, j] <- 0.3
          }
        }
      }
      cat(sprintf("Added within-block connections for signature %d (dims %d:%d)\n", 
                  k, start_dim, end_dim))
    }
  }

  print("Attention matrix W structure (showing non-zero elements):")
  W_display <- W_true
  W_display[W_display == 0] <- NA
  print(round(W_display, 2))

  print_slide("STEP 3: COMPUTE PSI DIRECTLY", 
  "SIMPLIFIED APPROACH:
  psi[k,d] = E_d[d]^T * W * E_k[k] / sqrt(L)

  This is the RAW attention score - no softmax, no projection!
  Each psi value measures embedding compatibility after transformation W.")

  # Step 4: Compute psi DIRECTLY (this is the key simplification)
  psi_true <- (E_d_true %*% W_true %*% t(E_k_true)) / sqrt(L)
  psi_true <- t(psi_true)  # Make it [K, D]

  print("True psi matrix structure [K x D] (first 6 diseases, all signatures):")
  print(round(psi_true[, 1:min(6, D)], 3))

  # Show psi statistics by signature
  cat("\nPsi statistics by signature (should show block structure):\n")
  for(k in 1:K) {
    diseases_in_sig <- which(disease_assignments == k)
    psi_in_block <- psi_true[k, diseases_in_sig]
    psi_out_block <- psi_true[k, -diseases_in_sig]

    cat(sprintf("Signature %d: In-block mean=%.3f, Out-block mean=%.3f, Ratio=%.2f\n",
                k, mean(psi_in_block), mean(psi_out_block), 
                mean(psi_in_block) / mean(psi_out_block)))
  }

  print_slide("STEP 4: GENERATE OUTCOME DATA", 
  "Generate disease outcomes using the true psi values:
  P(disease d | person i) ~ sigmoid(sum_k theta[i,k] * psi[k,d])

  This creates realistic disease patterns based on the embedding structure.")

  # Step 5: Generate outcome data using these psi values
  theta <- matrix(0, N, K)
  for(i in 1:N) {
    # Each person has different signature loadings (Dirichlet distribution)
    theta[i, ] <- rdirichlet(1, rep(2, K))
  }

  Y <- array(0, dim = c(N, D, 1))
  for(i in 1:N) {
    for(d in 1:D) {
      linear_pred <- sum(theta[i, ] * psi_true[, d])
      prob <- 1 / (1 + exp(-linear_pred))
      Y[i, d, 1] <- rbinom(1, 1, prob * 0.05)  # Lower rate for realism
    }
  }

  cat(sprintf("Generated %d total events across %d patients and %d diseases\n", 
              sum(Y), N, D))
  cat(sprintf("Event rate: %.4f (target was ~0.05)\n", mean(Y)))

  # Show disease-specific event rates
  disease_rates <- apply(Y, 2, mean)
  cat(sprintf("Disease event rates: Min=%.4f, Max=%.4f, Mean=%.4f\n", 
              min(disease_rates), max(disease_rates), mean(disease_rates)))

  return(list(
    E_d_true = E_d_true,
    E_k_true = E_k_true, 
    W_true = W_true,
    psi_true = psi_true,
    Y = Y,
    theta = theta,
    disease_assignments = disease_assignments,
    N = N, D = D, K = K, L = L
  ))
}

# Dirichlet distribution sampler
rdirichlet <- function(n, alpha) {
  k <- length(alpha)
  x <- matrix(rgamma(n * k, alpha, 1), n, k)
  x / rowSums(x)
}

# =============================================================================
# SLIDE 3: VISUALIZATION FUNCTIONS
# =============================================================================
create_comprehensive_plots <- function(data) {

  print_slide("CREATING VISUALIZATIONS", 
  "Now we'll visualize all components to see if they make sense:
  1. Embeddings in PCA space (diseases should cluster by signature)
  2. Attention matrix W structure (should be block-diagonal)
  3. Final psi values (should show clear block patterns)
  4. Raw embedding values (should show signature-specific patterns)")

  plots <- list()

  # 1. Embeddings in PCA space with better visualization
  E_d_pca <- prcomp(data$E_d_true)
  disease_df <- data.frame(
    PC1 = E_d_pca$x[,1],
    PC2 = E_d_pca$x[,2],
    Disease = paste0("D", 1:data$D),
    TrueCluster = factor(data$disease_assignments),
    Type = "Disease"
  )

  E_k_pca <- prcomp(data$E_k_true)
  signature_df <- data.frame(
    PC1 = E_k_pca$x[,1],
    PC2 = E_k_pca$x[,2],
    Disease = paste0("Sig", 1:data$K),
    TrueCluster = factor(1:data$K),
    Type = "Signature"
  )

  embedding_df <- rbind(
    disease_df[,c("PC1", "PC2", "Disease", "TrueCluster", "Type")],
    signature_df[,c("PC1", "PC2", "Disease", "TrueCluster", "Type")]
  )

  plots$p1 <- ggplot(embedding_df, aes(PC1, PC2, color = TrueCluster, shape = Type)) +
    geom_point(size = 2, alpha = 0.7) +
    scale_color_viridis_d(name = "Signature") +
    scale_shape_manual(values = c("Disease" = 16, "Signature" = 17)) +
    theme_minimal() +
    labs(title = "Disease & Signature Embeddings (PCA)",
         subtitle = paste("Diseases should cluster by signature assignment"),
         x = "First Principal Component", y = "Second Principal Component") +
    theme(legend.position = "bottom")

  # 2. Attention matrix W with better visualization
  W_df <- melt(data$W_true)
  colnames(W_df) <- c("Row", "Col", "Weight")

  plots$p2 <- ggplot(W_df, aes(Col, Row, fill = Weight)) +
    geom_tile() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                        midpoint = 0, name = "Weight") +
    theme_minimal() +
    labs(title = "Attention Matrix W",
         subtitle = "Block-diagonal structure preserves embedding dimensions",
         x = "Column", y = "Row")

  # 3. Psi values with clearer structure
  psi_df <- melt(data$psi_true)
  colnames(psi_df) <- c("Signature", "Disease", "Psi")

  plots$p3 <- ggplot(psi_df, aes(Disease, Signature, fill = Psi)) +
    geom_tile() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                        midpoint = 0, name = "Psi Value") +
    theme_minimal() +
    labs(title = "Psi Parameters (Disease-Signature Associations)",
         subtitle = "Block structure shows which diseases belong to which signatures",
         x = "Disease", y = "Signature") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

  # 4. Raw embedding values showing structure
  E_d_subset <- data$E_d_true[1:min(20, data$D), ]  # Show first 20 diseases
  E_k_subset <- data$E_k_true

  E_d_df <- melt(E_d_subset)
  colnames(E_d_df) <- c("Entity", "Dimension", "Value")
  E_d_df$Type <- "Disease"
  E_d_df$ID <- paste0("D", E_d_df$Entity)

  E_k_df <- melt(E_k_subset)
  colnames(E_k_df) <- c("Entity", "Dimension", "Value")
  E_k_df$Type <- "Signature"  
  E_k_df$ID <- paste0("S", E_k_df$Entity)

  embedding_values_df <- rbind(
    E_d_df[,c("Entity", "Dimension", "Value", "Type", "ID")],
    E_k_df[,c("Entity", "Dimension", "Value", "Type", "ID")]
  )

  plots$p4 <- ggplot(embedding_values_df, aes(Dimension, ID, fill = Value)) +
    geom_tile() +
    facet_wrap(~Type, scales = "free_y") +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                        midpoint = 0, name = "Value") +
    theme_minimal() +
    labs(title = "Raw Embedding Values",
         subtitle = "Block structure: each signature dominates specific dimensions",
         x = "Embedding Dimension", y = "Entity") +
    theme(axis.text.y = element_text(size = 8))

  # 5. Enhanced model summary with more metrics
  expected_psi <- matrix(-1, data$K, data$D)  # Background negative
  for(d in 1:data$D) {
    primary_sig <- data$disease_assignments[d]
    expected_psi[primary_sig, d] <- 1  # In-block positive
  }

  psi_correlation <- cor(as.vector(expected_psi), as.vector(data$psi_true))

  # Calculate additional metrics
  psi_block_ratio <- 0
  for(k in 1:data$K) {
    diseases_in_block <- which(data$disease_assignments == k)
    in_block_mean <- mean(data$psi_true[k, diseases_in_block])
    out_block_mean <- mean(data$psi_true[k, -diseases_in_block])
    psi_block_ratio <- psi_block_ratio + (in_block_mean / out_block_mean)
  }
  psi_block_ratio <- psi_block_ratio / data$K

  plots$p5 <- ggplot() +
    annotate("text", x = 0.5, y = 0.8, 
             label = paste("Block Structure Correlation:", round(psi_correlation, 3)), size = 5) +
    annotate("text", x = 0.5, y = 0.7, 
             label = paste("In-block/Out-block Ratio:", round(psi_block_ratio, 2)), size = 5) +
    annotate("text", x = 0.5, y = 0.6, 
             label = paste("Total Events:", sum(data$Y)), size = 4) +
    annotate("text", x = 0.5, y = 0.5, 
             label = paste("Event Rate:", round(mean(data$Y), 4)), size = 4) +
    annotate("text", x = 0.5, y = 0.4, 
             label = paste("Diseases per Signature:", data$D %/% data$K), size = 4) +
    annotate("text", x = 0.5, y = 0.3, 
             label = paste("Dimensions per Signature:", data$L %/% data$K), size = 4) +
    xlim(0, 1) + ylim(0, 1) +
    theme_void() +
    labs(title = "Model Performance Summary",
         subtitle = "Key metrics showing interpretability and structure")

  # 6. Parameter comparison
  direct_params <- data$K * data$D
  hierarchical_params <- data$D * data$L + data$K * data$L + data$L^2

  comparison_df <- data.frame(
    Method = c("Direct Psi", "Hierarchical"),
    Parameters = c(direct_params, hierarchical_params),
    Efficiency = c(1, hierarchical_params / direct_params)
  )

  plots$p6 <- ggplot(comparison_df, aes(Method, Parameters, fill = Method)) +
    geom_col() +
    geom_text(aes(label = Parameters), vjust = -0.5, size = 4) +
    scale_fill_viridis_d() +
    theme_minimal() +
    labs(title = "Parameter Efficiency",
         subtitle = paste("Ratio:", round(hierarchical_params/direct_params, 2), 
                         "x (", ifelse(hierarchical_params < direct_params, "EFFICIENT", "INEFFICIENT"), ")"),
         y = "Number of Parameters") +
    theme(legend.position = "none")

  return(plots)
}

# =============================================================================
# SLIDE 4: ENHANCED VALIDATION
# =============================================================================
validate_simplified_model <- function(data) {

  print_slide("VALIDATION: COMPREHENSIVE MODEL TESTING", 
  "Testing key properties with improved diagnostics:
  1. Perfect reconstruction (mathematical verification)
  2. Noise robustness (stability test)
  3. Block structure preservation (interpretability measure)
  4. Parameter efficiency (practical benefits)")

  # Test 1: Perfect reconstruction
  psi_reconstructed <- (data$E_d_true %*% data$W_true %*% t(data$E_k_true)) / sqrt(data$L)
  psi_reconstructed <- t(psi_reconstructed)

  reconstruction_correlation <- cor(as.vector(data$psi_true), as.vector(psi_reconstructed))
  reconstruction_mse <- mean((data$psi_true - psi_reconstructed)^2)

  cat(sprintf("Perfect reconstruction: correlation = %.6f, MSE = %.2e\n", 
              reconstruction_correlation, reconstruction_mse))

  # Test 2: Noise robustness with multiple noise levels
  noise_levels <- c(0.05, 0.1, 0.2)
  noise_correlations <- numeric(length(noise_levels))

  for(i in seq_along(noise_levels)) {
    E_d_noisy <- data$E_d_true + matrix(rnorm(data$D * data$L, 0, noise_levels[i]), 
                                       data$D, data$L)
    psi_noisy <- (E_d_noisy %*% data$W_true %*% t(data$E_k_true)) / sqrt(data$L)
    psi_noisy <- t(psi_noisy)
    noise_correlations[i] <- cor(as.vector(data$psi_true), as.vector(psi_noisy))

    cat(sprintf("Noise robustness (σ=%.2f): correlation = %.3f\n", 
                noise_levels[i], noise_correlations[i]))
  }

  # Test 3: Enhanced block structure analysis
  expected_pattern <- matrix(-1, data$K, data$D)
  for(d in 1:data$D) {
    primary_sig <- data$disease_assignments[d]
    expected_pattern[primary_sig, d] <- 1
  }

  pattern_correlation <- cor(as.vector(expected_pattern), as.vector(data$psi_true))

  # Additional block structure metrics
  block_separation_scores <- numeric(data$K)
  for(k in 1:data$K) {
    in_block_diseases <- which(data$disease_assignments == k)
    out_block_diseases <- setdiff(1:data$D, in_block_diseases)

    in_block_psi <- data$psi_true[k, in_block_diseases]
    out_block_psi <- data$psi_true[k, out_block_diseases]

    # t-test separation
    if(length(in_block_psi) > 1 && length(out_block_psi) > 1) {
      t_result <- t.test(in_block_psi, out_block_psi)
      block_separation_scores[k] <- abs(t_result$statistic)
    }

    cat(sprintf("Signature %d: In-block mean=%.3f (n=%d), Out-block mean=%.3f (n=%d), t=%.2f\n",
                k, mean(in_block_psi), length(in_block_psi), 
                mean(out_block_psi), length(out_block_psi),
                block_separation_scores[k]))
  }

  cat(sprintf("Overall block structure: correlation=%.3f, mean t-statistic=%.2f\n",
              pattern_correlation, mean(block_separation_scores, na.rm = TRUE)))

  # Test 4: Parameter efficiency analysis
  direct_params <- data$K * data$D
  hierarchical_params <- data$D * data$L + data$K * data$L + data$L^2
  efficiency_ratio <- hierarchical_params / direct_params

  # Calculate the breakeven point
  # Hierarchical < Direct when: D*L + K*L + L^2 < K*D
  # Solving for L: L^2 + L*(D+K) < K*D
  # For practical purposes: L < K*D/(D+K) approximately
  breakeven_L <- (data$K * data$D) / (data$D + data$K)

  cat(sprintf("\nParameter Analysis:\n"))
  cat(sprintf("  Direct parameters: %d\n", direct_params))
  cat(sprintf("  Hierarchical parameters: %d\n", hierarchical_params))
  cat(sprintf("  Efficiency ratio: %.2f\n", efficiency_ratio))
  cat(sprintf("  Breakeven embedding dimension: %.1f (current: %d)\n", 
              breakeven_L, data$L))

  success_criteria <- list(
    reconstruction = reconstruction_correlation > 0.999,
    noise_robustness = min(noise_correlations) > 0.7,
    block_structure = pattern_correlation > 0.3,
    parameter_efficiency = efficiency_ratio < 2.0
  )

  overall_success <- all(unlist(success_criteria))

  cat(sprintf("\nSUCCESS CRITERIA:\n"))
  for(criterion in names(success_criteria)) {
    status <- ifelse(success_criteria[[criterion]], "PASS", "FAIL")
    cat(sprintf("  %s: %s\n", criterion, status))
  }

  return(list(
    reconstruction_correlation = reconstruction_correlation,
    noise_correlations = noise_correlations,
    pattern_correlation = pattern_correlation,
    block_separation_scores = block_separation_scores,
    efficiency_ratio = efficiency_ratio,
    breakeven_L = breakeven_L,
    overall_success = overall_success
  ))
}

# =============================================================================
# SLIDE 5: MAIN SIMULATION
# =============================================================================
run_complete_simplified_simulation <- function() {

  print_slide("HIERARCHICAL EMBEDDINGS: COMPLETE DEMONSTRATION", 
  "This simulation demonstrates the simplified hierarchical embedding approach:

  CORE INSIGHT: psi[k,d] = E_d[d]^T * W * E_k[k] / sqrt(L)

  We'll use realistic but manageable scale to clearly show the benefits.")

  # Generate data with realistic scale
  data <- generate_simplified_embedding_data(N = 1000, D = 50, K = 10, L = 5)

  # Create comprehensive visualizations
  plots <- create_comprehensive_plots(data)

  # Validate the approach thoroughly
  validation <- validate_simplified_model(data)

  print_slide("INTERPRETATION AND NEXT STEPS", 
  "The simplified hierarchical embedding approach provides:

  1. SEMANTIC STRUCTURE: Diseases in embedding space reflect biological similarity
  2. PARAMETER EFFICIENCY: Fewer parameters when L is chosen appropriately  
  3. INTERPRETABILITY: Attention weights show disease-signature relationships
  4. EXTENSIBILITY: Easy to add medication effects, new diseases, etc.

  For your real problem (D=350, K=20), choose L=5-8 for optimal efficiency.")

  # Create final visualization
  final_plot <- grid.arrange(
    plots$p1, plots$p2, plots$p5,
    plots$p3, plots$p4, plots$p6,
    ncol = 3, nrow = 2,
    top = "Simplified Hierarchical Embeddings: Complete Analysis"
  )

  # Comprehensive summary
  cat("\n")
  cat("=", "\n")
  cat("FINAL SIMULATION RESULTS\n")
  cat("=", "\n")
  cat(sprintf("Reconstruction: %.6f (perfect=1.0)\n", validation$reconstruction_correlation))
  cat(sprintf("Noise robustness: %.3f (robust>0.8)\n", min(validation$noise_correlations))) 
  cat(sprintf("Block structure: %.3f (interpretable>0.3)\n", validation$pattern_correlation))
  cat(sprintf("Parameter efficiency: %.2fx (efficient<1.0)\n", validation$efficiency_ratio))
  cat(sprintf("Optimal L for your data: ≤%.0f (current: %d)\n", validation$breakeven_L, data$L))

  if(validation$overall_success) {
    cat("\n✓ SUCCESS: The simplified hierarchical embedding approach works!\n")
    cat("\nFor your real application:\n")
    cat("- Use L=5 with D=350, K=20 for maximum efficiency\n") 
    cat("- Expect ~73% parameter reduction vs direct psi learning\n")
    cat("- Maintain interpretable disease-signature relationships\n")
  } else {
    cat("\n✗ Issues detected - review the diagnostics above\n")
  }

  return(list(data = data, plots = plots, validation = validation))
}

# Run the complete simulation
results <- run_complete_simplified_simulation()
```

The `echo: false` option disables the printing of code (only output is displayed).