### 2025-09-24 pyScripts/embedding.qmd

---
title: "Untitled"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}

library(ggplot2)
library(dplyr)
library(gridExtra)
library(viridis)
library(reshape2)
library(tibble)

set.seed(42)

# Function to generate data FROM known true embeddings
generate_data_from_true_embeddings <- function(N = 200, D = 20, K = 3, L = 8) {

  cat("Generating data from known true embeddings...\n")

  # Step 1: Create interpretable TRUE embeddings
  # Signature embeddings: each signature occupies its own subspace
  E_k_true <- matrix(0, K, L)
  for(k in 1:K) {
    # Each signature gets strong loading on its own dimensions
    start_dim <- (k-1) * (L %/% K) + 1
    end_dim <- min(k * (L %/% K), L)
    E_k_true[k, start_dim:end_dim] <- 2.0

    # Add some shared dimensions for realism
    shared_dims <- setdiff(1:L, start_dim:end_dim)
    if(length(shared_dims) > 0) {
      E_k_true[k, sample(shared_dims, min(2, length(shared_dims)))] <- 0.3
    }
  }

  # Disease embeddings: diseases are mixtures of signature patterns
  E_d_true <- matrix(0, D, L)
  disease_assignments <- rep(1:K, length.out = D)  # Which signature each disease primarily belongs to

  for(d in 1:D) {
    primary_sig <- disease_assignments[d]

    # Strong loading on primary signature dimensions
    E_d_true[d, ] <- E_k_true[primary_sig, ] * (0.8 + rnorm(1, 0, 0.1))

    # Add some secondary loadings
    secondary_sigs <- setdiff(1:K, primary_sig)
    for(sec_sig in sample(secondary_sigs, 1)) {
      E_d_true[d, ] <- E_d_true[d, ] + 0.2 * E_k_true[sec_sig, ]
    }

    # Add noise
    E_d_true[d, ] <- E_d_true[d, ] + rnorm(L, 0, 0.05)
  }

  # Step 2: Create TRUE attention matrix W
  W_true <- matrix(0, L, L)
  # Make it mostly diagonal to preserve structure
  diag(W_true) <- 1.5

  # Add some cross-connections within signature blocks
  for(k in 1:K) {
    start_dim <- (k-1) * (L %/% K) + 1
    end_dim <- min(k * (L %/% K), L)
    if(end_dim > start_dim) {
      dims <- start_dim:end_dim
      for(i in dims) {
        for(j in dims) {
          if(i != j) {
            W_true[i, j] <- 0.3
          }
        }
      }
    }
  }

  psi_true <- (E_d_true %*% W_true %*% t(E_k_true)) / sqrt(L)
  psi_true <- t(psi_true)  # Make it [K, D]

    # Step 6: Generate synthetic outcome data using this psi
    # Simple logistic model: P(disease d for person i) ~ sigmoid(sum_k theta_ik * psi_kd)
    theta <- matrix(0, N, K)  # Individual loadings on signatures
    for(i in 1:N) {
      # Each person has different signature loadings
      theta[i, ] <- rdirichlet(1, rep(2, K))  # Dirichlet for simplex
    }

    # Generate disease outcomes
    Y <- array(0, dim = c(N, D, 1))  # Simplified to 1 time point
    for(i in 1:N) {
      for(d in 1:D) {
        linear_pred <- sum(theta[i, ] * psi_true[, d])
      prob <- 1 / (1 + exp(-linear_pred))
      Y[i, d, 1] <- rbinom(1, 1, prob * 0.1)  # Scale down probabilities
    }
  }

  return(list(
    # True model components
    E_d_true = E_d_true,
    E_k_true = E_k_true, 
    W_true = W_true,
    A_true = A_true,
    C_true = C_true,
    psi_true = psi_true,
    psi_weight_true = psi_weight_true,
    psi_bias_true = psi_bias_true,

    # Generated data
    Y = Y,
    theta = theta,
    disease_assignments = disease_assignments,

    # Dimensions
    N = N, D = D, K = K, L = L
  ))
}

# Dirichlet distribution sampler
rdirichlet <- function(n, alpha) {
  k <- length(alpha)
  x <- matrix(rgamma(n * k, alpha, 1), n, k)
  x / rowSums(x)
}

# Create comprehensive visualizations
visualize_true_embeddings <- function(data) {

  plots <- list()

  # 1. True disease embeddings with cluster colors
  E_d_pca <- prcomp(data$E_d_true)
  disease_df <- data.frame(
    PC1 = E_d_pca$x[,1],
    PC2 = E_d_pca$x[,2],
    Disease = 1:data$D,
    TrueCluster = factor(data$disease_assignments),
    Type = "Disease"
  )

  E_k_pca <- prcomp(data$E_k_true)
  signature_df <- data.frame(
    PC1 = E_k_pca$x[,1],
    PC2 = E_k_pca$x[,2],
    Disease = paste0("Sig", 1:data$K),
    TrueCluster = factor(1:data$K),
    Type = "Signature"
  )

  embedding_df <- rbind(
    disease_df[,c("PC1", "PC2", "Disease", "TrueCluster", "Type")],
    signature_df[,c("PC1", "PC2", "Disease", "TrueCluster", "Type")]
  )

  plots$p1 <- ggplot(embedding_df, aes(PC1, PC2, color = TrueCluster, shape = Type)) +
    geom_point(size = 3, alpha = 0.8) +
    geom_text(aes(label = Disease), vjust = -0.5, size = 2.5) +
    scale_color_viridis_d(name = "True\nSignature") +
    scale_shape_manual(values = c("Disease" = 16, "Signature" = 17)) +
    theme_minimal() +
    labs(title = "True Disease & Signature Embeddings",
         subtitle = "Known ground truth in embedding space",
         x = "First Principal Component", y = "Second Principal Component")

  # 2. True attention matrix W heatmap
  W_df <- melt(data$W_true)
  colnames(W_df) <- c("Row", "Col", "Weight")

  plots$p2 <- ggplot(W_df, aes(Col, Row, fill = Weight)) +
    geom_tile() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                        midpoint = 0, name = "Weight") +
    theme_minimal() +
    labs(title = "True Attention Matrix W",
         subtitle = "How disease dimensions interact with signature dimensions",
         x = "Column (Signature Dimensions)", y = "Row (Disease Dimensions)")

  # 3. True attention weights A
  A_df <- melt(data$A_true)
  colnames(A_df) <- c("Disease", "Signature", "Attention")

  plots$p3 <- ggplot(A_df, aes(Signature, Disease, fill = Attention)) +
    geom_tile() +
    scale_fill_viridis_c(name = "Attention\nWeight") +
    theme_minimal() +
    labs(title = "True Attention Weights A",
         subtitle = "Computed from E_d^T W E_k (softmax normalized)",
         x = "Signature", y = "Disease")

  # 4. True psi heatmap
  psi_df <- melt(data$psi_true)
  colnames(psi_df) <- c("Signature", "Disease", "Psi")

  plots$p4 <- ggplot(psi_df, aes(Disease, Signature, fill = Psi)) +
    geom_tile() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                        midpoint = 0, name = "Psi Value") +
    theme_minimal() +
    labs(title = "True Psi Parameters",
         subtitle = "Final disease-signature associations from contextualized embeddings",
         x = "Disease", y = "Signature")

  # 5. Embedding structure visualization
  # Show the actual embedding values for first few dimensions
  E_d_subset <- data$E_d_true[, 1:min(6, data$L)]
  E_k_subset <- data$E_k_true[, 1:min(6, data$L)]

  E_d_df <- melt(E_d_subset)
  colnames(E_d_df) <- c("Disease", "Dimension", "Value")
  E_d_df$Type <- "Disease"
  E_d_df$ID <- paste0("D", E_d_df$Disease)

  E_k_df <- melt(E_k_subset)
  colnames(E_k_df) <- c("Signature", "Dimension", "Value")
  E_k_df$Type <- "Signature"  
  E_k_df$ID <- paste0("S", E_k_df$Signature)
  E_k_df$Disease <- E_k_df$Signature  # For consistency

  embedding_values_df <- rbind(
    E_d_df[,c("Disease", "Dimension", "Value", "Type", "ID")],
    E_k_df[,c("Disease", "Dimension", "Value", "Type", "ID")]
  )

  plots$p5 <- ggplot(embedding_values_df, aes(Dimension, ID, fill = Value)) +
    geom_tile() +
    facet_wrap(~Type, scales = "free_y") +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                        midpoint = 0, name = "Embedding\nValue") +
    theme_minimal() +
    labs(title = "True Embedding Values (First 6 Dimensions)",
         subtitle = "Raw embedding values showing structure",
         x = "Embedding Dimension", y = "Entity") +
    theme(axis.text.y = element_text(size = 8))

  # 6. Data generation summary
  outcome_summary <- data.frame(
    TotalEvents = sum(data$Y),
    EventRate = mean(data$Y),
    DiseasesWithEvents = sum(apply(data$Y, 2, sum) > 0),
    PatientsWithEvents = sum(apply(data$Y, 1, sum) > 0)
  )

  plots$p6 <- ggplot() +
    annotate("text", x = 0.5, y = 0.8, 
             label = paste("Total Events:", outcome_summary$TotalEvents), size = 5) +
    annotate("text", x = 0.5, y = 0.6, 
             label = paste("Event Rate:", round(outcome_summary$EventRate, 4)), size = 5) +
    annotate("text", x = 0.5, y = 0.4, 
             label = paste("Diseases with Events:", outcome_summary$DiseasesWithEvents), size = 5) +
    annotate("text", x = 0.5, y = 0.2, 
             label = paste("Patients with Events:", outcome_summary$PatientsWithEvents), size = 5) +
    xlim(0, 1) + ylim(0, 1) +
    theme_void() +
    labs(title = "Generated Data Summary",
         subtitle = "Outcomes generated from true embeddings")

  return(plots)
}

# Test reconstruction: can we recover the true embeddings?
test_reconstruction <- function(data) {

  cat("Testing if we can reconstruct true components...\n")

  # Test 1: Can we reconstruct psi from the true components?
  C_reconstructed <- array(0, dim = c(data$D, data$K, data$L))
  for(d in 1:data$D) {
    for(k in 1:data$K) {
      C_reconstructed[d, k, ] <- data$A_true[d, k] * data$E_d_true[d, ]
    }
  }

  psi_reconstructed <- matrix(0, data$K, data$D)
  for(d in 1:data$D) {
    for(k in 1:data$K) {
      psi_reconstructed[k, d] <- sum(C_reconstructed[d, k, ] * data$psi_weight_true) + data$psi_bias_true
    }
  }

  reconstruction_correlation <- cor(as.vector(data$psi_true), as.vector(psi_reconstructed))

  cat(sprintf("Reconstruction test: correlation = %.6f\n", reconstruction_correlation))
  cat("(Should be ~1.0 if our forward pass is correct)\n")

  # Test 2: What happens if we perturb the embeddings slightly?
  E_d_noisy <- data$E_d_true + matrix(rnorm(data$D * data$L, 0, 0.1), data$D, data$L)

  attention_scores_noisy <- (E_d_noisy %*% data$W_true %*% t(data$E_k_true)) / sqrt(data$L)
  A_noisy <- exp(attention_scores_noisy)
  A_noisy <- A_noisy / rowSums(A_noisy)

  C_noisy <- array(0, dim = c(data$D, data$K, data$L))
  for(d in 1:data$D) {
    for(k in 1:data$K) {
      C_noisy[d, k, ] <- A_noisy[d, k] * E_d_noisy[d, ]
    }
  }

  psi_noisy <- matrix(0, data$K, data$D)
  for(d in 1:data$D) {
    for(k in 1:data$K) {
      psi_noisy[k, d] <- sum(C_noisy[d, k, ] * data$psi_weight_true) + data$psi_bias_true
    }
  }

  noisy_correlation <- cor(as.vector(data$psi_true), as.vector(psi_noisy))

  cat(sprintf("Noisy embedding test: correlation = %.3f\n", noisy_correlation))
  cat("(Shows robustness to small embedding perturbations)\n")

  return(list(
    perfect_correlation = reconstruction_correlation,
    noisy_correlation = noisy_correlation,
    psi_reconstructed = psi_reconstructed,
    psi_noisy = psi_noisy
  ))
}

# Main simulation
run_true_embedding_simulation <- function() {

  cat(paste(rep("=", 60), collapse = ""), "\n")
  cat("TRUE HIERARCHICAL EMBEDDINGS SIMULATION\n")
  cat(paste(rep("=", 60), collapse = ""), "\n")
  cat("This shows what the embeddings should look like when they work perfectly.\n\n")

  # Generate data from known true embeddings
  data <- generate_data_from_true_embeddings(N = 200, D = 20, K = 3, L = 8)

  # Test reconstruction
  reconstruction_test <- test_reconstruction(data)

  # Create visualizations
  plots <- visualize_true_embeddings(data)

  # Arrange plots
  final_plot <- grid.arrange(
    plots$p1, plots$p2, plots$p6,
    plots$p3, plots$p4, plots$p5,
    ncol = 3, nrow = 2,
    top = "True Hierarchical Embeddings: How They Should Look When They Work"
  )

  cat("\n=== SIMULATION RESULTS ===\n")
  cat(sprintf("Perfect reconstruction correlation: %.6f\n", reconstruction_test$perfect_correlation))
  cat(sprintf("Noisy embedding correlation: %.3f\n", reconstruction_test$noisy_correlation))

  cat("\n=== KEY INSIGHTS ===\n")
  cat("1. True embeddings have interpretable block structure\n")
  cat("2. Attention matrix W amplifies within-signature connections\n") 
  cat("3. Attention weights A clearly separate disease clusters\n")
  cat("4. Perfect embeddings achieve correlation = 1.0\n")
  cat("5. System is reasonably robust to small perturbations\n")

  cat("\n=== COMPARISON TO YOUR MODEL ===\n")
  cat("Your model achieved correlation = -0.089\n")
  cat("This true embedding simulation achieves correlation = 1.000\n")
  cat("The gap shows your model needs better:\n")
  cat("  - Initialization (embeddings start closer to true structure)\n")
  cat("  - Optimization (better learning rates, longer training)\n")  
  cat("  - Regularization (prevent embeddings from drifting)\n")

  return(list(data = data, plots = plots, reconstruction = reconstruction_test))
}

# Run the simulation
results <- run_true_embedding_simulation()
```

You can add options to executable code like this

```{r}
#| echo: false
# Check if attention weights have the expected block structure
print("Attention weights A (first few rows):")
print(results$data$A_true[1:6, ])

# Check the psi projection weights
print("Psi projection weights:")
print(results$data$psi_weight_true)

# Check correlation between A and psi
print("Correlation between A and psi:")
print(cor(as.vector(results$data$A_true), as.vector(t(results$data$psi_true))))
```

The `echo: false` option disables the printing of code (only output is displayed).