### 2025-10-18 presntations/heart_house_roundtable_lifecourse.qmd
update

---
title: "Cardiovascular Health and Disease Across the Life Course"
subtitle: "Many Roads to Rome"
author: "Sarah Urbut, MD PhD"
institute: "Massachusetts General Hospital"
date: today
format:
  revealjs:

    slide-number: true

    preview-links: auto
    footer: "Heart House Roundtable | Session 2"
    transition: slide
    background-transition: fade
    highlight-style: github
    code-fold: true
    fig-cap-location: bottom
    embed-resources: true
    reference-location: document
    smaller: false
    scrollable: true
editor: visual
---

```{python}
#| include: false
import numpy as np
import matplotlib.pyplot as plt
import pandas as pd
import seaborn as sns
import torch

# Set plotting style
plt.style.use('seaborn-v0_8-darkgrid')
sns.set_palette("husl")
```

# The Life Course Paradigm Shift {background-color="#2C3E50"}

## Current Risk Models: The 10-Year Problem

::::: columns
::: {.column width="50%"}
**Traditional Approach:**

-   Pooled Cohort Equations (PCE)
-   10-year ASCVD risk
-   Treatment at thresholds (≥7.5%)

**The Question:** \> "Who will have an event in the next 10 years?"
:::

::: {.column width="50%"}
**What's Missing:**

-   Young patients with high lifetime risk
-   Low short-term but high long-term risk
-   Different disease trajectories
-   Optimal intervention timing

**Better Question:** \> "What is their lifetime burden, and when should we intervene?"
:::
:::::

::: notes
For patients in their 40s and 50s—the critical decades—we're asking the wrong question. A 45-year-old with high LDL might have low 10-year risk but massive lifetime burden.
:::

## Many Roads to Rome

```{python}
#| echo: false
#| fig-width: 12
#| fig-height: 6

# Create synthetic trajectory data showing different paths to same endpoint
np.random.seed(42)
ages = np.arange(30, 80)

# Four different trajectories
fig, ax = plt.subplots(figsize=(12, 6))

# Trajectory 1: Early genetic risk (steep early rise)
genetic_risk = 0.1 + 0.02 * (ages - 30)**1.5 + np.random.normal(0, 0.05, len(ages))
genetic_risk = np.clip(genetic_risk, 0, 1)

# Trajectory 2: Metabolic (gradual middle age)
metabolic = 0.05 + 0.015 * (ages - 40)**1.2
metabolic[ages < 40] = 0.05
metabolic = np.clip(metabolic + np.random.normal(0, 0.03, len(ages)), 0, 1)

# Trajectory 3: Inflammatory (late onset, sharp rise)
inflammatory = np.zeros(len(ages))
inflammatory[ages > 55] = 0.02 * (ages[ages > 55] - 55)**1.8
inflammatory = np.clip(inflammatory + np.random.normal(0, 0.02, len(ages)), 0, 1)

# Trajectory 4: Healthy ager (stays low)
healthy = 0.02 + 0.001 * (ages - 30) + np.random.normal(0, 0.01, len(ages))
healthy = np.clip(healthy, 0, 0.15)

ax.plot(ages, genetic_risk, 'r-', linewidth=3, label='High Genetic Risk', alpha=0.8)
ax.plot(ages, metabolic, 'b-', linewidth=3, label='Metabolic Trajectory', alpha=0.8)
ax.plot(ages, inflammatory, 'g-', linewidth=3, label='Inflammatory Pathway', alpha=0.8)
ax.plot(ages, healthy, 'purple', linewidth=3, label='Healthy Ager', alpha=0.8)

# Mark critical intervention points
ax.axvspan(40, 60, alpha=0.1, color='orange', label='Critical Decades')

ax.set_xlabel('Age (years)', fontsize=14, fontweight='bold')
ax.set_ylabel('Cumulative Disease Risk', fontsize=14, fontweight='bold')
ax.set_title('Multiple Trajectories to Cardiovascular Disease', fontsize=16, fontweight='bold')
ax.legend(fontsize=12, loc='upper left')
ax.grid(True, alpha=0.3)
ax.set_ylim([0, 1])

plt.tight_layout()
plt.show()
```

::: callout-important
## Key Insight

Traditional risk models assume one pathway. Reality is far more complex—different trajectories require different intervention strategies.
:::

# CV Health Personas {background-color="#34495E"}

## Persona 1: High Genetic Risk

::::: columns
::: {.column width="60%"}
**Clinical Profile:**

-   High polygenic risk score (PRS)
-   Or familial hypercholesterolemia (FH)
-   LDL 180 mg/dL at age 35
-   CAD onset by age 50

**Research Evidence:**

-   *Ference et al., NEJM 2012*: Genetic variants that lower cholesterol even moderately over lifetime dramatically reduce CAD
-   *Khera et al., Nat Genet 2018*: PRS identifies high-risk individuals missed by traditional screening
:::

::: {.column width="40%"}
**Your Model Findings:**

```{python}
#| echo: false
#| fig-width: 5
#| fig-height: 4

# Simplified signature visualization
ages_short = np.arange(30, 82)
sig_genetic = 0.1 + 0.015 * (ages_short - 30) + np.random.normal(0, 0.02, len(ages_short))
sig_genetic = np.clip(sig_genetic, 0, 0.8)

fig, ax = plt.subplots(figsize=(5, 4))
ax.plot(ages_short, sig_genetic, 'r-', linewidth=3)
ax.axvline(45, color='orange', linestyle='--', linewidth=2, label='Signature rise')
ax.axvline(50, color='red', linestyle='--', linewidth=2, label='Clinical CAD')
ax.set_xlabel('Age', fontsize=11)
ax.set_ylabel('CVD Signature Loading', fontsize=11)
ax.set_title('Signature Changes Precede Diagnosis\nby 5-10 years', fontsize=12, fontweight='bold')
ax.legend(fontsize=9)
ax.grid(True, alpha=0.3)
plt.tight_layout()
plt.show()
```

**Signatures change 5-10 years before clinical diagnosis**
:::
:::::

::: notes
This is the "catch them early" persona. Current guidelines treat FH aggressively from childhood, but individuals with high PRS can have similar risk and are often missed.
:::

## Persona 2: Metabolic Trajectory

::::: columns
::: {.column width="50%"}
**Progression Path:**

1.  Prediabetes (age 40-45)
2.  Type 2 diabetes (age 45-50)
3.  Metabolic syndrome (age 50-55)
4.  CAD (age 55-60)

**Key Research Finding:**

*Urbut et al., medRxiv 2024*: \> "Metabolic signature (Sig 12) rises first, cardiovascular signature (Sig 5) follows 2-3 years later"

**The Intervention Window:** - When metabolic signature plateaus - But CVD signature accelerates - 2-3 years before clinical CAD
:::

::: {.column width="50%"}
**Signature Transition:**

```{python}
#| echo: false
#| fig-width: 6
#| fig-height: 5

# Metabolic to CVD signature transition
ages_met = np.arange(40, 70)

# Metabolic signature rises then plateaus
sig_metabolic = 0.2 + 0.025 * (ages_met - 40)
sig_metabolic[ages_met > 55] = 0.2 + 0.025 * 15  # Plateau
sig_metabolic = np.clip(sig_metabolic + np.random.normal(0, 0.02, len(ages_met)), 0, 1)

# CVD signature rises later
sig_cvd = np.zeros(len(ages_met))
sig_cvd[ages_met > 52] = 0.02 * (ages_met[ages_met > 52] - 52)**1.5
sig_cvd = np.clip(sig_cvd + np.random.normal(0, 0.015, len(ages_met)), 0, 1)

fig, ax = plt.subplots(figsize=(6, 5))
ax.plot(ages_met, sig_metabolic, 'b-', linewidth=3, label='Metabolic Signature (Sig 12)', alpha=0.8)
ax.plot(ages_met, sig_cvd, 'r-', linewidth=3, label='CVD Signature (Sig 5)', alpha=0.8)

# Mark transition window
transition_start = 53
transition_end = 57
ax.axvspan(transition_start, transition_end, alpha=0.2, color='orange', label='Transition Window')
ax.axvline(60, color='red', linestyle='--', linewidth=2, alpha=0.5, label='Clinical CAD')

ax.set_xlabel('Age (years)', fontsize=12, fontweight='bold')
ax.set_ylabel('Signature Loading (θ)', fontsize=12, fontweight='bold')
ax.set_title('Metabolic → Cardiovascular Transition\n(Intervention Window: 2-3 years before CAD)',
             fontsize=13, fontweight='bold')
ax.legend(fontsize=10, loc='upper left')
ax.grid(True, alpha=0.3)
ax.set_ylim([0, 0.8])

plt.tight_layout()
plt.show()
```
:::
:::::

## Persona 3: Inflammatory Pathway

::::: columns
::: {.column width="55%"}
**Clinical Picture:**

-   Rheumatoid arthritis, psoriasis, or lupus
-   Often normal lipids
-   "Low" 10-year PCE risk
-   But ongoing systemic inflammation

**Your Research:**

-   52 patients: rheumatologic → CV transitions
-   Signature deviations **5 years pre-CAD**
-   Multi-signature interactions:
    -   Metabolic + Inflammatory together
    -   Synergistic effect (9.1% \> 4.3% + 2.8%)

**Evidence Base:**

-   *ACC Statement on Inflammation*: Inflammation is therapeutic target in CVD
-   CANTOS trial: IL-1β inhibition reduces CV events
:::

::: {.column width="45%"}
**Signature Behavior:**

```{python}
#| echo: false
#| fig-width: 5
#| fig-height: 4.5

# Inflammatory pathway
ages_inflam = np.arange(45, 75)
sig_inflam = 0.3 + 0.01 * (ages_inflam - 45) + np.random.normal(0, 0.02, len(ages_inflam))
sig_cv_inflam = np.zeros(len(ages_inflam))
sig_cv_inflam[ages_inflam > 60] = 0.03 * (ages_inflam[ages_inflam > 60] - 60)**1.3
sig_cv_inflam = np.clip(sig_cv_inflam + np.random.normal(0, 0.01, len(ages_inflam)), 0, 1)

fig, ax = plt.subplots(figsize=(5, 4.5))
ax.plot(ages_inflam, sig_inflam, 'g-', linewidth=3, label='Inflammatory Sig', alpha=0.8)
ax.plot(ages_inflam, sig_cv_inflam, 'r-', linewidth=3, label='CVD Sig', alpha=0.8)
ax.axvline(70, color='red', linestyle='--', linewidth=2, alpha=0.5, label='Clinical CAD')
ax.fill_between(ages_inflam, 0, 1, where=(ages_inflam >= 65) & (ages_inflam <= 70),
                alpha=0.15, color='orange', label='Pre-CAD deviation')

ax.set_xlabel('Age (years)', fontsize=11)
ax.set_ylabel('Signature Loading', fontsize=11)
ax.set_title('Inflammatory → CVD Pathway\n(Often Missed by Traditional Models)',
             fontsize=12, fontweight='bold')
ax.legend(fontsize=9)
ax.grid(True, alpha=0.3)
ax.set_ylim([0, 0.7])

plt.tight_layout()
plt.show()
```

**Often missed:** Normal lipids but high inflammatory burden
:::
:::::

## Persona 4: Healthy Ager

::::: columns
::: {.column width="50%"}
**Characteristics:**

-   Low signature loadings across decades
-   Maintained cardiovascular health
-   Protective factors?
    -   Genetics
    -   Lifestyle
    -   Both?

**Model Performance:**

-   Baseline AUC: **0.898** (ASCVD)
-   1-year washout: **0.701**
-   2-year washout: **0.680**

**Key Question:** What distinguishes this group? Can we replicate their trajectory in others?
:::

::: {.column width="50%"}
```{python}
#| echo: false
#| fig-width: 6
#| fig-height: 5

# Healthy ager comparison
ages_full = np.arange(30, 80)

# Healthy agers
healthy_sig = 0.05 + 0.002 * (ages_full - 30) + np.random.normal(0, 0.01, len(ages_full))
healthy_sig = np.clip(healthy_sig, 0, 0.15)

# Average population
avg_sig = 0.05 + 0.01 * (ages_full - 30)**1.1 + np.random.normal(0, 0.03, len(ages_full))
avg_sig = np.clip(avg_sig, 0, 0.6)

# High risk
high_sig = 0.1 + 0.02 * (ages_full - 30)**1.3 + np.random.normal(0, 0.04, len(ages_full))
high_sig = np.clip(high_sig, 0, 1)

fig, ax = plt.subplots(figsize=(6, 5))
ax.fill_between(ages_full, 0, high_sig, alpha=0.3, color='red', label='High Risk Trajectories')
ax.fill_between(ages_full, 0, avg_sig, alpha=0.4, color='blue', label='Average Population')
ax.plot(ages_full, healthy_sig, 'purple', linewidth=4, label='Healthy Agers', alpha=0.9)

ax.set_xlabel('Age (years)', fontsize=12, fontweight='bold')
ax.set_ylabel('CVD Signature Loading', fontsize=12, fontweight='bold')
ax.set_title('Healthy Agers: Maintained Low Risk Across Life Course',
             fontsize=13, fontweight='bold')
ax.legend(fontsize=11, loc='upper left')
ax.grid(True, alpha=0.3)
ax.set_ylim([0, 1])

plt.tight_layout()
plt.show()
```
:::
:::::

::: callout-note
## Clinical Impact

Same 10-year risk ≠ Same intervention. Trajectory determines timing and intensity of prevention.
:::

# Dynamic Risk: Life Course Models {background-color="#2C3E50"}

## The Aladynoulli Model

:::::: columns
::: {.column width="50%"}
**What It Does:**

-   Longitudinal disease signatures over 30+ years
-   Bayesian survival framework
-   Dynamic risk updates

**What It Incorporates:**

1.  **Genetics** (fixed): Polygenic risk scores
2.  **Biomarkers** (dynamic): Serial lipids, BP, glucose
3.  **Imaging** (when available): CAC, plaque burden
4.  **Events** (transitions): New diagnoses as signals

**Scale:**

-   348 diseases
-   400,000 individuals (UK Biobank)
-   21 disease signatures
-   50+ time points (ages 30-80)
:::

:::: {.column width="50%"}
**Mathematical Framework:**

$$\pi_{i,d,t} = \kappa \times \sum_k \theta_{i,k,t} \times \phi_{k,d,t}$$

Where:

-   $\theta_{i,k,t}$ = signature proportions (from Gaussian process)
-   $\phi_{k,d,t}$ = disease probabilities
-   $\lambda_{i,k,t} \sim GP(\mu_k + G_i\gamma_k, K_\lambda)$ (temporal dynamics)

**Key Innovation:**

Signature **velocity** (rate of change) predicts risk:

::: callout-important
"Top quartile Sig 5 velocity has **HR=2.8** for CAD"
:::

This translates to **18 months earlier CAD onset**
::::
::::::

## Novel Biomarker: Signature Velocity

::::: columns
::: {.column width="50%"}
**Traditional Approach:**

-   Single time point measurement
-   Example: LDL = 150 mg/dL

**Life Course Approach:**

-   **Trajectory** and **velocity**
-   Example: LDL 120→180 over 10 years
-   Rate of change = 6 mg/dL/year

**Your Discovery:**

> "Fast signature progression → higher disease risk independent of absolute level"

**Clinical Translation:**

-   Top quartile velocity: HR = 2.8 for CAD
-   18 months earlier onset
-   New therapeutic target: Slow progression
:::

::: {.column width="50%"}
```{python}
#| echo: false
#| fig-width: 6
#| fig-height: 6

# Velocity analysis
ages_vel = np.arange(40, 70)

# Slow velocity
slow_vel = 0.1 + 0.005 * (ages_vel - 40) + np.random.normal(0, 0.01, len(ages_vel))
slow_vel = np.clip(slow_vel, 0, 1)

# Medium velocity
med_vel = 0.1 + 0.012 * (ages_vel - 40) + np.random.normal(0, 0.015, len(ages_vel))
med_vel = np.clip(med_vel, 0, 1)

# Fast velocity
fast_vel = 0.1 + 0.025 * (ages_vel - 40) + np.random.normal(0, 0.02, len(ages_vel))
fast_vel = np.clip(fast_vel, 0, 1)

fig, (ax1, ax2) = plt.subplots(2, 1, figsize=(6, 6))

# Top: Trajectories
ax1.plot(ages_vel, slow_vel, 'g-', linewidth=3, label='Slow Velocity (Q1)', alpha=0.8)
ax1.plot(ages_vel, med_vel, 'b-', linewidth=3, label='Medium Velocity (Q2-3)', alpha=0.8)
ax1.plot(ages_vel, fast_vel, 'r-', linewidth=3, label='Fast Velocity (Q4)', alpha=0.8)
ax1.set_ylabel('Signature Loading', fontsize=11, fontweight='bold')
ax1.set_title('Signature Trajectories by Velocity Quartile', fontsize=12, fontweight='bold')
ax1.legend(fontsize=9)
ax1.grid(True, alpha=0.3)

# Bottom: Hazard ratios
velocities = ['Q1\n(Slow)', 'Q2-3\n(Medium)', 'Q4\n(Fast)']
hrs = [1.0, 1.5, 2.8]
colors_hr = ['green', 'blue', 'red']

bars = ax2.bar(velocities, hrs, color=colors_hr, alpha=0.7, edgecolor='black', linewidth=2)
ax2.axhline(1, color='black', linestyle='--', linewidth=1.5, alpha=0.5)
ax2.set_ylabel('Hazard Ratio for CAD', fontsize=11, fontweight='bold')
ax2.set_xlabel('Velocity Quartile', fontsize=11, fontweight='bold')
ax2.set_title('Velocity Predicts CAD Risk (HR = 2.8 for Q4)', fontsize=12, fontweight='bold')
ax2.set_ylim([0, 3.5])
ax2.grid(True, alpha=0.3, axis='y')

# Add value labels
for i, (bar, hr) in enumerate(zip(bars, hrs)):
    height = bar.get_height()
    ax2.text(bar.get_x() + bar.get_width()/2., height + 0.1,
            f'HR={hr}', ha='center', va='bottom', fontweight='bold', fontsize=10)

plt.tight_layout()
plt.show()
```
:::
:::::

## Example: Same Risk, Different Trajectories

::: panel-tabset
### Case Comparison

**Two 55-year-old women, both with 10-year ASCVD risk of 7.5%:**

| Feature | Woman A | Woman B |
|------------------------|------------------------|------------------------|
| **LDL trajectory** | Stable 130 mg/dL × 20 years | 120→180 mg/dL over 15 years |
| **Inflammation** | Recent hsCRP uptick (1→4 mg/L) | Stable CRP \< 2 mg/L |
| **Signature pattern** | Inflammatory sig rising | Metabolic sig dominant |
| **PCE risk** | 7.5% | 7.5% |

### Traditional Model

**Both receive identical recommendations:**

-   Same 10-year risk (7.5%)
-   Same statin recommendation
-   Same LDL target (\<100 mg/dL)
-   No differentiation

### Life Course Model

**Different strategies based on trajectories:**

**Woman A (Inflammatory):**

-   Earlier intervention (age 50 vs 55)
-   Inflammatory biomarker monitoring
-   Consider imaging (CAC score)
-   Possibly anti-inflammatory Rx

**Woman B (Metabolic):**

-   Focus on metabolic control
-   Intensive lifestyle intervention
-   Early GLP-1/SGLT2i consideration
-   Tighter glucose monitoring
:::

# From Theory to Practice {background-color="#34495E"}

## Precision Prevention by Persona

::: panel-tabset
### High Genetic Risk

**Current Guidelines:**

-   LDL threshold ≥190 mg/dL (FH)
-   Age-independent statin for FH

**Life Course Approach:**

-   Genetic testing age 30-40
-   PRS-guided intervention
-   Earlier statin initiation
-   More aggressive LDL targets

**Evidence:**

-   AHA Statement: PRS highly predictive in younger populations
-   Ference et al.: Lifetime cholesterol lowering \> short-term intensive

**Implementation:**

-   Cost: \$50-200 for PRS
-   Coverage: Expanding but limited
-   Clinical workflow: Needs EHR integration

### Metabolic Trajectory

**Current Approach:**

-   Wait for diabetes (A1c ≥6.5%)
-   Reactive treatment

**Life Course Approach:**

-   Intervene at **transition point**
-   When metabolic sig plateaus
-   Before CVD sig accelerates

**Your Data:**

-   Transition occurs 2-3 years pre-CAD
-   Window for intensive prevention

**Interventions:**

-   Intensive lifestyle (DPP)
-   Early SGLT2i/GLP1-RA
-   Closer monitoring (3-6 months)

### Inflammatory Pathway

**Current Gap:**

-   Normal lipids = missed
-   Low PCE despite inflammation

**Life Course Approach:**

-   Inflammatory biomarkers (hsCRP, IL-6)
-   Imaging (CAC) for risk refinement
-   Earlier intervention in rheumatologic disease

**Your Finding:**

-   Signature deviations 5 years pre-CAD
-   52 patients with rheum→CV transitions

**Treatment Considerations:**

-   Optimize rheumatologic disease control
-   Consider colchicine
-   More aggressive lipid lowering
-   Serial imaging

### Healthy Agers

**Learn from Success:**

-   What protects this group?
-   Can we replicate their trajectory?

**Research Needed:**

-   Genomic studies
-   Lifestyle patterns
-   Biomarker profiles
-   Microbiome?

**Clinical Application:**

-   Risk stratification
-   Targeted prevention
-   Personalized goals
:::

## Implementation Challenges

:::::: columns
::: {.column width="33%"}
**Clinical Barriers:**

-   EHR integration
    -   No life course calculators
    -   Data fragmentation
    -   Manual entry burden
-   Decision support
    -   Need real-time risk updates
    -   Trajectory visualization
    -   Actionable recommendations
-   Workflow integration
    -   Time constraints
    -   Training needed
    -   Change management
:::

::: {.column width="33%"}
**Payer Barriers:**

-   Coverage policies
    -   "Preventive" Rx in "healthy" patients
    -   Genetic testing reimbursement
    -   Advanced imaging (CAC, CCTA)
-   Value demonstration
    -   Long-term outcomes
    -   Cost-effectiveness
    -   ROI uncertainty
-   Prior authorization
    -   Treatment thresholds
    -   Documentation burden
:::

::: {.column width="34%"}
**Communication Challenges:**

-   Patient understanding
    -   Lifetime vs 10-year risk
    -   Abstract concepts
    -   Motivation for prevention
-   Shared decision-making
    -   Risk/benefit discussions
    -   Treatment intensity
    -   Monitoring frequency
-   Tools needed
    -   Visual aids
    -   Personalized reports
    -   Decision aids
:::
::::::

::: callout-warning
## Key Question for Discussion

What's the threshold for payer coverage of preventive therapies in otherwise "healthy" high-lifetime-risk patients?
:::

# Looking Forward {background-color="#2C3E50"}

## Tools on the Horizon

::::: columns
::: {.column width="50%"}
**Current State:**

**LIVE-CVD Model:**

-   First lifelong benefit calculator
-   Integrates traditional risk factors
-   Projects lifetime ASCVD risk

**SCOT-HEART Trial:**

-   5-year: CT improves risk stratification
-   10-year: Sustained benefit
-   SCOT-HEART 2: Ongoing

**Your Model (Aladynoulli):**

-   Bayesian framework
-   Genetics + biomarkers + events
-   Dynamic, continuous updates
-   348 diseases, 21 signatures
:::

::: {.column width="50%"}
**Emerging Tools:**

**AI-Enhanced Imaging:**

-   Automated plaque quantification
-   Pericoronary fat analysis
-   ECG-based AI risk scores
-   Retinal imaging

**Polygenic Risk Scores:**

-   Million-variant PRS
-   Multi-ancestry validation
-   Clinical implementation starting

**EHR Integration:**

-   Real-time calculators
-   Passive monitoring
-   Alert systems
-   Population health dashboards

**Wearables & Sensors:**

-   Continuous BP monitoring
-   Glucose sensors (CGM)
-   Activity tracking
-   Sleep quality
:::
:::::

## The Promise: Precision Prevention

```{python}
#| echo: false
#| fig-width: 12
#| fig-height: 6

# Comparison of approaches
fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(12, 6))

# Left: Traditional (one-size-fits-all)
ages_trad = np.arange(30, 80)
trad_threshold = np.ones(len(ages_trad)) * 0.075  # 7.5% threshold

# All patients treated the same above threshold
patient1_trad = 0.05 + 0.008 * (ages_trad - 30) + np.random.normal(0, 0.01, len(ages_trad))
patient2_trad = 0.03 + 0.01 * (ages_trad - 30) + np.random.normal(0, 0.012, len(ages_trad))
patient3_trad = 0.04 + 0.009 * (ages_trad - 30) + np.random.normal(0, 0.011, len(ages_trad))

ax1.plot(ages_trad, patient1_trad, 'r-', linewidth=2, alpha=0.7, label='Patient 1')
ax1.plot(ages_trad, patient2_trad, 'b-', linewidth=2, alpha=0.7, label='Patient 2')
ax1.plot(ages_trad, patient3_trad, 'g-', linewidth=2, alpha=0.7, label='Patient 3')
ax1.axhline(0.075, color='orange', linestyle='--', linewidth=3, label='Treatment Threshold (7.5%)')
ax1.fill_between(ages_trad, 0.075, 1, alpha=0.15, color='orange')

ax1.set_xlabel('Age (years)', fontsize=12, fontweight='bold')
ax1.set_ylabel('10-Year ASCVD Risk', fontsize=12, fontweight='bold')
ax1.set_title('Traditional: One-Size-Fits-All\n(Same threshold for all)',
              fontsize=13, fontweight='bold')
ax1.legend(fontsize=10)
ax1.grid(True, alpha=0.3)
ax1.set_ylim([0, 0.6])

# Right: Life course (personalized)
# Different intervention points based on trajectory
patient1_lc = 0.1 + 0.02 * (ages_trad - 30)**1.3 + np.random.normal(0, 0.015, len(ages_trad))
patient2_lc = 0.05 + 0.015 * (ages_trad - 40)**1.1
patient2_lc[ages_trad < 40] = 0.05
patient3_lc = 0.02 + 0.003 * (ages_trad - 30) + np.random.normal(0, 0.008, len(ages_trad))

patient1_lc = np.clip(patient1_lc, 0, 1)
patient2_lc = np.clip(patient2_lc, 0, 1)
patient3_lc = np.clip(patient3_lc, 0, 0.3)

ax2.plot(ages_trad, patient1_lc, 'r-', linewidth=2, alpha=0.7, label='High Genetic (early Rx)')
ax2.plot(ages_trad, patient2_lc, 'b-', linewidth=2, alpha=0.7, label='Metabolic (mid Rx)')
ax2.plot(ages_trad, patient3_lc, 'g-', linewidth=2, alpha=0.7, label='Healthy Ager (late/no Rx)')

# Mark different intervention points
ax2.axvline(40, color='red', linestyle=':', linewidth=2, alpha=0.5)
ax2.axvline(50, color='blue', linestyle=':', linewidth=2, alpha=0.5)
ax2.text(40, 0.55, 'Genetic\nIntervention', ha='center', fontsize=9, color='red', fontweight='bold')
ax2.text(50, 0.55, 'Metabolic\nIntervention', ha='center', fontsize=9, color='blue', fontweight='bold')

ax2.set_xlabel('Age (years)', fontsize=12, fontweight='bold')
ax2.set_ylabel('Lifetime Disease Burden', fontsize=12, fontweight='bold')
ax2.set_title('Life Course: Precision Prevention\n(Trajectory-specific timing)',
              fontsize=13, fontweight='bold')
ax2.legend(fontsize=10, loc='upper left')
ax2.grid(True, alpha=0.3)
ax2.set_ylim([0, 0.6])

plt.tight_layout()
plt.show()
```

::: callout-tip
## The Goal

Tailor prevention to individual life course trajectories—not one-size-fits-all based on 10-year risk alone.
:::

## Key Discussion Questions

::: incremental
1.  **How do we reconcile population-level risk models with individualized prediction?**
    -   Especially when individuals may benefit from early intervention despite low calculated risk
2.  **Can life-course risk prediction be practically implemented?**
    -   What barriers do you foresee to payers covering preventive therapies in "healthy" patients?
3.  **How do we bridge lifetime risk reduction with current risk management?**
    -   Lower thresholds? Earlier age to start therapy? Public health efforts?
4.  **Communication strategies:**
    -   How do we make genetics, AI, and imaging meaningful to patient care?
    -   What tools help patients understand their lifetime trajectory?
5.  **Testing strategies:**
    -   Should everyone get CAC or CCTA every 10 years starting at age 40?
    -   Who should get genetic testing? When?
:::

# Summary {background-color="#2C3E50"}

## Key Takeaways

::::: columns
::: {.column width="50%"}
**The Problem:**

-   Traditional models miss young high-lifetime-risk patients
-   One-size-fits-all thresholds
-   No consideration of trajectory

**The Solution:**

-   Life course risk assessment
-   Multiple pathways to disease
-   Dynamic, updating models
-   Precision prevention

**The Evidence:**

-   Multiple trajectories exist (personas)
-   Signatures change years before diagnosis
-   Velocity predicts risk (HR=2.8)
-   Synergistic multi-signature effects
:::

::: {.column width="50%"}
**The Tools:**

-   Bayesian modeling (Aladynoulli)
-   Polygenic risk scores
-   AI-enhanced imaging
-   EHR integration coming

**The Challenges:**

-   Clinical workflow integration
-   Payer coverage
-   Patient communication
-   Training needed

**The Path Forward:**

-   Validate tools in diverse populations
-   Demonstrate cost-effectiveness
-   Build clinical decision support
-   Engage all stakeholders
:::
:::::

## Thank You

::::: columns
::: {.column width="50%"}
**Contact:**

Sarah Urbut, MD PhD Massachusetts General Hospital [surbut\@mgh.harvard.edu](mailto:surbut@mgh.harvard.edu)

**Preprint:**

Urbut et al., medRxiv 2024 doi: 10.1101/2024.09.29.24314557

**Code:**

[github.com/surbut/aladynoulli2](https://github.com/surbut/aladynoulli2)
:::

::: {.column width="50%"}
**Key Collaborators:**

-   UK Biobank participants
-   Mass General Brigham
-   PRIME-Heart Initiative

**Questions for Discussion:**

1.  Which personas resonate with your practice?
2.  What's your biggest barrier to life course prevention?
3.  Which tools are you most excited about?
4.  What would change your practice tomorrow?
:::
:::::

::: callout-note
## Ready for Discussion

Let's explore how we can move from theory to practice in the critical decades.
:::

# Backup Slides {background-color="#34495E"}

## Model Details: Aladynoulli

**Mathematical Framework:**

Disease probability at time $t$: $$\pi_{i,d,t} = \kappa \times \sum_{k=1}^K \theta_{i,k,t} \times \phi_{k,d,t}$$

**Components:**

-   $\theta_{i,k,t} = \text{softmax}(\lambda_{i,k,t})$ — signature proportions
-   $\lambda_{i,k,t} \sim GP(\mu_k + G_i\gamma_k, K_\lambda)$ — temporal dynamics with genetics
-   $\phi_{k,d,t} = \text{sigmoid}(\psi_{k,d} + GP(\mu_\phi, K_\phi))$ — disease probabilities

**Innovation:**

-   Gaussian processes for smooth temporal evolution
-   Genetic effects on signature trajectories
-   Proper censoring handling
-   Bayesian uncertainty quantification

## Validation Results

**UK Biobank (Primary):**

-   N = 400,000 individuals
-   348 diseases
-   21 signatures discovered
-   AUC = 0.898 (ASCVD, baseline)
-   AUC = 0.701 (1-year washout)

**Cross-Cohort Validation:**

-   Mass General Brigham
-   All of Us Research Program
-   79% signature concordance

**Robustness:**

-   Stable across ancestry groups
-   Stable across socioeconomic strata (TDI)
-   Population prevalence matches ONS/NHS

## Citations & References

::: smaller
1.  Goff et al. 2013 ACC/AHA Pooled Cohort Equations. *Circulation* 2014.
2.  Lloyd-Jones et al. Prediction of Lifetime Risk for CVD. *Circulation* 2006.
3.  Khera et al. Polygenic Prediction of Weight and Obesity Trajectories. *Cell* 2019.
4.  Grundy et al. 2018 AHA/ACC Cholesterol Guidelines. *Circulation* 2019.
5.  Ference et al. Effect of Long-Term Exposure to Lower LDL. *NEJM* 2012.
6.  AHA Scientific Statement on Polygenic Risk Scores, 2022.
7.  Ridker et al. Antiinflammatory Therapy with Canakinumab (CANTOS). *NEJM* 2017.
8.  Harshfield et al. Association Between Genetic Variation at ACE and Risk of COVID-19. *Circulation* 2020.
9.  Newby et al. Coronary CT Angiography and 5-Year Risk (SCOT-HEART). *NEJM* 2018.
10. Williams et al. Coronary Artery Plaque Characteristics (SCOT-HEART 10-year). *NEJM* 2024.
11. Urbut et al. Aladynoulli: Bayesian Survival Model for Disease Trajectories. *medRxiv* 2024.
:::