### 2025-10-19 presntations/heart_house_lifecourse_simple.qmd
Update

---
title: "Cardiovascular Health Across the Life Course"
subtitle: "Many Roads to Rome"
author: "Sarah Urbut, MD PhD"
institute: "Massachusetts General Hospital"
date: today
format:
  revealjs:
    slide-number: true
    preview-links: auto
    footer: "Heart House Roundtable | Session 2"
    transition: slide
    embed-resources: true
    smaller: false
editor: visual
---

```{python}
#| include: false
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns

np.random.seed(42)
plt.style.use('seaborn-v0_8-darkgrid')
sns.set_palette("husl")
```

# The Life Course Paradigm

## Current Risk Models: The 10-Year Problem

::: columns
::: {.column width="50%"}
**Traditional Approach:**

-   10-year ASCVD risk (PCE)
-   Treatment at thresholds (≥7.5%)
-   Static, single time point

**The Question:**

> "Who will have an event in the next 10 years?"
:::

::: {.column width="50%"}
**What's Missing:**

-   Young patients with high lifetime risk
-   Different disease trajectories
-   Optimal intervention timing
-   Information that accrues over time

**Better Question:**

> "What is their lifetime burden, and when should we intervene?"
:::
:::

## Many Roads to Rome

```{python}
#| echo: false
#| fig-width: 11
#| fig-height: 5.5

ages = np.arange(30, 80)

# Four trajectories
genetic = 0.1 + 0.02 * (ages - 30)**1.5 + np.random.normal(0, 0.05, len(ages))
genetic = np.clip(genetic, 0, 1)

metabolic = np.where(ages < 40, 0.05, 0.05 + 0.015 * (ages - 40)**1.2)
metabolic = np.clip(metabolic + np.random.normal(0, 0.03, len(ages)), 0, 1)

inflammatory = np.where(ages > 55, 0.02 * (ages - 55)**1.8, 0)
inflammatory = np.clip(inflammatory + np.random.normal(0, 0.02, len(ages)), 0, 1)

healthy = 0.02 + 0.001 * (ages - 30) + np.random.normal(0, 0.01, len(ages))
healthy = np.clip(healthy, 0, 0.15)

fig, ax = plt.subplots(figsize=(11, 5.5))
ax.plot(ages, genetic, 'r-', linewidth=3, label='Early Genetic Risk', alpha=0.8)
ax.plot(ages, metabolic, 'b-', linewidth=3, label='Metabolic Trajectory', alpha=0.8)
ax.plot(ages, inflammatory, 'g-', linewidth=3, label='Inflammatory', alpha=0.8)
ax.plot(ages, healthy, 'purple', linewidth=3, label='Healthy Ager', alpha=0.8)

ax.axvspan(40, 60, alpha=0.1, color='orange')
ax.text(50, 0.92, 'Critical Decades', ha='center', fontsize=13, fontweight='bold')

ax.set_xlabel('Age (years)', fontsize=13, fontweight='bold')
ax.set_ylabel('Cumulative CVD Risk', fontsize=13, fontweight='bold')
ax.set_title('Multiple Trajectories to Cardiovascular Disease', fontsize=15, fontweight='bold')
ax.legend(fontsize=11, loc='upper left')
ax.grid(True, alpha=0.3)
ax.set_ylim([0, 1])
plt.tight_layout()
plt.show()
```

# Four Trajectories

## Trajectory 1: Early Genetic Risk

::: columns
::: {.column width="55%"}
**Profile:**

-   High polygenic risk score (PRS) or FH
-   Elevated LDL from young age
-   CAD by age 45-50

**Key Evidence:**

-   *Ference, NEJM 2012*: Lifetime cholesterol lowering >> short-term intensive
-   *Khera, Nat Genet 2018*: PRS identifies missed high-risk individuals
-   *AHA 2022*: PRS highly predictive in younger populations

**The Concept:**

**Genetics provides early signal** → Lower threshold for intervention
:::

::: {.column width="45%"}
```{python}
#| echo: false
#| fig-width: 5
#| fig-height: 4.5

ages_short = np.arange(30, 75)

# Early vs late genetic risk
early_genetic = 0.15 + 0.018 * (ages_short - 30) + np.random.normal(0, 0.02, len(ages_short))
early_genetic = np.clip(early_genetic, 0, 1)

avg_risk = 0.05 + 0.008 * (ages_short - 30) + np.random.normal(0, 0.015, len(ages_short))
avg_risk = np.clip(avg_risk, 0, 0.6)

fig, ax = plt.subplots(figsize=(5, 4.5))
ax.fill_between(ages_short, 0, early_genetic, alpha=0.3, color='red', label='High PRS')
ax.fill_between(ages_short, 0, avg_risk, alpha=0.4, color='blue', label='Average PRS')

ax.axvline(40, color='red', linestyle='--', linewidth=2, alpha=0.6)
ax.text(41, 0.85, 'Early\nIntervention', fontsize=10, color='red', fontweight='bold')

ax.set_xlabel('Age', fontsize=11, fontweight='bold')
ax.set_ylabel('Cumulative Risk', fontsize=11, fontweight='bold')
ax.set_title('Genetics: Early Risk Signal', fontsize=12, fontweight='bold')
ax.legend(fontsize=10)
ax.grid(True, alpha=0.3)
plt.tight_layout()
plt.show()
```
:::
:::

## Trajectory 2: Metabolic Pathway

::: columns
::: {.column width="50%"}
**Progression:**

1.  Prediabetes (age 40-45)
2.  Type 2 diabetes (age 45-50)
3.  Metabolic syndrome (age 50-55)
4.  CAD (age 55-60)

**Key Evidence:**

-   DPP: Lifestyle intervention delays diabetes
-   Look AHEAD: Weight loss reduces CV events
-   EMPA-REG, LEADER: GLP-1/SGLT2i benefit

**The Concept:**

**Biomarkers accumulate** → Intervene at transition points
:::

::: {.column width="50%"}
```{python}
#| echo: false
#| fig-width: 5.5
#| fig-height: 5

ages_met = np.arange(35, 70)

# Metabolic markers increase
metabolic_markers = np.where(ages_met < 40, 0.05, 0.05 + 0.025 * (ages_met - 40))
metabolic_markers = np.clip(metabolic_markers + np.random.normal(0, 0.02, len(ages_met)), 0, 1)

fig, ax = plt.subplots(figsize=(5.5, 5))
ax.plot(ages_met, metabolic_markers, 'b-', linewidth=3, alpha=0.8)

# Mark key transitions
ax.axvline(45, color='orange', linestyle=':', linewidth=2)
ax.axvline(50, color='orange', linestyle=':', linewidth=2)
ax.text(47.5, 0.7, 'Diabetes\nOnset', ha='center', fontsize=9, color='orange', fontweight='bold')

ax.axvspan(50, 55, alpha=0.15, color='green')
ax.text(52.5, 0.85, 'Intervention\nWindow', ha='center', fontsize=9, fontweight='bold')

ax.set_xlabel('Age', fontsize=11, fontweight='bold')
ax.set_ylabel('Metabolic Risk', fontsize=11, fontweight='bold')
ax.set_title('Biomarkers Accumulate Over Time', fontsize=12, fontweight='bold')
ax.grid(True, alpha=0.3)
plt.tight_layout()
plt.show()
```
:::
:::

## Trajectory 3: Inflammatory Pathway

::: columns
::: {.column width="53%"}
**Profile:**

-   Rheumatologic disease (RA, psoriasis)
-   Often normal lipids
-   "Low" 10-year PCE risk
-   Chronic systemic inflammation

**Key Evidence:**

-   CANTOS: IL-1β inhibition reduces CV events
-   Colchicine trials: Anti-inflammatory benefit
-   ACC Statement: Inflammation as CV target

**The Concept:**

**Alternative risk pathway** → Missed by traditional models
:::

::: {.column width="47%"}
```{python}
#| echo: false
#| fig-width: 5
#| fig-height: 4.5

ages_inflam = np.arange(40, 75)

# Inflammation-driven risk
inflam_risk = 0.2 + 0.008 * (ages_inflam - 40) + np.random.normal(0, 0.02, len(ages_inflam))
inflam_risk = np.clip(inflam_risk, 0, 0.5)

traditional_risk = np.where(ages_inflam < 55, 0.03, 0.03 + 0.015 * (ages_inflam - 55))

fig, ax = plt.subplots(figsize=(5, 4.5))
ax.plot(ages_inflam, inflam_risk, 'g-', linewidth=3, label='Actual Risk\n(with inflammation)', alpha=0.8)
ax.plot(ages_inflam, traditional_risk, 'gray', linewidth=2, linestyle='--',
        label='PCE estimate\n(misses inflammation)', alpha=0.7)

ax.set_xlabel('Age', fontsize=11, fontweight='bold')
ax.set_ylabel('CVD Risk', fontsize=11, fontweight='bold')
ax.set_title('Traditional Models Miss\nInflammatory Risk', fontsize=12, fontweight='bold')
ax.legend(fontsize=9)
ax.grid(True, alpha=0.3)
plt.tight_layout()
plt.show()
```
:::
:::

## Trajectory 4: Healthy Ager

::: columns
::: {.column width="50%"}
**Characteristics:**

-   Maintained low risk across decades
-   Favorable genetics + lifestyle?
-   Resilience to aging

**Questions:**

-   What protects this group?
-   Can we replicate their trajectory?
-   Genomics? Lifestyle? Microbiome?

**The Concept:**

**Learn from success** → Identify protective factors
:::

::: {.column width="50%"}
```{python}
#| echo: false
#| fig-width: 5.5
#| fig-height: 5

ages_full = np.arange(30, 80)

healthy = 0.05 + 0.002 * (ages_full - 30) + np.random.normal(0, 0.01, len(ages_full))
healthy = np.clip(healthy, 0, 0.15)

average = 0.05 + 0.01 * (ages_full - 30)**1.1 + np.random.normal(0, 0.03, len(ages_full))
average = np.clip(average, 0, 0.6)

high = 0.1 + 0.02 * (ages_full - 30)**1.3 + np.random.normal(0, 0.04, len(ages_full))
high = np.clip(high, 0, 1)

fig, ax = plt.subplots(figsize=(5.5, 5))
ax.fill_between(ages_full, 0, high, alpha=0.25, color='red', label='High Risk')
ax.fill_between(ages_full, 0, average, alpha=0.35, color='blue', label='Average')
ax.plot(ages_full, healthy, 'purple', linewidth=4, label='Healthy Agers', alpha=0.9)

ax.set_xlabel('Age', fontsize=11, fontweight='bold')
ax.set_ylabel('CVD Risk', fontsize=11, fontweight='bold')
ax.set_title('Maintained Low Risk Across Life Course', fontsize=12, fontweight='bold')
ax.legend(fontsize=10, loc='upper left')
ax.grid(True, alpha=0.3)
ax.set_ylim([0, 1])
plt.tight_layout()
plt.show()
```
:::
:::

# The Life Course Approach

## Information Accrues Over Time

::: columns
::: {.column width="50%"}
**Starting Point (Age 30-40):**

-   **Genetics** (fixed)
    -   Polygenic risk scores
    -   Family history
    -   Monogenic conditions (FH)

**This provides the baseline trajectory**

**Over Time (Age 40-70):**

-   **Biomarkers** (dynamic)
    -   Serial lipids, BP, glucose
    -   Inflammatory markers
    -   Novel biomarkers

-   **Imaging** (periodic)
    -   CAC scoring
    -   Carotid ultrasound
    -   CT angiography

-   **Events** (sentinel)
    -   Diabetes diagnosis
    -   Hypertension
    -   Other comorbidities
:::

::: {.column width="50%"}
```{python}
#| echo: false
#| fig-width: 5.5
#| fig-height: 6

fig, ax = plt.subplots(figsize=(5.5, 6))

ages_info = np.arange(30, 75)

# Risk prediction uncertainty decreases over time
baseline_risk = 0.3 + 0.01 * (ages_info - 30)

# Wide uncertainty early (genetics only)
early_uncert = 0.25
mid_uncert = 0.15
late_uncert = 0.08

uncertainty = early_uncert - (early_uncert - late_uncert) * ((ages_info - 30) / (75 - 30))

upper = baseline_risk + uncertainty
lower = baseline_risk - uncertainty

# Plot narrowing uncertainty
ax.fill_between(ages_info, lower, upper, alpha=0.3, color='blue', label='Prediction Uncertainty')
ax.plot(ages_info, baseline_risk, 'b-', linewidth=3, label='Estimated Risk')

# Mark information addition points
ax.axvline(35, color='red', linestyle=':', linewidth=1.5, alpha=0.6)
ax.text(35, 0.85, 'Genetics', fontsize=8, rotation=90, va='bottom')

ax.axvline(45, color='orange', linestyle=':', linewidth=1.5, alpha=0.6)
ax.text(45, 0.85, 'Biomarkers', fontsize=8, rotation=90, va='bottom')

ax.axvline(55, color='green', linestyle=':', linewidth=1.5, alpha=0.6)
ax.text(55, 0.85, 'Imaging', fontsize=8, rotation=90, va='bottom')

ax.axvline(60, color='purple', linestyle=':', linewidth=1.5, alpha=0.6)
ax.text(60, 0.85, 'Events', fontsize=8, rotation=90, va='bottom')

ax.set_xlabel('Age', fontsize=12, fontweight='bold')
ax.set_ylabel('Risk Estimate', fontsize=12, fontweight='bold')
ax.set_title('Uncertainty Decreases as\nInformation Accumulates', fontsize=13, fontweight='bold')
ax.legend(fontsize=10, loc='upper left')
ax.grid(True, alpha=0.3)
ax.set_ylim([0, 1])

plt.tight_layout()
plt.show()
```

**Uncertainty decreases as information accumulates**
:::
:::

## Traditional vs. Life Course

```{python}
#| echo: false
#| fig-width: 12
#| fig-height: 5.5

fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(12, 5.5))

ages_comp = np.arange(30, 75)

# LEFT: Traditional - single threshold
patient1_trad = 0.04 + 0.009 * (ages_comp - 30) + np.random.normal(0, 0.01, len(ages_comp))
patient2_trad = 0.05 + 0.010 * (ages_comp - 30) + np.random.normal(0, 0.01, len(ages_comp))
patient3_trad = 0.03 + 0.008 * (ages_comp - 30) + np.random.normal(0, 0.01, len(ages_comp))

ax1.plot(ages_comp, patient1_trad, 'r-', linewidth=2.5, alpha=0.7, label='Patient A')
ax1.plot(ages_comp, patient2_trad, 'b-', linewidth=2.5, alpha=0.7, label='Patient B')
ax1.plot(ages_comp, patient3_trad, 'g-', linewidth=2.5, alpha=0.7, label='Patient C')
ax1.axhline(0.075, color='orange', linestyle='--', linewidth=3, label='Rx Threshold (7.5%)')
ax1.fill_between(ages_comp, 0.075, 0.6, alpha=0.12, color='orange')

ax1.set_xlabel('Age (years)', fontsize=12, fontweight='bold')
ax1.set_ylabel('10-Year ASCVD Risk', fontsize=12, fontweight='bold')
ax1.set_title('Traditional: One Threshold\n(Same Rx timing for all)', fontsize=13, fontweight='bold')
ax1.legend(fontsize=10)
ax1.grid(True, alpha=0.3)
ax1.set_ylim([0, 0.6])

# RIGHT: Life course - personalized timing
genetic_lc = 0.12 + 0.018 * (ages_comp - 30)**1.2 + np.random.normal(0, 0.015, len(ages_comp))
genetic_lc = np.clip(genetic_lc, 0, 1)

metabolic_lc = np.where(ages_comp < 40, 0.05, 0.05 + 0.015 * (ages_comp - 40)**1.1)
metabolic_lc = np.clip(metabolic_lc, 0, 1)

healthy_lc = 0.02 + 0.003 * (ages_comp - 30) + np.random.normal(0, 0.008, len(ages_comp))
healthy_lc = np.clip(healthy_lc, 0, 0.25)

ax2.plot(ages_comp, genetic_lc, 'r-', linewidth=2.5, alpha=0.7, label='High Genetic (Rx age 35)')
ax2.plot(ages_comp, metabolic_lc, 'b-', linewidth=2.5, alpha=0.7, label='Metabolic (Rx age 50)')
ax2.plot(ages_comp, healthy_lc, 'g-', linewidth=2.5, alpha=0.7, label='Healthy (Rx age 65+)')

# Mark interventions
ax2.axvline(35, color='red', linestyle=':', linewidth=2, alpha=0.5)
ax2.axvline(50, color='blue', linestyle=':', linewidth=2, alpha=0.5)
ax2.text(35, 0.55, 'Genetic\nRx', ha='center', fontsize=9, color='red', fontweight='bold')
ax2.text(50, 0.55, 'Metabolic\nRx', ha='center', fontsize=9, color='blue', fontweight='bold')

ax2.set_xlabel('Age (years)', fontsize=12, fontweight='bold')
ax2.set_ylabel('Lifetime Risk Burden', fontsize=12, fontweight='bold')
ax2.set_title('Life Course: Personalized Timing\n(Trajectory-specific Rx)', fontsize=13, fontweight='bold')
ax2.legend(fontsize=10, loc='upper left')
ax2.grid(True, alpha=0.3)
ax2.set_ylim([0, 0.6])

plt.tight_layout()
plt.show()
```

# Implementation

## Tools for Life Course Assessment

::: columns
::: {.column width="50%"}
**Available Now:**

-   **LIVE-CVD**: Lifelong benefit calculator
-   **PRS**: Clinical labs offer testing (\$50-200)
-   **CAC scoring**: Widely available imaging
-   **SCOT-HEART**: CT angiography trials

**On the Horizon:**

-   AI-enhanced imaging
-   EHR-embedded calculators
-   Continuous monitoring (wearables)
-   Multi-omics integration
:::

::: {.column width="50%"}
**Challenges:**

-   **Clinical:**
    -   EHR integration
    -   Workflow burden
    -   Training needed

-   **Payer:**
    -   Coverage for "healthy" patients
    -   Long-term value demonstration
    -   Reimbursement models

-   **Patient:**
    -   Understanding lifetime risk
    -   Motivation for prevention
    -   Health literacy
:::
:::

## Key Discussion Questions

1.  **How do we reconcile population models with individualized prediction?**

    -   When to override risk thresholds based on trajectory?

2.  **Can life-course prediction be practically implemented?**

    -   What barriers to payer coverage for preventive therapies?

3.  **How do we bridge lifetime risk with current practice?**

    -   Lower thresholds? Earlier age? Public health?

4.  **Communication:**

    -   How to make genetics, AI, imaging meaningful?

5.  **Testing strategies:**

    -   Universal CAC at age 40?
    -   Who gets genetic testing? When?

# Summary

## Key Takeaways

::: columns
::: {.column width="50%"}
**The Core Concept:**

-   Multiple trajectories to CVD
-   Genetics provides early signal
-   Information accumulates over time
-   Uncertainty decreases with data

**The Opportunity:**

-   Earlier intervention for high-risk trajectories
-   Avoid overtreatment of low-risk
-   Precision prevention timing
:::

::: {.column width="50%"}
**The Path Forward:**

-   Integrate genetics + biomarkers + imaging
-   Build clinical decision support
-   Demonstrate cost-effectiveness
-   Engage stakeholders (payers, clinicians, patients)

**The Question:**

> How do we move from 10-year risk thresholds to lifetime trajectory management?
:::
:::

## Thank You

**Sarah Urbut, MD PhD**

Massachusetts General Hospital

[surbut\@mgh.harvard.edu](mailto:surbut@mgh.harvard.edu)

**Questions for Discussion:**

1.  Which trajectories resonate with your practice?
2.  What's your biggest barrier to life course prevention?
3.  What would change your practice tomorrow?