### 2025-11-16 pyScripts/new_oct_revision/new_notebooks/Genetic_Discovery_Summary_CORRECTED.tex
Update

\documentclass[11pt]{article}
\usepackage[margin=0.8in]{geometry}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage[table]{xcolor}
\usepackage{caption}
\usepackage{multirow}

\begin{document}

\section*{Genetic Discovery Summary: 151 Genome-Wide Significant Loci}

\subsection*{Table 1: Loci Summary by Signature}

\begin{table}[h]
\centering
\caption{Genome-wide significant loci identified across 15 disease signatures with genetic signal. For Signature 5, novelty analysis was performed on all 127 genome-wide significant SNPs in the regions; other signatures show independent lead loci.}
\begin{tabular}{lp{4cm}rrp{6cm}}
\toprule
\textbf{Signature} & \textbf{Clinical Domain} & \textbf{Total} & \textbf{Unique*} & \textbf{Example Genes} \\
\midrule
\rowcolor{yellow!20}
SIG5 & Cardiovascular/Lipid & 127 & 10 & PCSK9, LPA, LDLR, APOB, APOE, PHACTR1, TRIB1, CELSR2, CDKN2B-AS1 \\
\midrule
SIG17 & GI/Colorectal & 19 & — & GATA3, BDNF-AS, COLCA1, SMAD3, SMAD7, GREM1, BMP5 \\
\midrule
SIG7 & Hypertension/Vascular & 14 & — & SH2B3, FGF5, IRF1, CNNM2, HLA-DQB1, KCNK3 \\
\midrule
SIG19 & Skin/Pigmentation & 11 & — & MC1R, TYR, SLC45A2, CASP8, TOX3, KRT5 \\
\midrule
SIG1 & Musculoskeletal & 8 & — & GDF5, PRKG1, LTBP1, TGFA, COL27A1 \\
\midrule
SIG0 & Heart Failure/Arrhythmia & 6 & — & PITX2, LPA, R3HDML, MAP3K7CL \\
SIG2 & Upper GI/GERD & 6 & — & CCKBR, NCAM1, FOXP1, CRLF1 \\
SIG3 & Vascular/Coagulation & 6 & — & F5, MSRB3, WNT7B, PLCD1 \\
SIG10 & Ophthalmologic & 6 & — & CFH, HTRA1, CDKN2B-AS1, ME3 \\
\midrule
SIG13 & Male Urogenital & 5 & — & NOTCH4, CLPTM1L, SLC24A4 \\
\midrule
SIG14 & Pulmonary/COPD & 4 & — & MUC5B, HYKK, CDH23, HHIP-AS1 \\
SIG15 & Diabetes/Metabolic & 4 & — & TCF7L2, HNF1B, UBE2E2, BTNL2 \\
\midrule
SIG8 & Female Reproductive & 3 & — & TERT, GREB1, EEFSEC \\
\midrule
SIG16 & Neurodegeneration/Sepsis & 1 & — & APOE \\
SIG18 & Gallbladder/Biliary & 1 & — & ABCG8 \\
SIG9 & Back Pain/Spine & 1 & — & NFU1 \\
\midrule
\textbf{TOTAL} & \textbf{Across 15 Signatures} & \textbf{151} & \textbf{10†} & \\
\bottomrule
\end{tabular}
\end{table}

\footnotesize{*Unique = Not present in any constituent trait GWAS within 1MB windows. Only systematically assessed for Signature 5 (cardiovascular).}

\footnotesize{†Empirically determined for Signature 5 only. Other signatures lack constituent trait overlap data for novelty assessment.}

\vspace{0.5cm}

\subsection*{Signature 5 Novelty Analysis}

\textbf{Key Finding:} When comparing Signature 5 (cardiovascular) SNPs to constituent trait GWAS for Angina, Coronary Atherosclerosis, Hypercholesterolemia, Myocardial Infarction, Acute IHD, and Chronic IHD (using 1MB windows):

\begin{itemize}
    \item \textbf{10 of 127 SNPs (7.9\%) are unique to Signature 5} — not found in any constituent trait GWAS
    \item \textbf{117 SNPs (92.1\%) overlap} with constituent traits, validating biological coherence
    \item \textbf{35 SNPs are highly pleiotropic} — associated with 4+ constituent cardiovascular traits
\end{itemize}

\textbf{The 10 Unique Signature 5 Loci} (highlighted in \colorbox{green!20}{green} in Table 2):
\begin{enumerate}
    \item rs6687726 (\textbf{IL6R}) — Interleukin-6 receptor, inflammatory pathway
    \item rs2509121 (\textbf{HYOU1}) — Hypoxia up-regulated protein, ER stress response
    \item rs4760278 (\textbf{R3HDM2}) — RNA binding protein
    \item rs1532085 (\textbf{LIPC}) — Hepatic lipase, HDL metabolism
    \item rs7168222 (\textbf{NR2F2-AS1}) — Nuclear receptor antisense RNA
    \item rs35039495 (\textbf{PLCG2}) — Phospholipase C gamma 2, platelet activation
    \item rs8121509 (\textbf{OPRL1}) — Opioid receptor, pain/stress signaling
    \item rs1499813 (\textbf{FNDC3B}) — Fibronectin domain, insulin signaling
    \item 4:96088139 (\textbf{UNC5C}) — Netrin receptor, axon guidance
    \item rs4732365 (\textbf{C7orf55}) — Unknown function
\end{enumerate}

\vspace{0.5cm}

\subsection*{Table 2: Representative Genetic Associations with Biological Validation}

\begin{longtable}{lllp{3cm}p{5.5cm}}
\caption{Selected loci demonstrating biological coherence of signature-based GWAS. \colorbox{green!20}{Green highlighting} indicates loci unique to Signature 5 (not in any constituent trait GWAS).} \\
\toprule
\textbf{Signature} & \textbf{SNP} & \textbf{Gene} & \textbf{P-value} & \textbf{Biological Context} \\
\midrule
\endfirsthead
\toprule
\textbf{Signature} & \textbf{SNP} & \textbf{Gene} & \textbf{P-value} & \textbf{Biological Context} \\
\midrule
\endhead
\midrule
\multicolumn{5}{r}{\textit{Continued on next page}} \\
\endfoot
\bottomrule
\endlastfoot

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 5: Cardiovascular/Lipid-Mediated Disease}} \\
\multicolumn{5}{l}{\footnotesize \textit{56 total loci; 10 unique to Sig5 (highlighted in green), 46 overlap with constituent traits}} \\
\midrule
SIG5 & rs10455872 & LPA & $2.8 \times 10^{-130}$ & Lipoprotein(a), strongest CAD association \\
SIG5 & rs1333042 & CDKN2B-AS1 & $9.0 \times 10^{-100}$ & 9p21 locus, CAD, MI, glaucoma (pleiotropy) \\
SIG5 & rs7412 & APOE & $1.8 \times 10^{-59}$ & APOE-$\varepsilon$2 allele, lipids, Alzheimer's \\
SIG5 & rs12740374 & CELSR2 & $1.7 \times 10^{-42}$ & Cadherin EGF LAG seven-pass, LDL-C \\
SIG5 & rs138294113 & LDLR & $5.9 \times 10^{-39}$ & LDL receptor, familial hypercholesterolemia \\
SIG5 & rs9349379 & PHACTR1 & $6.5 \times 10^{-30}$ & Phosphatase and actin regulator, CAD \\
SIG5 & rs11591147 & PCSK9 & $7.5 \times 10^{-25}$ & Proprotein convertase, LDL-C, CAD therapeutic target \\
SIG5 & rs2954031 & TRIB1 & $2.5 \times 10^{-15}$ & Tribbles pseudokinase 1, lipid metabolism \\
SIG5 & rs2119690 & LPL & $1.8 \times 10^{-10}$ & Lipoprotein lipase, triglyceride hydrolysis \\
\rowcolor{green!20}
SIG5 & \textbf{rs1499813} & \textbf{FNDC3B} & $\mathbf{1.1 \times 10^{-10}}$ & \textbf{UNIQUE: Fibronectin domain, adipogenesis, insulin signaling} \\
\rowcolor{green!20}
SIG5 & \textbf{rs2509121} & \textbf{HYOU1} & $\mathbf{1.0 \times 10^{-9}}$ & \textbf{UNIQUE: Hypoxia up-regulated protein, ER stress} \\
\rowcolor{green!20}
SIG5 & \textbf{rs4732365} & \textbf{C7orf55} & $\mathbf{1.1 \times 10^{-9}}$ & \textbf{UNIQUE: Chromosome 7 ORF, unknown function} \\
\rowcolor{green!20}
SIG5 & \textbf{rs8121509} & \textbf{OPRL1} & $\mathbf{3.7 \times 10^{-9}}$ & \textbf{UNIQUE: Opioid receptor, stress/pain signaling} \\
\rowcolor{green!20}
SIG5 & \textbf{rs6687726} & \textbf{IL6R} & $\mathbf{1.3 \times 10^{-8}}$ & \textbf{UNIQUE: Interleukin-6 receptor, systemic inflammation} \\
\rowcolor{green!20}
SIG5 & \textbf{rs35039495} & \textbf{PLCG2} & $\mathbf{2.6 \times 10^{-8}}$ & \textbf{UNIQUE: Phospholipase C, platelet activation, immunity} \\
\rowcolor{green!20}
SIG5 & \textbf{rs7168222} & \textbf{NR2F2-AS1} & $\mathbf{2.7 \times 10^{-8}}$ & \textbf{UNIQUE: Nuclear receptor antisense RNA} \\
\rowcolor{green!20}
SIG5 & \textbf{rs4760278} & \textbf{R3HDM2} & $\mathbf{3.2 \times 10^{-8}}$ & \textbf{UNIQUE: R3H domain containing 2, RNA binding} \\
\rowcolor{green!20}
SIG5 & \textbf{4:96088139} & \textbf{UNC5C} & $\mathbf{3.7 \times 10^{-8}}$ & \textbf{UNIQUE: Unc-5 netrin receptor, axon guidance} \\
\rowcolor{green!20}
SIG5 & \textbf{rs1532085} & \textbf{LIPC} & $\mathbf{3.8 \times 10^{-8}}$ & \textbf{UNIQUE: Hepatic lipase, HDL metabolism} \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 15: Diabetes and Metabolic Disease}} \\
\midrule
SIG15 & rs7903146 & TCF7L2 & $5.1 \times 10^{-28}$ & Transcription factor 7-like 2, strongest T2D association \\
SIG15 & rs11657964 & HNF1B & $2.8 \times 10^{-8}$ & Hepatocyte nuclear factor, MODY5, renal cysts \\
SIG15 & rs1496653 & UBE2E2 & $1.4 \times 10^{-8}$ & Near diabetes-associated locus \\
SIG15 & rs3806155 & BTNL2 & $3.3 \times 10^{-10}$ & MHC region, immune-metabolic interface \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 10: Age-Related Ophthalmologic Disease}} \\
\midrule
SIG10 & rs33944729 & CFH & $1.1 \times 10^{-11}$ & Complement factor H, macular degeneration \\
SIG10 & rs58077526 & HTRA1 & $6.1 \times 10^{-17}$ & Serine peptidase, age-related macular degeneration \\
SIG10 & rs7866783 & CDKN2B-AS1 & $1.9 \times 10^{-8}$ & 9p21, glaucoma, retinal disease (also in SIG5, SIG17, SIG19) \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 0: Heart Failure and Arrhythmia}} \\
\midrule
SIG0 & rs6843082 & PITX2 & $1.6 \times 10^{-17}$ & Paired-like homeodomain TF, atrial fibrillation \\
SIG0 & rs74617384 & LPA & $5.0 \times 10^{-10}$ & Lipoprotein(a), shared with SIG5 (pleiotropy) \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 17: Gastrointestinal/Colorectal}} \\
\midrule
SIG17 & rs12777423 & GATA3 & $3.9 \times 10^{-8}$ & Transcription factor, GI development \\
SIG17 & rs17309874 & BDNF-AS & $3.3 \times 10^{-11}$ & Brain-derived neurotrophic factor, gut-brain axis \\
SIG17 & rs7103178 & COLCA1 & $8.5 \times 10^{-9}$ & Colorectal cancer associated 1/2 \\
SIG17 & rs4939567 & SMAD7 & $5.4 \times 10^{-16}$ & TGF-$\beta$ signaling, colorectal cancer \\
SIG17 & rs58658771 & GREM1 & $1.3 \times 10^{-18}$ & BMP antagonist, colorectal cancer \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 14: Pulmonary/COPD}} \\
\midrule
SIG14 & rs35705950 & MUC5B & $1.3 \times 10^{-8}$ & Mucin 5B, idiopathic pulmonary fibrosis \\
SIG14 & rs72738786 & HYKK & $4.3 \times 10^{-19}$ & Near CHRNA5-CHRNA3-CHRNB4, lung function \\
SIG14 & rs6537293 & HHIP-AS1 & $4.3 \times 10^{-8}$ & Hedgehog interacting protein, COPD, lung function \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 13: Male Urogenital Disease}} \\
\midrule
SIG13 & rs460073 & CLPTM1L & $1.7 \times 10^{-14}$ & TERT-CLPTM1L region, prostate cancer, bladder cancer \\
SIG13 & rs2071278 & NOTCH4 & $3.3 \times 10^{-10}$ & MHC region, prostate associations (also in SIG3, SIG5, SIG7, SIG15, SIG17) \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 7: Hypertension and Vascular Regulation}} \\
\midrule
SIG7 & rs3184504 & SH2B3 & $1.1 \times 10^{-13}$ & SH2B adaptor protein 3, blood pressure, inflammation \\
SIG7 & rs12509595 & FGF5 & $1.6 \times 10^{-10}$ & Fibroblast growth factor 5, blood pressure \\
SIG7 & rs11242115 & IRF1 & $1.3 \times 10^{-8}$ & Interferon regulatory factor, immune-vascular \\
SIG7 & rs9275218 & HLA-DQB1 & $5.5 \times 10^{-17}$ & MHC class II (7 signatures total—maximum pleiotropy) \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 16: Neurodegeneration/Acute Illness}} \\
\midrule
SIG16 & rs429358 & APOE & $9.9 \times 10^{-14}$ & APOE-$\varepsilon$4, Alzheimer's, cognitive decline, sepsis \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 18: Gallbladder and Biliary Disease}} \\
\midrule
SIG18 & rs11887534 & ABCG8 & $1.3 \times 10^{-55}$ & ATP-binding cassette, cholesterol transport, gallstones \\
\midrule

\rowcolor{yellow!30}
\multicolumn{5}{l}{\textbf{SIGNATURE 19: Skin and Pigmentation}} \\
\midrule
SIG19 & rs1126809 & TYR & $4.7 \times 10^{-14}$ & Tyrosinase, melanin synthesis \\
SIG19 & rs16891982 & SLC45A2 & $3.5 \times 10^{-13}$ & Solute carrier, pigmentation \\
SIG19 & rs1805007 & MC1R & $1.5 \times 10^{-12}$ & Melanocortin 1 receptor, red hair, skin cancer \\

\end{longtable}

\subsection*{Key Findings}

\begin{enumerate}
    \item \textbf{151 genome-wide significant loci} identified across 15 disease signatures

    \item \textbf{Biological coherence:} Nearly all loci map to genes with established roles in signature-relevant pathways:
    \begin{itemize}
        \item SIG5 (cardiovascular): Lipid genes (PCSK9, LPA, LDLR, APOB, APOE, CELSR2, LPL, TRIB1)
        \item SIG15 (diabetes): Glucose regulation (TCF7L2, HNF1B)
        \item SIG10 (ophthalmologic): Macular degeneration genes (CFH, HTRA1)
        \item SIG14 (pulmonary): Lung function and fibrosis (MUC5B, HHIP)
        \item SIG19 (skin): Pigmentation genes (TYR, MC1R, SLC45A2)
    \end{itemize}

    \item \textbf{Novel discoveries:} \colorbox{green!20}{10 loci unique to Signature 5} not found in any constituent cardiovascular trait GWAS, including biologically relevant genes:
    \begin{itemize}
        \item \textbf{IL6R} (inflammation), \textbf{PLCG2} (platelet activation)
        \item \textbf{FNDC3B} (insulin signaling), \textbf{LIPC} (hepatic lipase, HDL metabolism)
        \item \textbf{HYOU1} (hypoxia/ER stress), \textbf{OPRL1} (stress response)
    \end{itemize}

    \item \textbf{Pleiotropy revealed:} 49 loci (32.5\%) affect multiple signatures:
    \begin{itemize}
        \item \textbf{HLA-DQB1}: 7 signatures (maximum pleiotropy)
        \item \textbf{CDKN2B-AS1}: SIG5, SIG10, SIG17, SIG19 (cardiovascular, ophthalmologic, GI, skin)
        \item \textbf{LPA}: SIG0 and SIG5 (heart failure and cardiovascular)
        \item \textbf{NOTCH4}: 6 signatures (MHC region, autoimmune-metabolic interface)
    \end{itemize}

    \item \textbf{Joint modeling power:} For Signature 5, 92.1\% of loci overlap with constituent traits (validating biology), while 7.9\% are unique discoveries—demonstrating both validation and enhanced discovery

    \item \textbf{Highly pleiotropic core:} 35 Signature 5 loci associate with 4+ cardiovascular conditions, revealing genetic architecture of disease COMBINATIONS not visible in single-trait GWAS

    \item \textbf{Phenome-wide coverage:} Genetic discoveries span cardiovascular, metabolic, pulmonary, GI, ophthalmologic, neurologic, reproductive, and dermatologic domains
\end{enumerate}

\subsection*{Detailed Example: rs1532085 (LIPC) — A Unique Signature 5 Discovery}

\begin{table}[h]
\centering
\caption{rs1532085 near LIPC demonstrates signature-based discovery of biologically meaningful loci}
\begin{tabular}{lll}
\toprule
\textbf{Analysis} & \textbf{P-value} & \textbf{Result} \\
\midrule
\rowcolor{red!10}
Angina (individual GWAS) & $p > 5 \times 10^{-8}$ & Not significant \\
\rowcolor{red!10}
MI (individual GWAS) & $p > 5 \times 10^{-8}$ & Not significant \\
\rowcolor{red!10}
Hypercholesterolemia (individual GWAS) & $p > 5 \times 10^{-8}$ & Not significant \\
\rowcolor{red!10}
Coronary atherosclerosis (individual GWAS) & $p > 5 \times 10^{-8}$ & Not significant \\
\rowcolor{red!10}
Acute IHD (individual GWAS) & $p > 5 \times 10^{-8}$ & Not significant \\
\rowcolor{red!10}
Chronic IHD (individual GWAS) & $p > 5 \times 10^{-8}$ & Not significant \\
\midrule
\rowcolor{green!30}
\textbf{Signature 5 (joint GWAS)} & $\mathbf{3.8 \times 10^{-8}}$ & \textbf{Genome-wide significant} \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Biological interpretation:} LIPC encodes hepatic lipase, a critical enzyme in HDL (high-density lipoprotein) metabolism and reverse cholesterol transport. Hepatic lipase plays multiple roles in cardiovascular disease:
\begin{itemize}
    \item Hydrolysis of triglycerides and phospholipids in HDL and LDL particles
    \item Regulation of HDL particle size and composition
    \item Modulation of hepatic lipoprotein uptake
    \item Effects on remnant lipoprotein clearance and atherosclerosis
\end{itemize}

This variant affects cardiovascular disease through \textit{distributed pleiotropic effects} across multiple manifestations (angina, MI, atherosclerosis, hypercholesterolemia)—each individually too weak to detect in single-trait GWAS, but collectively significant when modeled jointly through the cardiovascular signature.

\textbf{This exemplifies signature-based discovery:} The variant is at a well-established lipid metabolism gene (LIPC) but only detectable through joint modeling of related cardiovascular disease phenotypes. Individual cardiovascular endpoints dilute the signal, while the signature captures the aggregate lipid-mediated cardiovascular effect.

\vspace{1cm}

\subsection*{Interpretation for Reviewers}

\textbf{Novelty Framework:}
\begin{itemize}
    \item \textbf{7.9\% completely novel} (10/127 Sig5 SNPs) — not in constituent trait GWAS
    \item \textbf{92.1\% validate known biology} — overlap confirms biological coherence
    \item \textbf{27.6\% highly pleiotropic} (35/127 Sig5 SNPs) — affect 4+ cardiovascular traits
\end{itemize}

The 10 unique loci include genes with clear biological relevance (IL6R, PLCG2, FNDC3B, LIPC) that connect cardiovascular disease to inflammation, platelet function, insulin signaling, and HDL metabolism—pathways that are biologically well-established but missed when analyzing traditional cardiovascular endpoints in isolation.

\textbf{This is not statistical artifact. This is biological signal revealing disease processes and comorbidity mechanisms beyond single-disease risk factors.}

\end{document}