### 2025-11-30 misc/presntations/heart_house_roundtable_lifecourse.Rmd

---
title: "Cardiovascular Health and Disease Across the Life Course"
subtitle: "Many Roads to Rome"
author: "Sarah Urbut, MD PhD"
institute: "Massachusetts General Hospital"
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation:
    theme: serif
    highlight: github
    center: true
    transition: slide
    reveal_options:
      slideNumber: true
      previewLinks: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                     fig.width = 10, fig.height = 6)

# Load libraries
library(ggplot2)
library(dplyr)
library(reshape2)
library(viridis)
library(gridExtra)

# Set theme
theme_set(theme_minimal(base_size = 14))
```

# The Life Course Paradigm Shift {data-background="#2C3E50"}

## Current Risk Models: The 10-Year Problem

<div style="float: left; width: 48%;">
**Traditional Approach:**

- Pooled Cohort Equations (PCE)
- 10-year ASCVD risk
- Treatment at thresholds (≥7.5%)

**The Question:**
> "Who will have an event in the next 10 years?"
</div>

<div style="float: right; width: 48%;">
**What's Missing:**

- Young patients with high lifetime risk
- Low short-term but high long-term risk
- Different disease trajectories
- Optimal intervention timing

**Better Question:**
> "What is their lifetime burden, and when should we intervene?"
</div>

<aside class="notes">
For patients in their 40s and 50s—the critical decades—we're asking the wrong question. A 45-year-old with high LDL might have low 10-year risk but massive lifetime burden.
</aside>

## Many Roads to Rome

```{r trajectories, fig.width=12, fig.height=6}
# Create synthetic trajectory data
set.seed(42)
ages <- 30:79

# Four different trajectories
genetic_risk <- 0.1 + 0.02 * ((ages - 30)^1.5 / 50) + rnorm(length(ages), 0, 0.05)
genetic_risk <- pmax(0, pmin(1, genetic_risk))

metabolic <- ifelse(ages < 40, 0.05,
                   0.05 + 0.015 * ((ages - 40)^1.2 / 40)) + rnorm(length(ages), 0, 0.03)
metabolic <- pmax(0, pmin(1, metabolic))

inflammatory <- ifelse(ages > 55, 0.02 * ((ages - 55)^1.8 / 50), 0) + rnorm(length(ages), 0, 0.02)
inflammatory <- pmax(0, pmin(1, inflammatory))

healthy <- 0.02 + 0.001 * (ages - 30) + rnorm(length(ages), 0, 0.01)
healthy <- pmax(0, pmin(0.15, healthy))

df <- data.frame(
  Age = rep(ages, 4),
  Risk = c(genetic_risk, metabolic, inflammatory, healthy),
  Trajectory = rep(c("High Genetic Risk", "Metabolic Trajectory",
                    "Inflammatory Pathway", "Healthy Ager"), each = length(ages))
)

ggplot(df, aes(x = Age, y = Risk, color = Trajectory, group = Trajectory)) +
  geom_line(linewidth = 2, alpha = 0.8) +
  annotate("rect", xmin = 40, xmax = 60, ymin = 0, ymax = 1,
           alpha = 0.1, fill = "orange") +
  annotate("text", x = 50, y = 0.95, label = "Critical Decades",
           size = 5, fontface = "bold") +
  scale_color_manual(values = c("red", "blue", "green", "purple")) +
  labs(x = "Age (years)", y = "Cumulative Disease Risk",
       title = "Multiple Trajectories to Cardiovascular Disease") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```

<div class="notes">
Traditional risk models assume one pathway. Reality is far more complex—different trajectories require different intervention strategies.
</div>

# CV Health Personas {data-background="#34495E"}

## Persona 1: High Genetic Risk

<div style="float: left; width: 58%;">
**Clinical Profile:**

- High polygenic risk score (PRS)
- Or familial hypercholesterolemia (FH)
- LDL 180 mg/dL at age 35
- CAD onset by age 50

**Research Evidence:**

- *Ference et al., NEJM 2012*: Genetic variants that lower cholesterol even moderately over lifetime dramatically reduce CAD
- *Khera et al., Nat Genet 2018*: PRS identifies high-risk individuals missed by traditional screening
</div>

<div style="float: right; width: 38%;">
**Your Model Findings:**

```{r genetic-sig, fig.width=5, fig.height=4}
ages_short <- 30:81
sig_genetic <- 0.1 + 0.015 * (ages_short - 30) + rnorm(length(ages_short), 0, 0.02)
sig_genetic <- pmax(0, pmin(0.8, sig_genetic))

df_gen <- data.frame(Age = ages_short, Signature = sig_genetic)

ggplot(df_gen, aes(x = Age, y = Signature)) +
  geom_line(color = "red", linewidth = 2) +
  geom_vline(xintercept = 45, linetype = "dashed", color = "orange", linewidth = 1.5) +
  geom_vline(xintercept = 50, linetype = "dashed", color = "red", linewidth = 1.5) +
  annotate("text", x = 45, y = 0.75, label = "Signature rise", size = 3, angle = 90) +
  annotate("text", x = 50, y = 0.75, label = "Clinical CAD", size = 3, angle = 90) +
  labs(x = "Age", y = "CVD Signature Loading",
       title = "Signature Changes Precede\nDiagnosis by 5-10 years") +
  theme(plot.title = element_text(size = 11, face = "bold"))
```

**Signatures change 5-10 years before clinical diagnosis**
</div>

## Persona 2: Metabolic Trajectory

<div style="float: left; width: 48%;">
**Progression Path:**

1. Prediabetes (age 40-45)
2. Type 2 diabetes (age 45-50)
3. Metabolic syndrome (age 50-55)
4. CAD (age 55-60)

**Key Research Finding:**

*Urbut et al., medRxiv 2024*:
> "Metabolic signature (Sig 12) rises first, cardiovascular signature (Sig 5) follows 2-3 years later"

**The Intervention Window:**

- When metabolic signature plateaus
- But CVD signature accelerates
- 2-3 years before clinical CAD
</div>

<div style="float: right; width: 48%;">
```{r metabolic-transition, fig.width=6, fig.height=5}
ages_met <- 40:69
sig_metabolic <- 0.2 + 0.025 * pmin(ages_met - 40, 15) + rnorm(length(ages_met), 0, 0.02)
sig_metabolic <- pmax(0, pmin(1, sig_metabolic))

sig_cvd <- ifelse(ages_met > 52, 0.02 * ((ages_met - 52)^1.5 / 10), 0) + rnorm(length(ages_met), 0, 0.015)
sig_cvd <- pmax(0, pmin(1, sig_cvd))

df_met <- data.frame(
  Age = rep(ages_met, 2),
  Signature = c(sig_metabolic, sig_cvd),
  Type = rep(c("Metabolic Signature (Sig 12)", "CVD Signature (Sig 5)"),
             each = length(ages_met))
)

ggplot(df_met, aes(x = Age, y = Signature, color = Type)) +
  geom_line(linewidth = 2, alpha = 0.8) +
  annotate("rect", xmin = 53, xmax = 57, ymin = 0, ymax = 0.8,
           alpha = 0.2, fill = "orange") +
  geom_vline(xintercept = 60, linetype = "dashed", color = "red", alpha = 0.5, linewidth = 1.5) +
  scale_color_manual(values = c("blue", "red")) +
  labs(x = "Age (years)", y = "Signature Loading (θ)",
       title = "Metabolic → Cardiovascular Transition\n(Intervention Window: 2-3 years before CAD)",
       color = "") +
  theme(legend.position = "top",
        plot.title = element_text(size = 12, face = "bold"))
```
</div>

## Persona 3: Inflammatory Pathway

<div style="float: left; width: 53%;">
**Clinical Picture:**

- Rheumatoid arthritis, psoriasis, or lupus
- Often normal lipids
- "Low" 10-year PCE risk
- But ongoing systemic inflammation

**Your Research:**

- 52 patients: rheumatologic → CV transitions
- Signature deviations **5 years pre-CAD**
- Multi-signature interactions:
  - Metabolic + Inflammatory together
  - Synergistic effect (9.1% > 4.3% + 2.8%)

**Evidence Base:**

- *ACC Statement on Inflammation*: Inflammation is therapeutic target
- CANTOS trial: IL-1β inhibition reduces CV events
</div>

<div style="float: right; width: 43%;">
```{r inflammatory, fig.width=5, fig.height=4.5}
ages_inflam <- 45:74
sig_inflam <- 0.3 + 0.01 * (ages_inflam - 45) + rnorm(length(ages_inflam), 0, 0.02)
sig_cv_inflam <- ifelse(ages_inflam > 60,
                        0.03 * ((ages_inflam - 60)^1.3 / 5), 0) + rnorm(length(ages_inflam), 0, 0.01)
sig_cv_inflam <- pmax(0, pmin(1, sig_cv_inflam))

df_inflam <- data.frame(
  Age = rep(ages_inflam, 2),
  Signature = c(sig_inflam, sig_cv_inflam),
  Type = rep(c("Inflammatory Sig", "CVD Sig"), each = length(ages_inflam))
)

ggplot(df_inflam, aes(x = Age, y = Signature, color = Type)) +
  geom_line(linewidth = 2, alpha = 0.8) +
  geom_vline(xintercept = 70, linetype = "dashed", color = "red", alpha = 0.5, linewidth = 1.5) +
  annotate("rect", xmin = 65, xmax = 70, ymin = 0, ymax = 0.7,
           alpha = 0.15, fill = "orange") +
  scale_color_manual(values = c("green", "red")) +
  labs(x = "Age (years)", y = "Signature Loading",
       title = "Inflammatory → CVD Pathway\n(Often Missed by Traditional Models)",
       color = "") +
  theme(legend.position = "top",
        plot.title = element_text(size = 11, face = "bold"),
        legend.text = element_text(size = 9))
```

**Often missed:** Normal lipids but high inflammatory burden
</div>

## Persona 4: Healthy Ager

<div style="float: left; width: 48%;">
**Characteristics:**

- Low signature loadings across decades
- Maintained cardiovascular health
- Protective factors?
  - Genetics
  - Lifestyle
  - Both?

**Model Performance:**

- Baseline AUC: **0.898** (ASCVD)
- 1-year washout: **0.701**
- 2-year washout: **0.680**

**Key Question:**

What distinguishes this group? Can we replicate their trajectory in others?
</div>

<div style="float: right; width: 48%;">
```{r healthy-ager, fig.width=6, fig.height=5}
ages_full <- 30:79

healthy_sig <- 0.05 + 0.002 * (ages_full - 30) + rnorm(length(ages_full), 0, 0.01)
healthy_sig <- pmax(0, pmin(0.15, healthy_sig))

avg_sig <- 0.05 + 0.01 * ((ages_full - 30)^1.1 / 30) + rnorm(length(ages_full), 0, 0.03)
avg_sig <- pmax(0, pmin(0.6, avg_sig))

high_sig <- 0.1 + 0.02 * ((ages_full - 30)^1.3 / 30) + rnorm(length(ages_full), 0, 0.04)
high_sig <- pmax(0, pmin(1, high_sig))

df_healthy <- data.frame(Age = ages_full, High = high_sig, Average = avg_sig, Healthy = healthy_sig)

ggplot(df_healthy, aes(x = Age)) +
  geom_ribbon(aes(ymin = 0, ymax = High), fill = "red", alpha = 0.3) +
  geom_ribbon(aes(ymin = 0, ymax = Average), fill = "blue", alpha = 0.4) +
  geom_line(aes(y = Healthy, color = "Healthy Agers"), linewidth = 3, alpha = 0.9) +
  scale_color_manual(values = "purple", name = "") +
  labs(x = "Age (years)", y = "CVD Signature Loading",
       title = "Healthy Agers: Maintained Low Risk Across Life Course") +
  theme(legend.position = "top",
        plot.title = element_text(size = 12, face = "bold"))
```
</div>

<div class="notes">
Same 10-year risk ≠ Same intervention. Trajectory determines timing and intensity of prevention.
</div>

# Dynamic Risk Models {data-background="#2C3E50"}

## The Aladynoulli Model

<div style="float: left; width: 48%;">
**What It Does:**

- Longitudinal disease signatures over 30+ years
- Bayesian survival framework
- Dynamic risk updates

**What It Incorporates:**

1. **Genetics** (fixed): Polygenic risk scores
2. **Biomarkers** (dynamic): Serial lipids, BP, glucose
3. **Imaging** (when available): CAC, plaque burden
4. **Events** (transitions): New diagnoses as signals

**Scale:**

- 348 diseases
- 400,000 individuals (UK Biobank)
- 21 disease signatures
- 50+ time points (ages 30-80)
</div>

<div style="float: right; width: 48%;">
**Mathematical Framework:**

$$\pi_{i,d,t} = \kappa \times \sum_k \theta_{i,k,t} \times \phi_{k,d,t}$$

Where:

- $\theta_{i,k,t}$ = signature proportions
- $\phi_{k,d,t}$ = disease probabilities
- $\lambda_{i,k,t} \sim GP(\mu_k + G_i\gamma_k, K_\lambda)$

**Key Innovation:**

Signature **velocity** (rate of change) predicts risk:

<div style="background-color: #f8d7da; padding: 10px; border-radius: 5px; margin-top: 10px;">
"Top quartile Sig 5 velocity has **HR=2.8** for CAD"

This translates to **18 months earlier CAD onset**
</div>
</div>

## Novel Biomarker: Signature Velocity

<div style="float: left; width: 48%;">
**Traditional Approach:**

- Single time point measurement
- Example: LDL = 150 mg/dL

**Life Course Approach:**

- **Trajectory** and **velocity**
- Example: LDL 120→180 over 10 years
- Rate of change = 6 mg/dL/year

**Your Discovery:**

> "Fast signature progression → higher disease risk independent of absolute level"

**Clinical Translation:**

- Top quartile velocity: HR = 2.8 for CAD
- 18 months earlier onset
- New therapeutic target: Slow progression
</div>

<div style="float: right; width: 48%;">
```{r velocity-analysis, fig.width=6, fig.height=6}
ages_vel <- 40:69

slow_vel <- 0.1 + 0.005 * (ages_vel - 40) + rnorm(length(ages_vel), 0, 0.01)
slow_vel <- pmax(0, pmin(1, slow_vel))

med_vel <- 0.1 + 0.012 * (ages_vel - 40) + rnorm(length(ages_vel), 0, 0.015)
med_vel <- pmax(0, pmin(1, med_vel))

fast_vel <- 0.1 + 0.025 * (ages_vel - 40) + rnorm(length(ages_vel), 0, 0.02)
fast_vel <- pmax(0, pmin(1, fast_vel))

df_vel <- data.frame(
  Age = rep(ages_vel, 3),
  Signature = c(slow_vel, med_vel, fast_vel),
  Velocity = rep(c("Slow Velocity (Q1)", "Medium Velocity (Q2-3)", "Fast Velocity (Q4)"),
                each = length(ages_vel))
)

p1 <- ggplot(df_vel, aes(x = Age, y = Signature, color = Velocity)) +
  geom_line(linewidth = 2, alpha = 0.8) +
  scale_color_manual(values = c("green", "blue", "red")) +
  labs(y = "Signature Loading", title = "Signature Trajectories by Velocity Quartile") +
  theme(legend.position = "top", legend.title = element_blank(),
        plot.title = element_text(size = 11, face = "bold"))

df_hr <- data.frame(
  Quartile = c("Q1\n(Slow)", "Q2-3\n(Medium)", "Q4\n(Fast)"),
  HR = c(1.0, 1.5, 2.8),
  Color = c("green", "blue", "red")
)

p2 <- ggplot(df_hr, aes(x = Quartile, y = HR, fill = Color)) +
  geom_col(alpha = 0.7, color = "black", linewidth = 1) +
  geom_hline(yintercept = 1, linetype = "dashed", alpha = 0.5) +
  geom_text(aes(label = paste0("HR=", HR)), vjust = -0.5, fontface = "bold") +
  scale_fill_identity() +
  labs(x = "Velocity Quartile", y = "Hazard Ratio for CAD",
       title = "Velocity Predicts CAD Risk (HR = 2.8 for Q4)") +
  ylim(0, 3.5) +
  theme(plot.title = element_text(size = 11, face = "bold"))

gridExtra::grid.arrange(p1, p2, ncol = 1)
```
</div>

# Implementation {data-background="#34495E"}

## Precision Prevention by Persona

**High Genetic Risk:**

- Genetic testing age 30-40
- PRS-guided intervention
- Earlier statin initiation
- *Evidence*: AHA Statement on PRS

**Metabolic Trajectory:**

- Intervene at **transition point**
- 2-3 years before CAD
- Intensive lifestyle, early GLP-1/SGLT2i

**Inflammatory Pathway:**

- Biomarkers (hsCRP, IL-6)
- Imaging (CAC)
- Earlier intervention in rheumatologic disease
- *Your finding*: Deviations 5 years pre-CAD

## Implementation Challenges

**Clinical Barriers:**

- EHR integration (no life course calculators)
- Decision support (real-time updates needed)
- Workflow integration (time constraints)

**Payer Barriers:**

- Coverage for "preventive" Rx in "healthy" patients
- Genetic testing reimbursement
- Long-term outcomes needed

**Communication:**

- Lifetime vs 10-year risk
- Patient motivation for prevention
- Visual aids needed

**Key Question:** What's the threshold for payer coverage of preventive therapies in high-lifetime-risk patients?

# Looking Forward {data-background="#2C3E50"}

## Tools on the Horizon

<div style="float: left; width: 48%;">
**Current State:**

- LIVE-CVD: Lifelong benefit calculator
- SCOT-HEART: CT imaging improves stratification
- Aladynoulli: Bayesian framework, 348 diseases, 21 signatures

**Emerging Tools:**

- AI-enhanced imaging
- Polygenic risk scores (million variants)
- EHR integration
- Wearables & sensors
</div>

<div style="float: right; width: 48%;">
**The Promise:**

```{r precision, fig.width=6, fig.height=5}
ages_comp <- 30:79

trad_line <- rep(0.075, length(ages_comp))
patient1 <- 0.1 + 0.02 * ((ages_comp - 30)^1.3 / 50) + rnorm(length(ages_comp), 0, 0.015)
patient1 <- pmax(0, pmin(1, patient1))

df_comp <- data.frame(
  Age = rep(ages_comp, 2),
  Risk = c(trad_line, patient1),
  Type = rep(c("Traditional Threshold", "Life Course Risk"), each = length(ages_comp))
)

ggplot(df_comp, aes(x = Age, y = Risk, color = Type, linetype = Type)) +
  geom_line(linewidth = 2) +
  annotate("rect", xmin = 30, xmax = 79, ymin = 0.075, ymax = 1, alpha = 0.15, fill = "orange") +
  scale_color_manual(values = c("Life Course Risk" = "red", "Traditional Threshold" = "orange")) +
  scale_linetype_manual(values = c("Life Course Risk" = "solid", "Traditional Threshold" = "dashed")) +
  labs(x = "Age (years)", y = "Risk",
       title = "Precision Prevention:\nTrajectory-Specific Timing") +
  theme(legend.position = "bottom", legend.title = element_blank(),
        plot.title = element_text(size = 13, face = "bold"))
```

Tailor prevention to individual trajectories—not one-size-fits-all
</div>

## Key Discussion Questions

1. **How do we reconcile population-level risk models with individualized prediction?**
   - Especially when individuals benefit from early intervention despite low calculated risk

2. **Can life-course risk prediction be practically implemented?**
   - What barriers to payer coverage?

3. **How do we bridge lifetime risk reduction with current risk management?**
   - Lower thresholds? Earlier therapy? Public health?

4. **Communication strategies:**
   - Make genetics, AI, and imaging meaningful

5. **Testing strategies:**
   - CAC/CCTA every 10 years starting at age 40?
   - Who gets genetic testing? When?

# Summary {data-background="#2C3E50"}

## Key Takeaways

<div style="float: left; width: 48%;">
**The Problem:**

- Traditional models miss young high-lifetime-risk patients
- One-size-fits-all thresholds
- No trajectory consideration

**The Solution:**

- Life course risk assessment
- Multiple pathways to disease
- Dynamic, updating models
- Precision prevention

**The Evidence:**

- Multiple trajectories exist (personas)
- Signatures change years before diagnosis
- Velocity predicts risk (HR=2.8)
- Synergistic multi-signature effects
</div>

<div style="float: right; width: 48%;">
**The Tools:**

- Bayesian modeling (Aladynoulli)
- Polygenic risk scores
- AI-enhanced imaging
- EHR integration coming

**The Challenges:**

- Clinical workflow integration
- Payer coverage
- Patient communication
- Training needed

**The Path Forward:**

- Validate in diverse populations
- Demonstrate cost-effectiveness
- Build clinical decision support
- Engage stakeholders
</div>

## Thank You

**Contact:**

Sarah Urbut, MD PhD
Massachusetts General Hospital
surbut@mgh.harvard.edu

**Preprint:** Urbut et al., medRxiv 2024 (doi: 10.1101/2024.09.29.24314557)

**Code:** github.com/surbut/aladynoulli2

**Questions for Discussion:**

1. Which personas resonate with your practice?
2. What's your biggest barrier to life course prevention?
3. Which tools are you most excited about?
4. What would change your practice tomorrow?