### 2025-12-11 reviewer_response_todo.md
update

# Reviewer Response Completeness Check & TODO List

## ✅ ADDRESSED ITEMS

### Referee #1 Minor Details:
- ✅ Comment 4: Predictive medical research tradition - Steyerberg reference added (line 381)
- ✅ Comment 5: Interpretability and risk metrics - Addressed (lines 390-405)
- ✅ Comment 6: Temporal modeling vs age-dependent risk - Addressed (lines 407-428)
- ✅ Comment 7: Heritability estimates - Addressed with CVD comparison (lines 457-483)
- ✅ Comment 8: Joint phenotype modeling - Addressed (lines 485-490)
- ✅ Comment 9: AUC comparisons - Comprehensive comparisons added (lines 491-628)
- ✅ Comment 10: Age-specific discrimination - Clarified (lines 630-677)
- ✅ Comment 11: Figure organization - Addressed (lines 678-683)

### Referee #2 Minor Concerns:
- ✅ Cohort Definition - Detailed in Methods (lines 900-906)
- ✅ Genetic Analysis - Detailed ancestry handling (lines 907-912)
- ✅ Phenotype Handling - Comprehensive ICD/PheCode transformation details (lines 914-927)
- ✅ Analytical Decisions - Justification for 348 diseases and K=20 (lines 929-933)
- ✅ Code Availability - Fixed GitHub link (line 935)

### Referee #3 Minor Comments:
- ✅ Comment 8: Heterogeneity terminology - Clarified (line 1270)
- ✅ Comment 9: Signature count (20 vs 21) - Confirmed K=20 (line 1272)
- ✅ Comment 10: Fig 2B x-axis - Clarified (line 1274)
- ✅ Comment 11: Fig 4D clustering - Addressed (line 1276)
- ✅ Comment 12: SBayesS/LDpred - Acknowledged (line 1278)
- ✅ Comment 13: geq1000 notation - Corrected to ≥1,000 (line 1280)
- ✅ Comment 14: Fig 4B clusters - Clarified methodology (line 1282)
- ✅ Comment 15: Harrell's C - Explained why AUC is appropriate (line 1284)
- ✅ Comment 16: AUC invariance - Acknowledged limitation (line 1286)
- ✅ Comment 17: Computational intensity - Addressed (line 1288)

---

## ❌ TODO ITEMS

### 1. **Missing Reference: "Stires and Briggs"**
   - **Location**: Reviewer comment line 55 mentions "Stires and Briggs" as a book reference
   - **Current Status**: Only Steyerberg (2019) is cited in response (line 381)
   - **Action Needed**: 
     - Verify if "Stires and Briggs" is a real reference (may be a typo or alternative name)
     - If valid, add to the response alongside Steyerberg
     - If not found, confirm Steyerberg alone is sufficient (it's a well-known comprehensive text)

### 2. **Computational Hours Placeholder**
   - **Location**: Line 1288 in reviewer_response.tex
   - **Current Text**: "full UK Biobank analysis (39 subsets) required ~[X] compute hours"
   - **Action Needed**: 
     - Fill in actual compute hours or provide estimate
     - If exact number unavailable, provide reasonable estimate based on:
       - 200 epochs per subset
       - Number of subsets (39)
       - Approximate time per epoch
     - Alternative: State "approximately X hours" or provide range

### 3. **Simulation Study Mention**
   - **Location**: Simulation study exists in current.tex (lines 495-503) but not explicitly mentioned in reviewer response
   - **Current Status**: Simulation study validates model recovery but not highlighted in response
   - **Action Needed**: 
     - Consider adding brief mention in "Model Robustness and Validity" section
     - Or add to "Model Validity and Learning Dynamics" (Analysis 21)
     - Note: Not explicitly requested by reviewers, but demonstrates model validation

---

## 📊 ASSESSMENT: Code Detail & Simulation Comprehensiveness

### Code Detail: ✅ **COMPREHENSIVE**
- ✅ GitHub repository link fixed and accessible
- ✅ All 26 analysis notebooks documented
- ✅ Code structure explained (PyTorch implementation)
- ✅ Training scripts, prediction pipelines documented
- ✅ Vectorized implementation mentioned
- ✅ Computational details provided (epochs, convergence, inference time)
- ✅ Data availability commitments detailed

### Simulation Study: ✅ **ADEQUATE BUT COULD BE ENHANCED**

**Current Simulation (current.tex lines 495-503):**
- ✅ Uses model itself as generative model (good validation approach)
- ✅ Tests recovery of: cluster structure, temporal dynamics, genetic effects
- ✅ Reports: correct cluster count (5/5), Jaccard similarity (0.795)
- ✅ Demonstrates model can recover known parameters

**Potential Enhancements (not critical but would strengthen):**
- Could add: Sensitivity analysis (varying N, D, K)
- Could add: Performance under different noise levels
- Could add: Comparison with alternative initialization strategies
- **Note**: Current simulation is sufficient for validation purposes

---

## 🎯 PRIORITY ACTIONS

**HIGH PRIORITY:**
1. Fill in computational hours placeholder [X] → actual number/estimate
2. Verify/address "Stires and Briggs" reference

**LOW PRIORITY (Optional):**
3. Add brief mention of simulation study to response (if space allows)

---

## 📝 NOTES

- Most minor details are comprehensively addressed
- Methods section in current.tex appears thorough
- Simulation study validates model appropriately
- Main gaps are: computational hours number and potential missing reference